---
title: "ANALY_explr_dep_brain_males"
author: "Niamh MacSweeney"
date: "03/10/2022"
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

ABCD Exploratory Analysis Script - Brain Structure ~ brain structure -> Males

#Introduction

The purpose of this script is to re-run our pilot analyses (see ANALY_pilot... scripts) but using the same timepoints for our main analysis, pubertal timing (Y1) and depression (Y3), and imaging (Y2) to identify other brain regions that may act as mediators that were not detected in the pilot analysis. 
------------------------------------------------------------------------

####Workflow

1.  Setup: Load data, merge to create master dataframe.
2.  Pilot prep: Split into males and females.
3.  Set up  function
4.  Run models

####STEP 1: SETUP Load libraries and set wd

```{r, load libraries, set wd}


library(tidyverse)
library(lme4)
library(lmerTest)
library(nlme)
library(psych)
library(ggplot2)
library(readr)
library(pbapply)
library(gridExtra)
library(sandwich)
library(stringr)
library(kableExtra)
#library(epiDisplay)
library(gmodels)
library(caret)


setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/analy")
```

Load data cleaned and tided data

```{r,load data}

#Cleaned data in MPlus format 

targetdata <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/explr_analy_M_data.R4.0.rds")


```
## Adjust variable names 

Mplus files need to be less than 8 chars in length so we will need to add tags back onto the subcortical measure so that we can adjust for multiple comparisons in batches (e.g., cortical, subcortical etc.) 

```{r, edit subcortical names}

subcortVars <- c("V.aa", "V.amygdala",
                  "V.caudate", "V.hpus", 
                  "V.pallidum","V.putamen",
                  "V.tp", "V.vedc"
                  )

targetdata <- targetdata %>% 
  rename(Vsc.aa = V.aa,
         Vsc.amygdala = V.amygdala,
         Vsc.caudate = V.caudate,
         Vsc.hpus = V.hpus,
         Vsc.pallidum = V.pallidum,
        Vsc.putamen = V.putamen,
        Vsc.tp = V.tp,
        Vsc.vedc = V.vedc)
  

```

####STEP 2: Model set up We are using X.Shen's reg_phewasStyle_Random.R function to run the mixed effects models for both unilateral and bilateral brain regions. Note that we will use an average measure across the left and right hemispheres for bilateral brain regions. 

We need to change "src_subject_id" to "f.eid" so that function will run.

```{r, colnames}


colnames(targetdata)[1]='f.eid'
colnames(targetdata) #check it worked

#check for duplicate columns
unique_names <- unique(colnames(targetdata)) #no duplicates 

#rename site_id_y1 to site_id (should have 21 sites)
targetdata <- targetdata %>% 
  rename(site_id = site_id_y1)
colnames(targetdata)

```

Define Basic Model for pilot analysis

IV: Depressive symptoms

DV: Brain Measures

Covariates: Fixed = age, race/ethnicity; dti mean motion for (DTI models)

Random: family ID, scanner ID.

Set categorical variables

```{r, set categorical variables}

targetdata$race_y0 = as.factor(targetdata$race_y0)
targetdata$income = as.factor(targetdata$income)
targetdata$fam_id = as.factor(targetdata$fam_id)
targetdata$scnr_id = as.factor(targetdata$scnr_id)
targetdata$sex= as.factor(targetdata$sex)
targetdata$f.eid = as.factor(targetdata$f.eid)
targetdata$site_id = as.factor(targetdata$site_id)
str(targetdata)

```

##Define function 


```{r, function}

source('/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/functions/reg_phewasStyle_GLMM_outcome_brain.R')

```

#STEP 3: Define variables:

- scale variables first 

```{r, scale vars}
targetdata=targetdata
colnames(targetdata)

#scale all numeric except population weighting variables and depression variable (outcome)


#change integer variables to numeric for ease of scaling later 
intCols <- targetdata[sapply(targetdata, class) == 'integer']
intCols <- colnames(intCols)

targetdata <- targetdata %>% 
  mutate_at(intCols, as.numeric)

# Get continuous variables we want to z-scale
conCols <- targetdata[sapply(targetdata, class) == 'numeric']
conCols <- colnames(conCols)
conCols <- conCols[!conCols %in% c("pws","y_dep_y3")] 
#don't standardise outcome as it is count data. pws = population weighting variable

#make list of factor columns for ease of sorting new df later
facCols <- targetdata[sapply(targetdata, class) == 'factor']
facCols <- colnames(facCols)

# Factorise character variables, z-scale continuous, 
targetdata <- targetdata %>%
  mutate_at(facCols, as.factor) %>% 
  mutate_at(conCols, scale) %>% 
  mutate_at(conCols, as.numeric)

str(targetdata)
head(targetdata)
summary(targetdata)


```

Set up outcome (dependent variable): Brain measures

```{r, set up DVs}
#select imaging vars
ls.dep.fs.short=colnames(targetdata)[grep('^V\\.|^SA\\.|^T\\.|^SL\\.|^Vsc\\.|^FA\\.|^MD\\.',colnames(targetdata))]
ls.dep.all = ls.dep.fs.short
ls.dep.all=ls.dep.all[!duplicated(ls.dep.all)]

```

Set up Independent Variable (or factor): Brain measures

```{r, set up IVs}

ls.factor=colnames(targetdata)[grep('^pt_m',colnames(targetdata))]
```

Combine DVs and IVs so that we get a list of the DV and IV

```{r, combine DVs and IVs}

ls.dep.factor.combo=expand.grid(ls.dep.all,ls.factor,stringsAsFactors = F)

```

##Set up covariates:
For this exploratory analyses, we are running the base model

Fixed covs: age, race/ethnicity. 

Random: family, site


```{r, set up covs}

ls.models=data.frame(dependent=ls.dep.factor.combo$Var1,
                     factor=ls.dep.factor.combo$Var2,
                     covs='',stringsAsFactors = F)

ls.models$covs=paste0(c('age_y2','race_y0'),collapse='+') 
#just list fixed covariates here. Random effects are specified in function

#add dti mean motion (fd) as cov for dti models
ls.models$covs[grepl('FA\\.|MD\\.',ls.models$dependent)]=paste0(ls.models$covs[grepl('FA\\.|MD\\.',ls.models$dependent)],'+dti_mot')


```

#### STEP 4: Specify models

USER: Make sure that the model names specified below match the model names defined in function. 

```{r, specify models}

ls.models$model.est=''
ls.models$model.est[grep('hemi',ls.models$covs)]='bilateral' #we are defining here for purposes of function but we are not using the bilateral elements as we have used an average measure - therefore, the model will use lme models for all other cases. 
ls.models$model.est[ls.models$model.est=='']='lme'


```

Divide models into bulk measures and individual regions

Bulk (global) measures

```{r, bulk measures}
ls.model.bulk=ls.models[grep('\\.glb',ls.models$dependent),]
ls.model.bulk$p_batch=1

```

Individual regions

```{r, individual regions}

ls.model.region=ls.models[!grepl('\\.glb',ls.models$dependent),] 

ls.model.region$p_batch=99999
target.model=ls.model.region
ls.dep.cate=c('T\\.','V\\.','SA\\.','SL\\.', 'Vsc\\.', 'FA\\.','MD\\.') #Vsc = subcortical measures

cate.no = 1
for (dep in ls.dep.cate){
        loc = grepl(dep,target.model$dependent)
        target.model$p_batch[loc]=cate.no
        cate.no=cate.no+1
}

ls.model.region=target.model


#Check impact of WBV. 
# for individual cortical and subcort region models add whole brain volume as a covariate
ls.model.region.covWholeB=ls.model.region
ls.model.region.covWholeB$covs[!grepl('FA\\.|MD\\.',ls.model.region.covWholeB$dependent)]=
      paste0(ls.model.region.covWholeB$covs[!grepl('FA\\.|MD\\.', ls.model.region.covWholeB$dependent)],'+ WBV')

```

#### STEP 5: Run Models

```{r, run models}

#Global (bulk) models
result.males.pt_brain.bulk=reg_phewasStyle(ls.model.bulk,dat_short=targetdata, correctByFactor = T)

#Regional (Individual) models
result.males.pt_brain.region=reg_phewasStyle(ls.model.region,dat_short=targetdata, correctByFactor = T)

#covary for WBV
result.males.pt_brain.region.covWholeB=reg_phewasStyle(ls.model.region.covWholeB,dat_short=targetdata, correctByFactor = T)

#add TRUE column if p.corrected < 0.002 which means that all values of ≤0.001

#global results - none significant at this threshold
result.males.pt_brain.bulk$Sig <- as.numeric((as.character(result.males.pt_brain.bulk[,"p.corrected"]))) < 0.002


#regional results
result.males.pt_brain.region$Sig <- as.numeric((as.character(result.males.pt_brain.region[,"p.corrected"]))) < 0.002

#cov for whole brain 
result.males.pt_brain.region.covWholeB$Sig <-  as.numeric((as.character(result.males.pt_brain.region.covWholeB[,"p.corrected"]))) < 0.002

#Save as R.Data file
save(result.males.pt_brain.bulk,result.males.pt_brain.region,file='/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/exploratory_analysis/males_pt_brain_explr.RData')

#Save results as .csv files

#global results
write.csv(result.males.pt_brain.bulk, file = '/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/exploratory_analysis/sig.males.pt.brain.bulk.csv')

#regional results 
write.csv(result.males.pt_brain.region, file = '/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/exploratory_analysis/sig.males.pt.brain.regional.csv')

write.csv(result.males.pt_brain.region.covWholeB, file = '/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/exploratory_analysis/sig.males.pt.brain.regional.cov.WholeB.csv')

```


Double check function outputs -- looks good! 
```{r, test function}

CCfitBase <- lmer(Vsc.vedc ~ pt_m + age_y3 + race_y0
                            + (1|scnr_id) + (1|fam_id),
                            control = lmerControl(optimizer ="bobyqa"),
                            data = targetdata,
                            na.action = na.exclude
                       )
summary(CCfitBase)

```


 \#\#\#\#\#\#\#\#\#\#\#\#\#\# End of Script \#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#
