---
title: "Brain Heatmaps for manuscript"
author: Niamh MacSweeney"
date: "21/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

We will use this script to visualise the brain associations for our Puberty-Depression RR.

#Set Up

Load in packages
```{r, ggseg}

library(tidyverse)
library(ggseg)
library(gridExtra)
library(hrbrthemes) #nice ggplot themes
library(gcookbook)


setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs")


```
#General introduction to GGSEG - what can it help us do?

## hemisphere stacking
we can stack hemisphere if we want

```{r, stacked}
ggseg(position="stacked")
```
Now we want to colour in our segmentated area

```{r, fill}
#here we see the novelty of the underlying code, basically works like a GGplot syntax :) 
ggseg(mapping=aes(fill=region))
      

```

```{r, fill}
#we can also colour and fill but avoid showing the legend from prev. figure and add in a title :) 
ggseg(mapping=aes(fill=region), position="stacked", show.legend = F) +
  ggtitle("first atlas")
  
  # +theme_darkbrain() #can add this on if you want a dark background
```
So this atlas quite nicely helps us view cortical areas. But what about subcortical? Well there is the ASEG
subcortical atlas

```{r, ASEG/SUBCORT}
ggseg(atlas=aseg,mapping=aes(fill=region)) +
  ggtitle("Automatic subcortical segmentation")
```

#Visualise ROIs 

##Pubertal timing ~ Brain associations that were significant at p (FDR corrected) ≤0.001

###FEMALES
Group by cortical measurement type:

1. Cortical thickness
2. Cortical volume
3. Sulcal Depth 
4. Subcortical regions - ventral diencephalon. 


### Cortical thickness; region, beta value (standardised)
PDS-Brain

paracentral -0.12
rostral middle frontal -0.11
insula -0.10
caudal middle frontal -0.09
superior frontal -0.09
lingual -0.08
inferior parietal -0.08
superior parietal -0.08

### Volume, region, beta value (standardised)
bankssts	-0.09
caudal middle frontal -0.09
insula -0.09
paracentral	-0.09
precuneus	-0.09
rostral middle frontal	-0.09
superior frontal	-0.09
superior parietal	-0.08
middle temporal	-0.08
fusiform	-0.08


## Subcortical - not available in ggseg parcellation. 
Accumbens area	-0.09


###Males
Only significant associations is ventral diecephalon (0.08)


```{R, add results}
#make dataframe with regions, Beta values and group type
ptCorticalResults = tibble(
              region=c("paracentral", "rostral middle frontal", "insula", "caudal middle frontal", "superior frontal",
                       "lingual", "inferior parietal", "superior parietal", 
                       "bankssts", "caudal middle frontal", "insula", "paracentral", "precuneus", 
                       "rostral middle frontal", "superior frontal", "superior parietal", "middle temporal",
                       "fusiform"),
              beta= c(-0.12,  -0.11, -0.10, -0.09, -0.09, -0.08, -0.08, -0.08,  -0.09, -0.09, -0.09, -0.09, -0.09,
                      -0.09,	-0.09, -0.08, -0.08, -0.08),
              group = c("thickness", "thickness", "thickness","thickness","thickness",
                        "thickness", "thickness", "thickness", 
                        "volume", "volume", "volume", "volume", "volume",
                        "volume", "volume", "volume", "volume", "volume")
              )
    #stringsAsFactors = FALSE


```

Plot
```{r, plot puberty ROIs}

###Cortical ROIs #####
#via ggseg
ptCorticalResults %>% 
  group_by(group) %>% 
  ggseg(atlas = dk, 
      colour = "black",
      size = .2, 
      hemi = "right",
      position = "stacked",
      mapping = aes(fill = beta)) +
  scale_fill_gradient(low="deepskyblue",high="midnightblue") +
  labs(title = "Females: Cortical regions associated with earlier pubertal timing") +
  theme_custombrain(plot.background = "white",text.colour = "black", text.size = 10, text.family = "mono") +
  theme(plot.margin=unit(c(2,2,2,3),'mm'),
        plot.title = element_text(lineheight=1, face="bold", vjust=2, hjust=0.5,size=12)) +
  facet_wrap(~group)
ggsave("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/females_expl_PT_cortical_ROI_heatmap.png")


```



