---
title: "PREP_pt_expl.Rmd"
author: "Niamh MacSweeney"
date: "22/09/2022"
output:
  pdf_document: default
  html_document: default
---
#### Introduction

The purpose of this script is to:
tidy and score the pubertal development scale (PDS) data (caregiver report) from ABCD Release 4.0
to prepare a final, clean, sample for use in models. 



We will also be creating a pubertal timing measure for the gonadal and adrenal PDS items to run exploratory analysis in our RR. From our main analysis, we see that earlier pubertal timing is associated with increased depression (Beta = 0.146, p=<0.001). We would like to see whether any of the individual items are driving this association. 

## SETUP
Load libraries needed and set working directory

```{r, libraries}

library(tidyr)
library(dplyr)
library(tidyverse)
library(psych)
library(ggplot2)
library(ggstatsplot)
library(GGally)
library(stringr)
library(outliers)
library(jtools) 
library(readr)
library(gridExtra)
library(cowplot)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")
getwd()

```


```{r, load data}

pds <-rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/Physical_Health/abcd_ppdms01.rds")

#load in puberty category data
pub_cat <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/Physical_Health/abcd_ssphp01.rds")

head(pds, n=10) #first few observations
str(pds)  #types of variables

#Convert age in months to years, save new variable
pds <- pds %>%
  mutate(
    age_years = interview_age/12 
  )

colnames(pds) #check new variable exists

#extract puberty category variables from pub_cat

pub_cat <- pub_cat %>% 
  select(src_subject_id, eventname, sex, pds_p_ss_female_category, pds_p_ss_male_category) %>% 
  filter(eventname == "1_year_follow_up_y_arm_1") #reduce to year 1


```


```{r, reduce to Year 1}
# N=11,225
 pds_y1 <- pds %>% 
  filter(eventname == "1_year_follow_up_y_arm_1")

#merge pub_cat with pds_y1

pds_y1 <- pds_y1 %>% 
  left_join(y=pub_cat, by= c("src_subject_id", "sex"))

```

CLEAN DATA
#STEP 1: Column formatting, missing values, male/female item mismatches 

```{r, column formatting}
#check for sex NAs
pds_y1 %>% 
  dplyr::count(sex) # no NAs detected

#Check that males have answered male items, females answered female items
#Note: pds_m5 and pds_m4 mean should equal 0 for females, and pds_f4 and pds_f5b mean should equal 0 for males. 

describeBy(pds_y1, group=pds_y1$sex) 

#remove one mismatch - one female participant answered male PDS questions
#lines below will remove any participant that had "F" as sex and did not have NA for the male PDS Qs. 
c(which(pds_y1$sex == "F" & !is.na(pds_y1$pds_m4_p) | !is.na(pds_y1$pds_5b_p)))
 pds_y1 <- pds_y1[-c(which(pds_y1$sex == "F" & !is.na(pds_y1$pds_m4_p))),]
 
 #N = 11244
 
#check it worked -- it did. 
describeBy(pds_y1, group=pds_y1$sex)


#Final sample for Year 1 PDS caregiver report: MALES = 5872, FEMALES = 5352, TOTAL N = 11224
xtabs(~pds_y1$sex)

```

Current sample for PDS caregiver report: MALES = 6196, FEMALES = 5680, TOTAL N = 11876

#STEP 2: Prepare data for scoring - remove "I don't know = "999" values

```{r, remove missing values}

#Inspect variables for "999" values
table(pds_y1$pds_1_p,pds_y1$sex)
table(pds_y1$pds_2_p,pds_y1$sex)
table(pds_y1$pds_3_p,pds_y1$sex)
table(pds_y1$pds_m5_p,pds_y1$sex)
table(pds_y1$pds_m4_p,pds_y1$sex)
table(pds_y1$pds_f4_p,pds_y1$sex)
table(pds_y1$pds_f5b_p,pds_y1$sex)

#PDS Recode: Change "999 = I don't know" to NA for these variables only as it is not numerically meaningful. 
pds_y1$pds_1_p[pds_y1$pds_1_p== 999] <- NA
pds_y1$pds_2_p[pds_y1$pds_2_p== 999] <- NA
pds_y1$pds_3_p[pds_y1$pds_3_p== 999] <- NA
pds_y1$pds_m5_p[pds_y1$pds_m5_p== 999] <- NA
pds_y1$pds_m4_p[pds_y1$pds_m4_p== 999] <- NA
pds_y1$pds_f4_p[pds_y1$pds_f4_p== 999] <- NA
pds_y1$pds_f5b_p[pds_y1$pds_f5b_p== 999] <- NA

#Recheck tables to see 999 has been changed to NA - looks good!
table(pds_y1$pds_1_p,pds_y1$sex)
table(pds_y1$pds_2_p,pds_y1$sex)
table(pds_y1$pds_3_p,pds_y1$sex)
table(pds_y1$pds_m5_p,pds_y1$sex)
table(pds_y1$pds_m4_p,pds_y1$sex)
table(pds_y1$pds_f4_p,pds_y1$sex)
table(pds_y1$pds_f5b_p,pds_y1$sex)

```

#STEP 3: Examine correlation PDS items with age
```{r, examine data}

#STEP 1: Examine correlation of raw data for PDS caregiver report
#Split df by sex and examine valid PDS variables for each.
pds_m <- filter(pds_y1, sex== "M")
pds_f <- filter(pds_y1, sex== "F")

#Group male PDS correlations with age
pds_m_tmp=c("pds_1_p", "pds_2_p", "pds_3_p", "pds_m5_p", "pds_m4_p", "age_years")
pds_age_m =ggpairs(pds_m, pds_m_tmp, title = "Within PDS Males")
pds_age_m

#Group female PDS correlations with age
pds_f_tmp=c("pds_1_p", "pds_2_p", "pds_3_p", "pds_f4_p", "pds_f5b_p", "age_years")
pds_age_f =ggpairs(pds_f, pds_f_tmp, title = "Within PDS Females")
pds_age_f


```

Correlation of PDS items with eachother 

```{r, corr plot}

library(corrplot)
#females
corColsF <- pds_f %>% 
  dplyr::select(pds_1_p, pds_2_p, pds_3_p, pds_f4_p, pds_f5b_p)
F <- cor(corColsF, use = "pairwise.complete.obs")
corrplot(F)

#males
corColsM <- pds_m %>% 
  dplyr::select(pds_1_p, pds_2_p, pds_3_p, pds_m4_p, pds_m5_p)
M <- cor(corColsM, use = "pairwise.complete.obs")
corrplot(M)

```


#STEP 4:PDS TOTAL SCORE (5 OUT OF 5 VARIABLES NEEDED)

First, select valid IDs

```{r, select valid IDs for PDS total}

#Sum PDS scores across rows for males. Then change female values for this field to NA
pds_y1$pds_tot_m <-rowSums(pds_y1[,c("pds_1_p", "pds_2_p", "pds_3_p", "pds_m5_p", "pds_m4_p")], na.rm=T)
pds_y1$pds_tot_m[pds_y1$pds_tot_m==0] <-NA
pds_y1$pds_tot_m[pds_y1$sex== "F"] <- NA

#Sum PDS scores across rows for females. Then change male values for this field to NA
pds_y1$pds_tot_f <-rowSums(pds_y1[,c("pds_1_p", "pds_2_p", "pds_3_p", "pds_f4_p", "pds_f5b_p")], na.rm=T)
pds_y1$pds_tot_f[pds_y1$pds_tot_f==0] <-NA
pds_y1$pds_tot_f[pds_y1$sex=="M"] <- NA

#Count number of Qs per subject in males that are not NA/DK
pds_y1$pds_m_Qcount <- 
rowSums(!is.na(pds_y1[,c("pds_1_p", "pds_2_p", "pds_3_p","pds_m5_p", "pds_m4_p")]))

#Change female values to NA for this Qcount_m col.
pds_y1$pds_m_Qcount[pds_y1$sex=="F"] <- NA 

#Count number of Qs per subject in females that are not NA/DK
pds_y1$pds_f_Qcount <- 
rowSums(!is.na(pds_y1[,c("pds_1_p", "pds_2_p", "pds_3_p", "pds_f4_p", "pds_f5b_p")]))

#Change male values to NA for this Qcount_f col.
pds_y1$pds_f_Qcount[pds_y1$sex=="M"] <- NA 

#This will give total number of NAs in males per pds column. 
apply(subset(pds_y1, sex=="M")[,c("pds_1_p", "pds_2_p", "pds_3_p","pds_m5_p", "pds_m4_p")],
      2, function(x) sum(is.na(x)))
#And total number of NAs in females per pds column
apply(subset(pds_y1, sex=="F")[,c("pds_1_p", "pds_2_p", "pds_3_p", "pds_f4_p", "pds_f5b_p")], 
      2, function(x) sum(is.na(x)))

#If any NAs for males. Then code to invalid.
pds_y1$pds_m_pt_valid <- NA
pds_y1$pds_m_pt_valid <- ifelse(pds_y1$sex=="M"
      & pds_y1$pds_m_Qcount==5,1,0)  #5 denotes the number of PDS variables needed, 1=valid, 0= invalid.

#If any NAs for females. Then code to invalid.
pds_y1$pds_f_pt_valid <- NA
pds_y1$pds_f_pt_valid <- ifelse(pds_y1$sex=="F" 
      & pds_y1$pds_f_Qcount==5,1,0)  #5 denotes the number of PDS variables needed, 1=valid, 0= invalid. 


#Count number of valid participants for males (N= 5440) and females (N=4960)
pds_y1 %>% dplyr::count(pds_m_pt_valid)
pds_y1%>% dplyr::count(pds_f_pt_valid)

##Remember total sample for males and females
xtabs(~pds_y1$sex) #We will use this to determine how many males and females were excluded
```

Total males valid = 5440 (Male total: 5872) Excluded = 432
Total females valid = 4960 (Female total: 5352) Excluded = 392
Total excluded = 824 

Second, generate PDS total score for valid IDs

```{r, generate PDS total}

#Calculate PDS total for males for only those with valid data (see definition of valid above)
pds_y1$pds_tot_m <- pds_y1$pds_tot_m
pds_y1$pds_tot_m[pds_y1$pds_m_pt_valid==0] <- NA

#Calculate PDS total for females for only those with valid data (see definition of valid above)
pds_y1$pds_tot_f <- pds_y1$pds_tot_f
pds_y1$pds_tot_f[pds_y1$pds_f_pt_valid==0] <- NA


#Create column for PDS total for males and females together
pds_y1 <- pds_y1 %>% 
  mutate(pds_tot_all = if_else(is.na(pds_tot_m), pds_tot_f, pds_tot_m))


#Generate new dataframe with complete cases for pubertal timing --- N =10,400
#This will allow us to save the linear model outputs as a new column. 
pds_timing_y1 <- pds_y1 %>% select(src_subject_id,sex, age_years, 
                                pds_m_pt_valid , pds_f_pt_valid, 
                                pds_tot_m, pds_tot_f, pds_tot_all,
                                pds_1_p,  pds_2_p,  pds_3_p, pds_m5_p, pds_m4_p,
                                pds_f4_p, pds_f5b_p, pds_p_ss_female_category, pds_p_ss_male_category) #keep individual PDS items

#Remove participants with invalid data for generating pubertal timing score. 
pds_timing_y1 <- pds_timing_y1[-c(which(pds_timing_y1$sex == "M" & pds_timing_y1$pds_m_pt_valid == 0), 
                    which(pds_timing_y1$sex == "F" & pds_timing_y1$pds_f_pt_valid == 0)),]


#Get summary stats (e.g., mean, range) for year 1 so that we can compare it to baseline

pds_timing_y1 %>% 
  group_by(sex) %>% 
  summary()
```

Total number of valid participant for PDS total = 10,400 (This matches above total calculations for M and F separately)

Year 1 summary stats: Mean PDS total for Males = 7.763; Mean PDS total for Females = 10.42


#STEP 5:GONADAL + ADRENAL MEASURE

Gonadal items map onto sex hormones: estradiol and testosterone
Adrenal items map onto DHEA

Gonadal PDS items:
Females: Growth spurt, breast development and menarche (pds_1_p, pds_f4_p, pds_f5b_p)
Males: Growth spurt, deepening of voice, facial hair (pds_1_p, pds_m4_p, pds_m5_p)

Adrenal items (both sexes): Pubic and body hair, skin changes (pds_2_p, pds_3_p)

```{r, generate gonadal and adrenal measures}

#gonadal
pds_timing_y1<- pds_timing_y1 %>% 
  mutate(gonad_avg = case_when(sex == "F" ~ rowMeans(pds_timing_y1[, c("pds_1_p", "pds_f4_p", "pds_f5b_p")]),
                               sex == "M" ~ rowMeans(pds_timing_y1[, c("pds_1_p", "pds_m4_p", "pds_m5_p")])
  ))

#adrenal
pds_timing_y1<- pds_timing_y1 %>% 
  mutate(adren_avg = case_when(sex == "F" ~ rowMeans(pds_timing_y1[, c("pds_2_p", "pds_3_p")]),
                               sex == "M" ~ rowMeans(pds_timing_y1[, c("pds_2_p", "pds_3_p")])
  ))


```

Let's plot the distribution of PDS total score, average gonadal and adrenal score for males and females using bar plots (discrete data) or density plots (continuous data)

```{r, plot distributions}

#convert sex variable to factor

colnames(pds_timing_y1)

pds_timing_y1$sex <- as.factor(pds_timing_y1$sex)

#distribution looks to be heavily skewed, which is to be expected given sample.

#males
pds_tot_bar <- ggplot(data=pds_timing_y1, aes(x= pds_tot_m)) +
  geom_bar(fill = "#00BFC4") +
  ggtitle("PDS total males") +
  labs(x = "Total PDS") +
  theme(plot.title = element_text(face = "bold"))
print(pds_tot_bar)

#Females
pds_tot_bar <- ggplot(data=pds_timing_y1, aes(x= pds_tot_f)) +
  geom_bar(fill = "#F8766D") +
  ggtitle("PDS total females") +
  labs(x = "Total PDS") +
  theme(plot.title = element_text(face = "bold"))
print(pds_tot_bar)

#show on same bar plot 
pds_tot <- ggplot(pds_timing_y1, aes(pds_tot_all))

pds_tot + geom_bar(aes(fill = sex)) +
  ggtitle("PDS total score") +
  labs (x = "Total PDS") +
  theme(plot.title = element_text(face = "bold"))
pds_tot


#make histogram of gonadal and adrenal development within each group 
gonad_hist <- pds_timing_y1 %>%
  ggplot(aes(x=gonad_avg, color=sex, fill=sex)) +
  geom_histogram(bins = 7, alpha = 0.5, position = "identity") +
  ggtitle("Gonadal score") +
   labs (x = "Gonadal PDS") +
  theme(plot.title = element_text(face = "bold"))
gonad_hist

adren_hist <-  pds_timing_y1 %>%
  ggplot(aes(x=adren_avg, color=sex, fill=sex)) +
  geom_histogram(bins = 7, alpha = 0.5, position = "identity") +
  ggtitle("Adrenal score") +
   labs (x = "Adrenal PDS") +
  theme(plot.title = element_text(face = "bold"))
adren_hist
  
#tidy pds category data

cat_df <- pds_timing_y1 %>% mutate(pds_cat = coalesce(pds_p_ss_female_category, pds_p_ss_male_category)) %>%
         select(src_subject_id, pds_cat, sex)

str(cat_df)
cat_df$pds_cat <- as.factor(cat_df$pds_cat)

#get levels in factor
levels <- cat_df %>% 
  group_by(pds_cat) %>%
  summarise(no_rows = length(pds_cat))

#rename factor levels 
cat_df$pds_cat <- recode_factor(cat_df$pds_cat, 
                      `1` = "Pre",
                      `2` = "Early",
                      `3` = "Mid",
                      `4` = "Late",
                      `5` = "Post",
                      `NA` = "NA"
)

category_hist <- cat_df %>%
  ggplot(aes(x=pds_cat, color=sex, fill=sex)) +
  geom_histogram(bins = 7, alpha = 0.5, position = "identity", stat="count") +
  ggtitle("PDS Category") +
  theme(plot.title = element_text(face = "bold"))
category_hist

```


Step 6:  GENERATE PUBERTAL TIMING SCORE for Gonadal and Adrenal and total PDS score

PDS score is regressed on age for males and females separately.
The standardised residual obtained is used as the pubertal timing score.


### Pubertal timing total score  (let's still run this so we can compare)
```{r, pt PDS total}

#### Male pubertal timing linear model ####
#model summary
pt_m_lm <- lm(pds_tot_m~age_years, na.action=na.exclude, data = pds_timing_y1) # create linear model, and specify na.exclude so that df size stays the same.
summ(pt_m_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 
#Note: the resid function in lm allows us to have the desired NA values for female rows. 

pds_timing_y1$pt_m <- resid(pt_m_lm)
summary(pds_timing_y1$pt_m)

org_m_QQplot <- plot(pt_m_lm, which= c(2)) #get QQplot


#### Female pubertal timing linear model ####
#model summary
pt_f_lm <- lm(pds_tot_f~age_years, na.action=na.exclude, data = pds_timing_y1) # create linear model

summ(pt_f_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default )

#obtain standardised residual = measure of pubertal timing. Save as new variable. 
pds_timing_y1$pt_f <- resid(pt_f_lm)
summary(pds_timing_y1$pt_f)

org_f_QQplot <- plot(pt_f_lm, which= c(2))

```

Gonadal and adrenal timing
```{r, pt gonadal and adrenal average}

#generate separate male and female gonadal items before generating timing measure

pds_timing_y1$gonad_m <- NA
pds_timing_y1$gonad_m <- pds_timing_y1

pds_timing_y1 <- pds_timing_y1 %>% 
  mutate(gonad_f = case_when(sex == "F" ~ rowMeans(pds_timing_y1[, c("pds_1_p", "pds_f4_p", "pds_f5b_p")])),
         gonad_m = case_when(sex == "M" ~ rowMeans(pds_timing_y1[, c("pds_1_p", "pds_m4_p", "pds_m5_p")])),
         adren_f = case_when(sex == "F" ~ rowMeans(pds_timing_y1[, c("pds_2_p", "pds_3_p")])),
         adren_m = case_when(sex == "M" ~ rowMeans(pds_timing_y1[, c("pds_2_p", "pds_3_p")]))
         )

#GONADAL TIMING

#males 
pt_m_lm <- lm(gonad_m ~age_years, na.action=na.exclude, data = pds_timing_y1) # create linear model, and specify na.exclude so that df size stays the same.
summ(pt_m_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 
#Note: the resid function in lm allows us to have the desired NA values for female rows. 

pds_timing_y1$gonad_pt_m <- resid(pt_m_lm)
summ_g_m <- summary(pds_timing_y1$gonad_pt_m)
summ_g_m

org_m_QQplot <- plot(pt_m_lm, which= c(2)) #get QQplot


#Females
pt_f_lm <- lm(gonad_f ~age_years, na.action=na.exclude, data = pds_timing_y1) # create linear model, and specify na.exclude so that df size stays the same.
summ(pt_f_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 
#Note: the resid function in lm allows us to have the desired NA values for female rows. 

pds_timing_y1$gonad_pt_f <- resid(pt_f_lm)
summ_g_f <- summary(pds_timing_y1$gonad_pt_f)
summ_g_f 

org_m_QQplot <- plot(pt_f_lm, which= c(2)) #get QQplot

## ADRENAL TIMING 

#Males 
pt_m_lm <- lm(adren_m ~age_years, na.action=na.exclude, data = pds_timing_y1) # create linear model, and specify na.exclude so that df size stays the same.
summ(pt_m_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 
#Note: the resid function in lm allows us to have the desired NA values for female rows. 

pds_timing_y1$adren_pt_m <- resid(pt_m_lm)
summ_g_m <- summary(pds_timing_y1$adren_pt_m)
summ_g_m

org_m_QQplot <- plot(pt_m_lm, which= c(2)) #get QQplot

#Females 
pt_f_lm <- lm(adren_f ~age_years, na.action=na.exclude, data = pds_timing_y1) # create linear model, and specify na.exclude so that df size stays the same.
summ(pt_f_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 
#Note: the resid function in lm allows us to have the desired NA values for female rows. 

pds_timing_y1$adren_pt_f <- resid(pt_f_lm)
summ_g_f <- summary(pds_timing_y1$adren_pt_f)
summ_g_f

org_m_QQplot <- plot(pt_f_lm, which= c(2)) #get QQplot

#check new dataframe 
head(pds_timing_y1)


```

STEP 7: Export clean data for Year 1. 

Change working directory and save clean file there. 
Note: clean files should be saved in data folder in ABCD_puberty_RR project folder.
```{r, export cleaned data}

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data")
saveRDS(pds_timing_y1,"pds_timing_y1_R4.0.rds")

```

####------------ END OF MAIN SCRIPT -----------------####




















###Outliers 

Outliers were only present for the original data, outlier sensitivity analysis not necessary for log10 and INT transformations. 

Now let's examine outliers (±5 SD from mean)
Note: We are defining outliers as ±5SD from mean as with a sample this large (N= ~10,000), 
we should expect some responses greater than 3SD (commonly used outlier cutoff) if the data are really normally distributed. 

As sensitivity analysis, we will run models with and without ouliers removed for PDS.
This will be for males only as no outlier detected for females. 
```{r, outliers}
#Males
outl_pds_m_upper <- (mean(pds_timing$pds_tot_m, na.rm=T) + 5*(sd(pds_timing$pds_tot_m, na.rm=T))) # = 16.45
outl_pds_m_lower <- (mean(pds_timing$pds_tot_m, na.rm=T) - 5*(sd(pds_timing$pds_tot_m, na.rm=T))) # = -2.05

#Plot outliers. Focus on +5SD from mean as negative values that are -5SD aren't valid here. 
hist(pds_timing$pds_tot_m, main="", xlab="PDS TOTAL_MALES")
abline(v=outl_pds_m_upper,col="red")
abline(v=outl_pds_m_lower,col="red")

outl_pds_m <- pds_timing %>% 
  filter(pds_tot_m > outl_pds_m_upper) #There are 10 outlier values

#create new variable with outliers removed for later sensitivity analysis
#duplicate pds_tot_m and name is pds_tot_m_no_outl
pds_timing <- pds_timing %>% 
  mutate(pds_tot_m_no_outl = pds_tot_m) 

#Change outlier values for pds_tot_m_no_outl to NA
#tried to do with dplyr syntax below and couldn't get working
pds_timing$pds_tot_m_no_outl[pds_timing$pds_tot_m_no_out > outl_pds_m_upper] <- NA

#dplyr syntax
#pds_timing <- pds_timing %>% 
  #mutate(pds_tot_m_no_outl = na_if(pds_tot_m_no_outl, pds_total_males_p > outl_pds_m_upper))

#check it worked - it did! 
outl_pds_m_check <- pds_timing %>% 
  filter(pds_tot_m_no_outl > outl_pds_m_upper)

#Females
outl_pds_f_upper <- (mean(pds_timing$pds_tot_f, na.rm=T) + 5*(sd(pds_timing$pds_tot_f, na.rm=T))) # = 21.88
outl_pds_f_lower <- (mean(pds_timing$pds_tot_f, na.rm=T) - 5*(sd(pds_timing$pds_tot_f, na.rm=T))) # = -4.13

#Plot outliers. Focus on +5SD from mean as negative values that are -5SD aren't valid here. 
hist(pds_timing$pds_tot_f, main="", xlab="PDS TOTAL_FEMALES")
abline(v=outl_pds_f_upper,col="red")
abline(v=outl_pds_f_lower,col="red")

#There are no outlier values for females at threshold of 5SD. 

#another method using dplyr
#pds_out_m_tmp <- pds_timing %>% 
#select(pds_total_males_p) %>% 
#filter(mean(pds_total_males_p) + 5*(sd(pds_total_males_p)) | mean(pds_total_males_p) - 5*(sd(pds_total_males_p))) %>% 
# summarise_each(funs(mean(., na.rm = TRUE)))

```

For sensitivity analysis with outliers removed (original data only)
``` {r, male pt, outliers removed}

pt_m_lm_no_outl <- lm(pds_tot_m_no_outl~age_years, na.action=na.exclude, data = pds_timing) # create linear model, and specify na.exclude so that df size stays the same.
summ(pt_m_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 
#Note: the resid function in lm allows us to have the desired NA values for female rows. 

pds_timing$pt_m_no_outl <- resid(pt_m_lm)
summary(pds_timing$pt_m_no_outl)

```








