---
title: "PREP_covs_R3.0"
author: "Niamh MacSweeney"
date: "5/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The purpose of this script is to tidy and quality control the covariates for our RR analysis

Covariates: Age, BMI, race/ethnicity, family income, site (for not imaging models), scanner_id (for imaging models), parental current mood, population weighting variable, family id. 

This data frame was downloaded from Release 4.0 and the DEAP portal which works off the ABCD data release (3.0) for the BMI and income variables. 

## Setup

Load libraries, set working directory.

```{r, setup}

library(psych)
library(tidyverse)
library(ggstatsplot)
library(readr)
library(gridExtra)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")
getwd()

```

##LOAD DATA

Note: Renaming variables to make the names more intuitive. 

Need to insert table here of all the covariates used.... 

```{r, covariate table}

vars <- factor(c("src_subject_id", "eventname", "sex", "age_years", "bmi", "race.6level", "demo_comb_income_p",  "abcd_site", "high.educ", "fsqc_qu_motion", "asr_scr_depress_r", "rel_family_id"))

var_lables <- factor(c("Participant ID", "Data collection wave", "Sex at birth", "Age in years", "Body Mass Index", "Race (6-level categorisation", "Household income", "Study site", "Highest education of parents", "FreeSurfer QC motion", "Parent (self-report) depressive problems", "Family ID"))

```

Note: While we are primarily using Release 4.0, we will use some covariates (race.6.level, BMI) from DEAP (still using Release 3.0) as these variables have been derived by DEAP and are not available on the NDA site. There have been no updates/no changes on these variables between release 3.0 and 4.0. Additionally, for our covariates, we are using baseline and year 1 data (all available in 3.0), so are confident that this won't impact our analyses. 


```{r, load data}
#import site info from longitudinal tracking file
site <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/Culture_and_Environment/abcd_lt01.rds")

#import family ID to track family and population stratification weight
family <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/Culture_and_Environment/acspsw03.rds")

#note: DEAP is still waiting to be updated to release4.0.
covs <-rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/DEAP_data_downloads/covs_puberty_RR.rds") 

#parent mental health data
parent_mh <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/Mental_Health/abcd_asrs01.rds")

#household income info
demo <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/Mental_Health/pdem02.rds")


#import dti for additional DTI motion covariates
dti <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/MRI_T_roi/abcd_dti_p101.rds")

#import MRI info for scanner manufacturer ID
mri_info <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/MRI_T_roi/abcd_mri01.rds")

#import cleaned cortical measures to get WBV

cortical <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/cortical_year2_R4.0.rds")

```

See the following data dictionary pages for further info on variables below:

Longitudinal tracking info (site): https://nda.nih.gov/data_structure.html?short_name=abcd_lt01

Family info: https://nda.nih.gov/data_structure.html?short_name=acspsw03

MRI info: https://nda.nih.gov/data_structure.html?short_name=abcd_mri01

```{r, reduce to variables of interest}
#rename variables so consistent across scripts and more intuitive
covs <- covs %>% 
  rename(eventname = event_name, sex = sex_at_birth, bmi = anthro_bmi_calc)
colnames(covs)

#extract bmi and race from covs df
covs <- covs %>% 
  select(src_subject_id, eventname, sex, bmi, race.6level)

#parent mood
parent_mh <- parent_mh %>% 
  select(src_subject_id, eventname, sex, asr_scr_depress_r)

#extract site id (this give site for each event name) and subject id 
site <- site %>% 
  select(src_subject_id, eventname, sex, interview_age, site_id_l)

#extract family id, acs stratification weight 
family <- family %>% 
  select(src_subject_id, eventname, sex, rel_family_id, acs_raked_propensity_score)

#extract household income (note: income also available in covs df but var extracted here is more detailed, i.e., has more levels)
demo <- demo %>% 
  select(src_subject_id, eventname, demo_comb_income_v2)

#extract dti motion covariates
dti <- dti %>%
  select(src_subject_id, eventname, sex, dmri_dti_meanmotion, dmri_dti_meanrot, dmri_dti_meantrans)

#create event name column for cortical df to keep consistent with others. This column isn't present because it was removed at earlier QC point when  dataframes were separated into baseline and year 2 before pilot analysis.

cortical$eventname <- "2_year_follow_up_y_arm_1"

#extract WBV, ICV 
WBV <- cortical %>%
  select(src_subject_id, eventname, WBV, ICV)


#rename MRI info vars
mri_info <- mri_info %>%
  rename(scanner_id = mri_info_deviceserialnumber)
#extract MRI info
mri_info <- mri_info %>%
  select(src_subject_id, eventname, sex, mri_info_visitid, scanner_id, mri_info_studydate)


```
Let's explore family ID structure 

```{r, family ID}

family_id_tbl <- family %>%
  group_by(eventname) %>% 
  summarise(no_rows = length(family))
print(family_id_tbl)

```

##Extra info on income variable. 

Income data is only available at baseline 

Income level descriptions:

1= Less than $5,000; 2=$5,000 through $11,999; 3=$12,000 through $15,999; 4=$16,000 through $24,999; 5=$25,000 through $34,999; 6=$35,000 through $49,999; 7=$50,000 through $74,999; 8= $75,000 through $99,999; 9=$100,000 through $199,999; 10=$200,000 and greater. 999 = Don't know ; 777 = Refuse to answer

# CONVERT TO WIDE AND TIDY

To derive variables for different timepoints, it is easier to convert the dataframe to wide format. 

```{r, convert to wide, filter}

#convert first covariate dataframe to wide
covs_wide <- reshape(covs, idvar = "src_subject_id", timevar = "eventname", direction = "wide")
#rename variables of interest
covs_wide <- covs_wide %>% 
  rename( bmi_y0 = bmi.baseline_year_1_arm_1, 
          bmi_y1 = bmi.1_year_follow_up_y_arm_1,
          bmi_y2 = bmi.2_year_follow_up_y_arm_1,
          sex = sex.baseline_year_1_arm_1, #using baseline sex variable
          race_y0 = race.6level.baseline_year_1_arm_1,
          race_6m = race.6level.6_month_follow_up_arm_1,
          race_y1 = race.6level.1_year_follow_up_y_arm_1,
          race_18m = race.6level.18_month_follow_up_arm_1,
          race_y2 = race.6level.2_year_follow_up_y_arm_1
          )
colnames(covs_wide)

#keep vars needed for imputation
covs_wide <- covs_wide %>% 
  select(src_subject_id, bmi_y0, bmi_y1, bmi_y2, sex, race_y0, race_6m, race_y1, race_18m, race_y2)

#convert family df to wide
family_wide <- reshape(family, idvar = "src_subject_id", timevar = "eventname", direction = "wide")
family_wide <- family_wide %>% 
  rename(sex = sex.baseline_year_1_arm_1, #baseline sex variable 
         rel_family_id = rel_family_id.baseline_year_1_arm_1,#rel_family_id collected at baseline
         acs_raked_propensity_score_y0 = acs_raked_propensity_score.baseline_year_1_arm_1,
         acs_raked_propensity_score_y1 = acs_raked_propensity_score.1_year_follow_up_y_arm_1
         ) 
colnames(family_wide)

#select vars 
family_wide <- family_wide %>% 
  select(src_subject_id,sex, rel_family_id, acs_raked_propensity_score_y0, acs_raked_propensity_score_y1)

#convert site df to wide 
site_wide <- reshape(site, idvar = "src_subject_id", timevar = "eventname", direction = "wide")
site_wide <- site_wide %>% 
  rename(sex = sex.baseline_year_1_arm_1,
         site_id = site_id_l.baseline_year_1_arm_1,
         site_id_6m = site_id_l.6_month_follow_up_arm_1,
        site_id_y1 = site_id_l.1_year_follow_up_y_arm_1,
        site_id_18m = site_id_l.18_month_follow_up_arm_1,
        site_id_y2 = site_id_l.2_year_follow_up_y_arm_1,
        site_id_y3 = site_id_l.3_year_follow_up_y_arm_1,
        interview_age = interview_age.baseline_year_1_arm_1 #baseline age
           )
colnames(site_wide)

#select vars
site_wide <- site_wide %>% 
  select(src_subject_id, sex, site_id, site_id_6m, site_id_y1, site_id_18m, site_id_y2, site_id_y3, interview_age)

#convert parent mental health df to wide 
#no parent depression data available for year 1
parent_mh_wide <- reshape(parent_mh, idvar = "src_subject_id", timevar = "eventname", direction = "wide")
parent_mh_wide <- parent_mh_wide %>% 
  rename(sex = sex.baseline_year_1_arm_1,
         parent_dep_y0 = asr_scr_depress_r.baseline_year_1_arm_1,
         parent_dep_y2 = asr_scr_depress_r.2_year_follow_up_y_arm_1,
         parent_dep_y3 = asr_scr_depress_r.3_year_follow_up_y_arm_1)
colnames(parent_mh_wide)

#select vars of interest
parent_mh_wide <- parent_mh_wide %>% 
  select(src_subject_id, sex, parent_dep_y0, parent_dep_y2, parent_dep_y3)

#convert to demo to wide
demo_wide <- reshape(demo, idvar = "src_subject_id", timevar = "eventname", direction = "wide")
#rename
demo_wide <- demo_wide %>% 
  rename(household_income_y0 = demo_comb_income_v2.baseline_year_1_arm_1)

#convert wbv to wide
wbv_wide <- reshape(WBV, idvar = "src_subject_id", timevar = "eventname", direction = "wide")
wbv_wide <- wbv_wide %>% 
  rename(WBV = WBV.2_year_follow_up_y_arm_1,
         ICV = ICV.2_year_follow_up_y_arm_1)
  
#convert dti vars to wide
dti_wide <- reshape(dti, idvar = "src_subject_id", timevar = "eventname", direction = "wide")
dti_wide <- dti_wide %>% 
  rename(sex = sex.baseline_year_1_arm_1,
         dti_mean_motion_y0 = dmri_dti_meanmotion.baseline_year_1_arm_1,
         dti_mean_motion_y2 = dmri_dti_meanmotion.2_year_follow_up_y_arm_1)
colnames(dti_wide)

#select vars of interest 
dti_wide <- dti_wide %>% 
  select(src_subject_id, sex, dti_mean_motion_y0, dti_mean_motion_y2)

#convert mri info vars to wide 

mri_wide <- reshape(mri_info, idvar = "src_subject_id", timevar = "eventname", direction = "wide")
mri_wide <- mri_wide %>% 
  rename(sex = sex.baseline_year_1_arm_1,
         scanner_id_y2 = scanner_id.2_year_follow_up_y_arm_1)
mri_wide <- mri_wide %>% 
  select(src_subject_id, sex, scanner_id_y2)

```

##MERGE DATAFRAMES

```{r, merge dataframes}

#merge dfs by id and sex. N should be equal to 11,876

#merge site with parent_mh 
covs_main <- site_wide  %>% 
  left_join(y=parent_mh_wide, by= c("src_subject_id", "sex"))

#merge with family
covs_main <- covs_main %>% 
  left_join(y=family_wide, by= c("src_subject_id", "sex"))

#merge with remaining covs
covs_main <- covs_main  %>% 
  left_join(y=covs_wide, by= c("src_subject_id", "sex")) 

#merge with dti motion covs
covs_main <- covs_main  %>% 
  left_join(y=dti_wide, by= c("src_subject_id", "sex")) 

#merge with mri info covs
covs_main <- covs_main  %>% 
  left_join(y=mri_wide, by= c("src_subject_id", "sex"))

#merge with wbv wide and demo wide (only by subject id)
covs_main <- covs_main %>% 
  left_join(y=wbv_wide, by= "src_subject_id")

#merge demo wide
covs_main <- covs_main %>% 
  left_join(y=demo_wide, by= "src_subject_id")

#inspect main df after merge
colnames(covs_main)
head(covs_main)

```
##INSPECT DATA

```{r, tidy up variables}

#change age to years instead of months 
covs_main <- covs_main %>%
  mutate(
    age_years = interview_age/12 
  )

# #check number of scanner ids: There are 30 different scanners used across 21 sites 
# scanner_id_tbl <- covs_all %>%
#   group_by(scanner_id) %>% 
#   summarise(no_rows = length(scanner_id))
# print(scanner_id_tbl)
# 
# #inspect what household income info is available per wave
# income <- covs_all %>% 
#   select(household_income, eventname)
# 
# income_tbl <- covs_all %>%
#   group_by(eventname) %>% 
#    summarise(no_rows = length(household_income))
# print(income_tbl)
# 
# parent_dep <-  covs_all %>% 
#   select(asr_scr_depress_r, eventname)
# 
# parent_dep_tbl <- covs_all %>%
#   group_by(eventname) %>% 
#    summarise(no_rows = length(parent_dep))
# print(parent_dep_tbl)

```


##MAIN ANALYSIS (full model)

```{r, check covs dataframe}

colnames(covs_main)
summary(covs_main)

```
Hypothesis 1 (no imaging data)

Basic Model : Age, race, family, site 

Full model: age, race, family, site + BMI + income + parental mood  + population weight. 

All covariates are from baseline expect for parental mood, which is taken from year 2. Although our outcome measure is youth depression at Year 3, there was no parental mood available from year 3 in release 4.0. 

See PREP_imputation script for exploration and imputation of missing data

##Export cleaned data 

```{r, export data}
setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data")

saveRDS(covs_main, "covs_main_R4.0.rds")
```
####------------ END OF MAIN SCRIPT -----------------####



# PILOT ANALYSIS: Core Covariates (Basic Model)

For reference only

- Fixed: age, race/ethnicity
- Random: family/scanner 


```{r, BASELINE DF and QC}
#### Baseline #####
covs_pilot <- covs_all %>% 
  filter(eventname=="baseline_year_1_arm_1") #N=11,878

### Core covariates for basic model ###

#Age --- NAs = 2
summary(covs_pilot$age_years)

#Race/ethnicity --- NAs = 171

#Note: We are using a 6-level race measure instead of race_etnicity variable in an effort to make the measure more representative. "race.6level" includes White, Black, Asian, AIAN/NHPI (Asian Indian/Alaskan American/Native Hawaiian/Pacific Islander), Other, Mixed. 
summary(covs_pilot$race.6level)

#family: NAs = 2
summary(covs_pilot$rel_family_id)

#Scanner ID--- NAs = 77
#use code below to get info on each level of char variable (incl. NAs)
scanner_tbl <- covs_pilot %>%
  group_by(scanner_id) %>% 
  summarise(no_rows = length(scanner_id))
print(scanner_tbl)



```


