---
title: "PREP_imaging_R3.0"
author: "Niamh MacSweeney" Credit to Gladi Thng and Xueyi Shen for template scripts.
date: "8/3/2021"
output: html_document
---
#### Introduction
The purpose of this script is to quality control and process the ABCD imaging data (cortical, diffusion/white matter, subcortical)

#Helpful Documents:

Follow the QC criteria recommended by ABCD in NDA 2.0.1 and 3.0.1 Imaging Instrument Release Notes: https://nda.nih.gov/study.html?id=634
Refer to the sMRI and dMRI release notes, also found at https://nda.nih.gov/study.html?id=634
Data dictionaries (located in same folder as corresponding data) are also useful for understanding variables. 

The current preprocessing followed the recommendations by Hagler et al., 2019 Neuroimage paper: https://www.sciencedirect.com/science/article/pii/S1053811919306822

Helpful link for grep syntax: https://biocorecrg.github.io/CRG_RIntroduction/regular-expressions-to-find-more-flexible-patterns.html

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#SETUP
Load libraries needed and set working directory.

```{r, setup}

library(tidyverse)
library(dplyr)
library(broom)
library(ggplot2)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")



```


#STEP 1: LOAD DATA FOR QUALITY CONTROL AND SELECT VALID IDS

Need satisfactory T1 raw images, Freesurfer (FS) and DTI QC, reduce to baseline data only.
```{r, load data and tidy}

qc.t1 <-rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/MRI_QC/mriqcrp102.rds") #t1 image QC file 
qc.fs <-rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/MRI_QC/freesqc01.rds") #freesurfer QC file

qc.criteria <- full_join(qc.t1,qc.fs) #merge two dataframes

#reduce to baseline data only, N=11,815
qc.criteria <- qc.criteria %>% 
  filter(eventname == "baseline_year_1_arm_1")

```


Select IDs that passed QC for T1 FS and DTI

field name: iqc_t1_ok_ser = T1: Number of series that are complete and passed QC. Valid IDs: iqc_t1_ok_ser > 0
field name: fsqc_qc = FS QC score. Valid IDs: fsqc_qc = 1
field name: iqc_dmri_ok_ser = DTI: Number of series that are complete and passed QC. Valid IDs: iqc_dmri_ok_ser > 0

```{r, valid IDs}

t1.ids.pass <- qc.criteria %>% 
  select(src_subject_id,iqc_t1_ok_ser,fsqc_qc) %>%  
  filter(iqc_t1_ok_ser >0) %>% # t1 raw scans that passed QC
  filter(fsqc_qc==1) # passed freesurfer QC
t1.ids.pass <- t1.ids.pass$src_subject_id # make new column for valid t1 IDs --- N = 11,236

dti.ids.pass <- qc.criteria %>% 
  select(src_subject_id,iqc_dmri_ok_ser,iqc_t1_ok_ser,fsqc_qc) %>% 
  filter(iqc_dmri_ok_ser>0) %>% #DTI scans that passed QC
  filter(iqc_t1_ok_ser>0) %>% 
  filter(fsqc_qc==1) 
dti.ids.pass <- dti.ids.pass$src_subject_id #make new column for valid DTI IDs ---- N = 10,701

```

#STEP 2: PROCESS IMAGING DATA

#### Cortical Measures #####

#Using cortical volume, surface area, thickness and sulcal depth. 

```{r, load cortical measures data}

abcd.smrip101 <- rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/abcd_smrip101.rds")

abcd.cortical<- abcd.smrip101 %>%  #make df name more intuitive
  filter(eventname == "baseline_year_1_arm_1") %>% #baseline only --- N= 11736
  .[,grep("subject|thick_cdk|sulc_cdk|area_cdk|vol_cdk", colnames(.))] #extract variables of interest using grep
```

Rename columns to make more intuitive
```{r, rename columns}
colnames(abcd.cortical)=gsub('smri_','',colnames(abcd.cortical))
colnames(abcd.cortical)=gsub('_cdk_','.APARC.',colnames(abcd.cortical)) #APARC = cortical parcellation in FS syntax
colnames(abcd.cortical)=gsub('thick.','thk.',colnames(abcd.cortical))
colnames(abcd.cortical)=gsub('area.','sa.',colnames(abcd.cortical))
```

#Create ICV and remove wholeb and csf
```{r}

abcd.cortical$ICV_ASEG = abcd.cortical$vol.ASEG.intracranialv
abcd.cortical=abcd.cortical[,!grepl('vol.ASEG.intracranialv',colnames(abcd.cortical))]

```

Rearrange col names so that hemisphere (lh or rh or bl (bilateral - have one measure)) is indicated as the start of col name
```{r, further renaming}

colnames(abcd.cortical)[grep('lh$',colnames(abcd.cortical))]=paste0('lh.',gsub('lh$','',colnames(abcd.cortical)))[grep('lh$',colnames(abcd.cortical))]
colnames(abcd.cortical)[grep('rh$',colnames(abcd.cortical))]=paste0('rh.',gsub('rh$','',colnames(abcd.cortical)))[grep('rh$',colnames(abcd.cortical))]
colnames(abcd.cortical)[grep('^vol.|^sa.|^sulc.|^thk.',colnames(abcd.cortical))]=paste0('bl.',colnames(abcd.cortical))[grep('^vol.|^sa.|^sulc.|^thk.',colnames(abcd.cortical))]
```

We are only interested in overall total values for this analysis so remove totals for lh and rh

```{r}
cols.omit <- abcd.cortical %>% .[grep(c("^rh.+\\.total$|^lh.+\\.total$|^rh.+\\.mean$|^lh.+\\.mean$"),colnames(.))] 
abcd.cortical <- abcd.cortical[,!colnames(abcd.cortical) %in% colnames(cols.omit)]
```

Select IDs that passed QC check ----  N=11232
```{r, select QC'd IDs}

abcd.cortical.final <- abcd.cortical %>% 
  .[.$src_subject_id %in% t1.ids.pass,] 

colnames(abcd.cortical.final)

```

Check distribution for global (total) vol and sa. Looks normally distributed so all good!

```{r, check distribution vol and sa}

abcd.cortical.final %>%
  select(contains(".total")) %>% 
  gather() %>% 
  ggplot(aes(as.numeric(value)))+
  facet_wrap(~ key, scales = "free") +
  geom_histogram(fill = "deepskyblue4") #Looks fine. 

```

Check distribution for global (mean) sulc and thk. Looks okay too!

```{r, check distribution sulc and thk}

abcd.cortical.final %>%
  select(contains(".mean")) %>% 
  gather() %>% 
  ggplot(aes(as.numeric(value)))+
  facet_wrap(~ key, scales = "free") +
  geom_histogram(fill = "deepskyblue4", 
                 bins = 30) 

```

#### DTI/White matter Measures ####

Note to self: check how tracks are being defined.  

Using fractional anisotropy and mean diffusivity. 

```{r, load DTI measures}

abcd.dti <- rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/abcd_dti_p101.rds")

abcd.dti <- abcd.dti %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% #reduce to baseline --- N = 11,736
  .[,grep("subject|_dtifa_|_dtimd_", colnames(.))] %>% #select FA and MD cols
  .[,grep('subject|fiberat',colnames(.))] #include DTI atlas tract adjustment

```

Tidy up column names
```{r, rename columns}

colnames(abcd.dti)=gsub('dmri_','',colnames(abcd.dti))
colnames(abcd.dti)=gsub('fiberat_','',colnames(abcd.dti))

colnames(abcd.dti)[grep('\\.all&rh$',colnames(abcd.dti))]=
  paste0('rhtotal.',gsub('\\.all|rh$','',colnames(abcd.dti)))[grep('\\.all|rh$',colnames(abcd.dti))]

#rearrange col names as done for cortical measures to make reading easier. 
colnames(abcd.dti)[grep('lh$',colnames(abcd.dti))]=paste0('lh.',gsub('lh$','',colnames(abcd.dti)))[grep('lh$',colnames(abcd.dti))]
colnames(abcd.dti)[grep('rh$',colnames(abcd.dti))]=paste0('rh.',gsub('rh$','',colnames(abcd.dti)))[grep('rh$',colnames(abcd.dti))]
colnames(abcd.dti)[grep('\\_all',colnames(abcd.dti))]=paste0('bl.',colnames(abcd.dti))[grep('\\_all',colnames(abcd.dti))]
colnames(abcd.dti)[grep('^dti',colnames(abcd.dti))]=paste0('bl.',colnames(abcd.dti))[grep('^dti',colnames(abcd.dti))]


```

Remove average FA and MD without corpus callosum
```{r, select variables of interest}

#remove average FA and MD without corpus callosum
cols.omit <- abcd.dti %>% .[grep("_allfcc$|_allfib$",colnames(.))] 
abcd.dti <- abcd.dti[,!colnames(abcd.dti) %in% colnames(cols.omit)] #77 variables left. 

```

Select IDs that pass QC. Total N = 10,699
```{r, select valid IDs}

abcd.dti.final<- abcd.dti %>% 
  .[.$src_subject_id %in% dti.ids.pass,] #N= 10,699

colnames(abcd.dti.final) #get column names
```

Check distribution of FA and MD

The MD values do appear to be skewed. However, instead of removing outliers, we will conduct analysis with and without the outliers removed as sensitivity analysis. With a sample size this large, it is difficult to know whether there are false or spurious values.

Removing outliers of ±5 SD does not change the distribution of FA but does significantly change the distribution of MD. 

```{r, check distribution}

dti.dist.incl.outl <- abcd.dti.final %>%
  select(ends_with("allfibers"))%>% 
  gather() %>% 
  ggplot(aes(as.numeric(value)))+
  facet_wrap(~ key, scales = "free") +
  geom_histogram(fill = "deepskyblue4",
                 bins = 30) 
dti.dist.incl.outl


```

Outlier treatment

Note here, using bl.dti.fa(or md)_allfibers variables when removing outliers. 

```{r, define outliers}

#Define upper and lower limits (±5 SD)
outl_fa_upper <- (mean(abcd.dti.final$bl.dtifa_allfibers, na.rm=T) +5*(sd(abcd.dti.final$bl.dtifa_allfibers, na.rm=T))) #=0.611
outl_fa_lower <- (mean(abcd.dti.final$bl.dtifa_allfibers, na.rm=T) -5*(sd(abcd.dti.final$bl.dtifa_allfibers, na.rm=T))) #=0.391
outl_md_upper <- (mean(abcd.dti.final$bl.dtimd_allfibers, na.rm=T) +5*(sd(abcd.dti.final$bl.dtimd_allfibers, na.rm=T))) #=0.916
outl_md_lower <- (mean(abcd.dti.final$bl.dtimd_allfibers, na.rm=T) -5*(sd(abcd.dti.final$bl.dtimd_allfibers, na.rm=T))) #=0.681

```
```{r, plot FA outliers}
hist(abcd.dti.final$bl.dtifa_allfibers, main="", xlab="FA ALL FIBERS")
abline(v=outl_fa_upper,col="red")
abline(v=outl_fa_lower,col="red")

```
```{r, plot MD outliers}
hist(abcd.dti.final$bl.dtimd_allfibers, main="", xlab="MD ALL FIBERS")
abline(v=outl_md_upper,col="red")
abline(v=outl_md_lower,col="red")

```
```{r, identify number of FA and MD outliers} 

fa_outl <- abcd.dti.final %>% 
  filter(bl.dtifa_allfibers > outl_fa_upper |
           bl.dtifa_allfibers < outl_fa_lower) #14 outliers present

md_outl <- abcd.dti.final %>% 
  filter(bl.dtimd_allfibers > outl_md_upper |  #6 outliers present 
           bl.dtimd_allfibers < outl_md_lower)

```

Create new FA and MD variables with outliers removed for later sensitivity analysis. N =10,699 before outlier removal. 

I can't get this to work at the moment. I want to do a sensitivity analysis with and without outliers for the DTI measures but I'm finding it difficult to remove the outlier IDs and then save as a new variable because the "all fibers" variable is a kind of grouping measure for many FA and MD measures. 
```{r, outlier removal}

#new FA variable excl. outliers
abcd.dti.final <- abcd.dti.final %>% 
  mutate(bl.dtifa_allfibers_excl_outl = bl.dtifa_allfibers) 

#new MD variable excl. outliers
abcd.dti.final <- abcd.dti.final %>% 
  mutate(bl.dtimd_allfibers_excl_outl = bl.dtimd_allfibers) 


#Change outlier values to NA

#FA
abcd.dti.final$bl.dtifa_allfibers_excl_outl[abcd.dti.final$bl.dtifa_allfibers_excl_outl > outl_fa_upper |abcd.dti.final$bl.dtifa_allfibers_excl_outl < outl_fa_lower] <- NA
#MD
abcd.dti.final$bl.dtimd_allfibers_excl_outl[abcd.dti.final$bl.dtimd_allfibers_excl_outl > outl_md_upper |abcd.dti.final$bl.dtimd_allfibers_excl_outl < outl_md_lower] <- NA

```
```{r, check it worked} 

fa_outl_check <- abcd.dti.final %>% 
  filter(bl.dtifa_allfibers > outl_fa_upper |
           bl.dtifa_allfibers < outl_fa_lower) #14 outliers present

md_outl_check <- abcd.dti.final %>% 
  filter(bl.dtimd_allfibers > outl_md_upper |  #6 outliers present 
           bl.dtimd_allfibers < outl_md_lower)

#THE ABOVE DOES NOT WORK AT THE MOMENT FOR THE REASONS OUTLINED ABOVE. FOR THE MOMENT, I WIL TRY AND PROCEED WITH THE REST OF QC AND ASK PEOPLE FOR THEIR HELP. 

```


Below, I've created a dataframe with the outlier excluded. I think I will have to export this as a separate RDS file and then run the sensitivity analysis separately.... 

```{r, identify outliers and create additional variables for sensitivity analysis} 


abcd.dti.final.excl.outl <- abcd.dti.final <- filter(abcd.dti.final,!is.na(bl.dtifa_allfibers))

abcd.dti.final.excl.outl[abs(scale(abcd.dti.final.excl.outl$bl.dtifa_allfibers))>5,
                         grep('dtifa',colnames(abcd.dti.final.excl.outl))]=NA

abcd.dti.final.excl.outl[abs(scale(abcd.dti.final.excl.outl$bl.dtimd_allfibers))>5,
                         grep('dtimd',colnames(abcd.dti.final.excl.outl))]=NA


```

Note, #DTI data final N = 10,643 (excluding outliers)

Subcortical measures (before QC, N =11,736)

These are not included in the Puberty RR but I am processing them here for future reference
```{r, load in subcortical measures}

abcd.subcort <- rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/abcd_smrip201.rds")

abcd.subcort <- abcd.subcort %>% 
filter(eventname == "baseline_year_1_arm_1") %>% #reduce to baseline --- N = 11,736
.[,grep("subject|smri_vol_scs_", colnames(.))] %>%   #select smri vol data
.[,grep("subject|aa|amygdala|caudate|hpus|pallidum|putamen|tp|cranial|subcortical",colnames(.))] #select columns overlap with ENIGMA 

```
```{r, tidy column names}

colnames(abcd.subcort)=gsub('smri_','',colnames(abcd.subcort))
colnames(abcd.subcort)=gsub('_scs_','.ASEG.',colnames(abcd.subcort))

```

Create ICV and remove wholeb, csf
Note: abcd.subcort$ICV_derived = abcd.subcort$vol.ASEG.wholeb+abcd.subcort$vol.ASEG.csf
```{r, create ICV and remove wholeb, csf}
abcd.subcort$ICV_ASEG =abcd.subcort$vol.ASEG.intracranialv
abcd.subcort=abcd.subcort[,!grepl('vol.ASEG.intracranialv',colnames(abcd.subcort))]
```


```{r, further tidy of column names}

colnames(abcd.subcort)[grep('l$|lh$',colnames(abcd.subcort))]=paste0('lh.',gsub('l$|lh$','',colnames(abcd.subcort)))[grep('l$|lh$',colnames(abcd.subcort))]
colnames(abcd.subcort)[grep('r$|rh$',colnames(abcd.subcort))]=paste0('rh.',gsub('r$|rh$','',colnames(abcd.subcort)))[grep('r$|rh$',colnames(abcd.subcort))]
colnames(abcd.subcort)[grep('^vol.',colnames(abcd.subcort))]=paste0('bl.',colnames(abcd.subcort))[grep('^vol.',colnames(abcd.subcort))]

```

Select IDs that passed QC, N=11,232
```{r, select valid IDs}

abcd.subcort.final<- abcd.subcort %>% 
  .[.$src_subject_id %in% t1.ids.pass,]   #N= 11232
```

Check distribution of subcortical measures

```{r, check distribution}

abcd.subcort.final %>%
  select(!starts_with("src"))%>% 
  gather() %>% 
  ggplot(aes(as.numeric(value)))+
  facet_wrap(~ key, scales = "free") +
  geom_histogram(fill = "deepskyblue4",
                 bins =30) #all look okay

```

Let's look at the final variables for subcortical measures (17 in total) after QC
```{r, subcortical column names}

colnames(abcd.subcort.final)
```
 EXPORT CLEANED DATA

Change working directory and save clean file there. 
Note: clean files should be saved in data folder in ABCD_puberty_RR project folder.
```{r, export cleaned data}

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data")
saveRDS(abcd.cortical.final,"cortical_cleaned_R3.0.rds")
saveRDS(abcd.dti.final,"dti_cleaned_R3.0.rds")
saveRDS(abcd.subcort.final,"subcort_cleaned_R3.0.rds")

```

####------------ END OF MAIN SCRIPT -----------------####