---
title: "RR Sample Sizes"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#####Introduction

We need to generate information about what our sample size will be for our main analysis, 

In order to be included in the final analysis, the following criteria must be met:

1. Must be an unrelated individual 
2. Must not have depressive symptoms at year 1
3. Must have puberty, imaging, and depression data at year 1, year 2 and year 3 respectively. 

We will run FIML for people who have missing covariate data. 

#####Required Inputs

Cleaned data Rds files are located in /Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data

Depressive symptoms: CBCL_dep_Y1_R4.0.rds (Year 3 = main outcome measure) and CBCL_dep_Y1_R4.0.rds (need to control for depression at y1)

Brain structure: cortical_Y2_R4.0.rds + subcort_Y2_R4.O.rds + dti_Y2_R4.0.rds

Covariates — Covariates: covs_pilot_R4.0.rds

####SET UP

```{r, load libraries and set wd}

library(tidyr)
library(dplyr)
library(tidyverse)
library(psych)
library(ggplot2)
library(ggstatsplot)
library(GGally)
library(stringr)
library(outliers)
library(jtools) 
library(readr)
library(gridExtra)
library(cowplot)
library(ggpubr)
library(DiagrammeR)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")
getwd()

```

Load data 

```{r, load data}

#Depression data 
dep_y1 <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/CBCL_dep_y1_R4.0.rds")

dep_y3 <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/CBCL_dep_y3_R4.0.rds")

#Imaging data
cortical <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/cortical_year2_R4.0.rds")

subcort <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/subcort_year2_R4.0.rds")

dti <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/dti_year2_R4.0.rds")

#puberty data

pds <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/pds_timing_y1_R4.0.rds")

#Covariates
covs <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/covs_pilot_R4.0.rds")

```

Tidy data

```{r, tidy data}

#select variables of interest for dep_y1

dep_y1 <- dep_y1 %>% 
  select(src_subject_id, dep_status_y1_t)

pds <- pds %>% 
  select(src_subject_id, sex, pds_tot_all, pt_m, pt_f)

```

Select valid IDs

```{r, valid IDs}

#### CBCL #####

valid_ids <-  dep_y1 %>% 
  select(src_subject_id, dep_status_y1_t) %>%  
  filter(dep_status_y1_t == "not depressed") #only keep individuals that were not depressed at y1. 
valid_ids <- valid_ids$src_subject_id

dep_y3_final <- dep_y3 %>% 
  .[.$src_subject_id %in% valid_ids,] #N= 5835

```

Merge dataframes
```{r, merge}

df <- dep_y3_final %>% 
  left_join(y=cortical, by= "src_subject_id") %>% 
  left_join(y=subcort, by= "src_subject_id") %>% 
  left_join(y=dti, by= "src_subject_id") %>% 
  left_join(y = pds, by= "src_subject_id") %>% 
  left_join(y=covs, by="src_subject_id")

#rename duplicate vars 
df <- df %>% 
  rename(sex = sex.x)

#remove sex.y 
df <- df %>% 
  select((-sex.y))

```

Check complete cases
```{r, complete cases}

#Make male and female def 

males_df <- df %>% #N=2844
  filter(sex == "M")

females_df <- df %>% #N=2635
  filter(sex == "F")

#remove pt_timing variables before checking complete cases because there will be many NAs for the sex specific cols. 

df_complete <- df %>% 
  dplyr::select(- c(pt_f, pt_m))

sum(complete.cases(df_complete)) #N=3872 for all variables. 

dep_pt <- df %>% 
  select("src_subject_id", "cbcl_scr_syn_withdep_r", "pds_tot_all", "eventname", "bmi", "household_income", "race.6level","asr_scr_depress_r",  "site_id_l", "age_years")

sum(complete.cases(dep_pt)) #5078

sum(complete.cases(df$pds_tot_all)) #N=5479

sum(complete.cases(df$acs_raked_propensity_score)) #5835

```
Make flow chart - do this in stage 2

```{r, flow chart}

```



