---
title: "PREP_CBCL_R4.0"
Date: 16/03/22
Author: Niamh MacSweeney
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Introduction

The purpose of this script is to generate depression severity measure (continuous variable) from the Child Behaviour Checklist (CBCL) using ABCD Release 4.0 to prepare a final, clean, sample for use in models.

Script workflow:
1. Setup: load data, tidy, reduce to year 3 wave, and get sample size
2. Step 1: Inspect data. 

Resources:

- Please consult this guide for an overview of the DSM scales: https://aseba.org/wp-content/uploads/DSM-Oriented-Guide-for-the-ASEBA.pdf

A note on T-scores from the ASEBA guide:

"Normalized T scores are standard scores that are based on the percentiles occupied by raw scale scores in the distribution of scores obtained by individuals in the relevant nor- mative sample. Consequently, the T score obtained by an individual tells us approximately how high (in terms of a percentile) the individual’s scale score is, compared to the scores obtained by individuals in the relevant normative sample"

Because substantial percentages of individuals in the normative samples obtained very low scores on the DSM-oriented and syndrome scales (e.g., scores of 0 or 1), the T score assignments start at 50, which represents the 50th percentile of scores in a normative sample. In other words, all raw scale scores that were in the lowest 50 percent of the distribution are assigned a T score of 50.

However, for statistical analyses, raw scale scores are often preferable, because they reflect all the variation that actually occurs in scores obtained by the individuals whose data are being analyzed. T scores, by contrast, lump together some raw scale scores, such as the different low scores that are given a T score of 50"

SETUP Load libraries needed and set working directory

```{r, libraries}

library(tidyr)
library(dplyr)
library(tidyverse)
library(psych)
library(ggplot2)
library(ggstatsplot)
library(GGally)
library(stringr)
library(outliers)
library(jtools) 
library(readr)
library(gridExtra)
library(cowplot)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")
getwd()

```

LOAD DATA, REDUCE VARIABLES, CREATE DATAFRAME

Variables of interest: short names (as per NDA data dictionary)

- DSM-5 dpression scale (T-Score) — i.e., standarised score: cbcl_scr_dsm5_depress_t

-CBCL 




```{r, load data and tidy}

cbcl <-rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/Mental_Health/abcd_cbcls01.rds")

head(cbcl, n=10) #first few observations
str(cbcl)  #types of variables
names(cbcl) #List variables in dataframe

#reduce to baseline data only, N= 6251
cbcl <- cbcl %>% 
  filter(eventname == "3_year_follow_up_y_arm_1")

#extract all variables containing "dep" for depression variables

cbcl_dep <- cbcl %>% 
  select(src_subject_id,interview_date, interview_age, sex, eventname, contains("dep"))

colnames(cbcl_dep) #check correct columns have been extracted
 
summary(cbcl$cbcl_scr_dsm5_depress_r)

dsm_dep_plot <- cbcl_dep %>% 
  ggplot(aes(x = cbcl_scr_dsm5_depress_r)) +
  geom_histogram(bins = 15,
                colour = "skyblue4",
                 fill = "skyblue3") +
  labs(title = "CBCL Depression", x = "CBCL_withdrawn_Depression (raw score)") +
  scale_x_continuous(breaks=seq(0, 20, 3), limits = c(0,20))
  
print(dsm_dep_plot)
plot(cbcl_dep$cbcl_scr_syn_withdep_r)

```
