---
title: "Creating DSM-V MDD severity measure in ABCD"
author: Xueyi Shen
date:  "`r Sys.Date()`"
output: github_document
---

Note: This script was used by Niamh to create lifetime MDD and DS severity measures for use in models. 

* * *
### Contact and credit

X Shen (xueyi.shen@ed.ac.uk)


* * *
### R settings

Required packages: 
 
 - dplyr
 - readr

Extra packages for Rmarkdown:

 - rmarkdown
 - knitr

Required inputs:

 - directory where tabulated rds files are stored (e.g. 'release2.0.1/iii.data')
 - a [csv file](https://github.com/xshen796/ABCD_UoEPsych/raw/main/DataManagement/Phenotype/MDD_and_Depressive_symptoms/ls.questionnaires.csv) with all data fields needed
 
ABCD data: [release 3.0.1](https://nda.nih.gov/data_dictionary.html?source=ABCD%2BRelease%2B3.0&submission=ALL)
 
* * *
### Main script body

#### Basic settings (on Eddie)
```{r}
library(dplyr)
library(knitr)
tab.dat.dir='/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data'
```


* * *
#### Merge and create one data obj for analysis
```{r}
# merge and save the master file  ----------------------------------------------------------------------
dir.var.ls='/Volumes/GenScotDepression/users/niamh/ABCD_UoEPsych/DataManagement/Phenotype/MDD_and_Depressive_symptoms/ls.questionnaires.csv'
ls.var=read.csv(dir.var.ls,stringsAsFactors=FALSE,sep = '\t',quote = '')
ls.files.tocatch=paste0(ls.var$Category,'/',ls.var$File.name)
ls.files.tocatch=unique(ls.files.tocatch)
for (i in ls.files.tocatch){
      tmp.file=readRDS(paste0(tab.dat.dir,'/',i,'.rds'))
      ## follow-up track file have duplicated IDs. Only keep the baseline data. 
      if (length(unique(tmp.file$src_subject_id))<length(tmp.file$src_subject_id)){
            tmp.file = tmp.file[order(tmp.file$interview_date),]
            tmp.file = tmp.file[!duplicated(tmp.file$src_subject_id),]}
      
      if (i==ls.files.tocatch[1]){questn.dat=tmp.file}else{
                  cols.keep=!grepl('interview_date|interview_age|sex|gender|eventname',colnames(tmp.file))
                  tmp.file=tmp.file[,cols.keep]
                  questn.dat=merge(questn.dat,tmp.file,by='src_subject_id',all=T)                  
            }
}
# grep('\\.x$',colnames(IM.dat))  # check if there is any duplicated colnames
rm(i,tmp.file)
questn.dat=questn.dat[,ls.var$Field.name]
```

```{r}
targetdat=questn.dat
targetdat.clean=targetdat
```


* * *
#### Create summary variables for life-time MDD diagnosis

Define function
```{r}
sum_diagnosis <- function(x,diagnosis,dic) {
      tmp.dic = dic[grep(diagnosis,dic$Description),]
      tmp.field = tmp.dic$Field.name
      dat = x[,colnames(x) %in% tmp.field]
      tmp.output = rowSums(dat==1,na.rm=T)
      tmp.output[tmp.output>0]=1
      tmp.output[rowSums(!is.na(dat))==0]=NA
      return(tmp.output)
} 
```

Child report:
```{r}
mdd.diagnosis.dic = ls.var[ls.var$File.name %in% 'abcd_ksad501',]
diagnosis.dat.tomerge=
      data.frame(src_subject_id=targetdat.clean$src_subject_id,
                 KSADS.MDD.y=sum_diagnosis(targetdat.clean,'Major Depressive Disorder',mdd.diagnosis.dic))
targetdat.clean=merge(targetdat.clean,diagnosis.dat.tomerge)
```

Caregiver report:
```{r}
mdd.diagnosis.dic = ls.var[ls.var$File.name %in% 'abcd_ksad01',]
diagnosis.dat.tomerge=
      data.frame(src_subject_id=targetdat.clean$src_subject_id,
                 KSADS.MDD.p=sum_diagnosis(targetdat.clean,'Major Depressive Disorder',mdd.diagnosis.dic))
targetdat.clean=merge(targetdat.clean,diagnosis.dat.tomerge)
```

Clean up data fields used for life-time MDD diagnosis
```{r}
targetdat.clean=data.frame(targetdat.clean[,!grepl('^ksads_1_8',colnames(targetdat.clean))])
```


* * *
#### Create measures for depressive symptoms based on DSM-V criteria

[A helpful summary](https://www.psnpaloalto.com/wp/wp-content/uploads/2010/12/Depression-Diagnostic-Criteria-and-Severity-Rating.pdf)

A summary of [MDD symptomology items in ABCD](https://github.com/xshen796/ABCD_UoEPsych/blob/main/DataManagement/Phenotype/MDD_and_Depressive_symptoms/TableS1.DepreSymp_Items.xlsx?raw=true)

A summary of [DSM-V criteria for deriving depressive symptoms](https://github.com/xshen796/ABCD_UoEPsych/blob/main/DataManagement/Phenotype/MDD_and_Depressive_symptoms/TableS2_DepreSymptom_Derivation.xlsx?raw=true)

##### Total number of depressive symptoms reported

Prepare function
```{r}
ls.items.Symp = list(
                    'Depressed Mood',
                    'Anhedonia',
                    'Fatigue',
                    'Concentration Disturbance',
                    'Decreased Self-Esteem',
                    'Guilt',
                    'Hopeless',
                    'Impairment in functioning due to depression',
                    'Indecision',
                    'Irritability',  
                     c('Insomnia','Hypersomnia'),
                     c('Weight Loss','Weight Gain'),
                     c('Decreased Appetite','Increased Appetite'),
                     c('Psychomotor Agitation','Psychomotor Retardation'),
                     c('Suicid|suicid|Attempt|SelfInjur'))
sum_diagnosis <- function(x,tmp.dic,ls.items.groupSymp) {
      tmp.field = tmp.dic$Field.name
      dat = x[,colnames(x) %in% tmp.field]
      
      tmp.group.field = lapply(ls.items.groupSymp,FUN=function(x) tmp.dic$Field.name[grep(paste0(x,collapse='|'),tmp.dic$Description)])
      tmp.group.output = lapply(tmp.group.field,FUN=function(x) rowSums(data.frame(dat[,x])==1,na.rm=T)) %>%
                         bind_cols 
      tmp.group.output[tmp.group.output>1]=1
      tmp.output=tmp.group.output %>%
                         rowSums(.,na.rm=T)
      return(tmp.output)
} 
```



=> Child report
Prepare data dictionaries
```{r}
new.dic = ls.var[grep('^Symptom -|Suicid|suicid|Attempt|SelfInjur',ls.var$Description),] %>%  # Items under MDD category in ABCD (symptom) + suicidality & self-injurial behaviour (diagnosis)
          .[grep('_ksad',.$File.name),] %>%
          .[grep('t$',.$Field.name),]
new.dic$Description=gsub(',','',new.dic$Description)
new.dic.ever = new.dic[!grepl('^Symptom - No two month symptom',new.dic$Description),]
new.dic.current = new.dic.ever[grep('Present',new.dic.ever$Description),]
new.dic.past = new.dic.ever[grep('Past',new.dic.ever$Description),]
new.dic.current = new.dic.current[order(new.dic.current$Description),]
new.dic.past = new.dic.past[order(new.dic.past$Description),]
```


The new data dictionaries look like:
```{r}
glimpse(new.dic.current)
```
```{r}
glimpse(new.dic.past)
```

Create life-time scores for each symptomology item: tmp.dat.ever.y
```{r}
for (i in 1:nrow(new.dic.current)){
    tmp.field.current=new.dic.current$Field.name[i]
    tmp.field.past=new.dic.past$Field.name[i]
    tmp.dat.col=rowSums(targetdat.clean[,c(tmp.field.current,tmp.field.past)],na.rm=T)
    tmp.dat.col[rowSums(is.na(targetdat.clean[,c(tmp.field.current,tmp.field.past)]))]=NA
    if(i==1){tmp.dat.ever=tmp.dat.col}else{tmp.dat.ever=cbind(tmp.dat.ever,tmp.dat.col)}
}
colnames(tmp.dat.ever)=new.dic.current$Field.name
tmp.dat.ever[tmp.dat.ever>1]=1
tmp.dat.ever.y=tmp.dat.ever
dic.current.y=new.dic.current
```

Create data and remove participants with all NAs for either MDD mood symptoms or self-harm/suicidality data
```{r}
# create data
Depre_symptoms.DSM.tomerge=
      data.frame(src_subject_id=targetdat.clean$src_subject_id,
                 KSADS.Dep_symptoms_sum_ever.y=sum_diagnosis(tmp.dat.ever,new.dic.current,ls.items.Symp),
                 KSADS.Dep_symptoms_sum_current.y=sum_diagnosis(targetdat.clean,new.dic.current,ls.items.Symp))
# set NAs
tmp.loc.symp = rowSums(!is.na(tmp.dat.ever[,new.dic.current$Field.name[grep('^Symptom',new.dic.current$Description)]]))==0
tmp.loc.diagn = rowSums(!is.na(tmp.dat.ever[,new.dic.current$Field.name[grep('^Diagnosis',new.dic.current$Description)]]))==0
tmp.loc = tmp.loc.symp==T&tmp.loc.diagn==T 
Depre_symptoms.DSM.tomerge$KSADS.Dep_symptoms_sum_ever.y[tmp.loc]=NA
tmp.loc = rowSums(is.na(targetdat.clean[,new.dic.current$Field.name]))
tmp.loc.symp = rowSums(!is.na(targetdat.clean[,new.dic.current$Field.name[grep('^Symptom',new.dic.current$Description)]]))==0
tmp.loc.diagn = rowSums(!is.na(targetdat.clean[,new.dic.current$Field.name[grep('^Diagnosis',new.dic.current$Description)]]))==0
tmp.loc = tmp.loc.symp==T&tmp.loc.diagn==T 
Depre_symptoms.DSM.tomerge$KSADS.Dep_symptoms_sum_current.y[tmp.loc]=NA
```

Merge with other phenotypes
```{r}
targetdat.clean=merge(targetdat.clean,Depre_symptoms.DSM.tomerge,by='src_subject_id',all.x=T)
```




=> Caregiver report
Prepare data dictionaries
```{r}
new.dic = ls.var[grep('^Symptom -|Suicid|suicid|Attempt|SelfInjur',ls.var$Description),] %>%  # Items under MDD category in ABCD (symptom) + suicidality & self-injurial behaviour (diagnosis)
          .[grep('_ksad',.$File.name),] %>%
          .[grep('p$',.$Field.name),]
new.dic.ever = new.dic[!grepl('^Symptom - No two month symptom',new.dic$Description),]
new.dic.current = new.dic.ever[grep('Present',new.dic.ever$Description),]
new.dic.past = new.dic.ever[grep('Past',new.dic.ever$Description),]
new.dic.current = new.dic.current[order(new.dic.current$Description),]
new.dic.past = new.dic.past[order(new.dic.past$Description),]
```

Create life-time scores for each symptomology item
```{r}
for (i in 1:nrow(new.dic.current)){
    tmp.field.current=new.dic.current$Field.name[i]
    tmp.field.past=new.dic.past$Field.name[i]
    tmp.dat.col=rowSums(targetdat.clean[,c(tmp.field.current,tmp.field.past)],na.rm=T)
    tmp.dat.col[rowSums(is.na(targetdat.clean[,c(tmp.field.current,tmp.field.past)]))]=NA
    if(i==1){tmp.dat.ever=tmp.dat.col}else{tmp.dat.ever=cbind(tmp.dat.ever,tmp.dat.col)}
}
colnames(tmp.dat.ever)=new.dic.current$Field.name
tmp.dat.ever[tmp.dat.ever>1]=1
tmp.dat.ever.p=tmp.dat.ever
dic.current.p=new.dic.current
```

The new data dictionaries look like:
```{r}
glimpse(new.dic.current)
```
```{r}
glimpse(new.dic.past)
```

Create data and remove participants with all NAs for either MDD mood symptoms or self-harm/suicidality data
```{r}
# create data
Depre_symptoms.DSM.tomerge=
      data.frame(src_subject_id=targetdat.clean$src_subject_id,
                 KSADS.Dep_symptoms_sum_ever.p=sum_diagnosis(tmp.dat.ever,new.dic.current,ls.items.Symp),
                 KSADS.Dep_symptoms_sum_current.p=sum_diagnosis(targetdat.clean,new.dic.current,ls.items.Symp))
# set NAs
tmp.loc.symp = rowSums(!is.na(tmp.dat.ever[,new.dic.current$Field.name[grep('^Symptom',new.dic.current$Description)]]))==0
tmp.loc.diagn = rowSums(!is.na(tmp.dat.ever[,new.dic.current$Field.name[grep('^Diagnosis',new.dic.current$Description)]]))==0
tmp.loc = tmp.loc.symp==T&tmp.loc.diagn==T 
Depre_symptoms.DSM.tomerge$KSADS.Dep_symptoms_sum_ever.p[tmp.loc]=NA
tmp.loc = rowSums(is.na(targetdat.clean[,new.dic.current$Field.name]))
tmp.loc.symp = rowSums(!is.na(targetdat.clean[,new.dic.current$Field.name[grep('^Symptom',new.dic.current$Description)]]))==0
tmp.loc.diagn = rowSums(!is.na(targetdat.clean[,new.dic.current$Field.name[grep('^Diagnosis',new.dic.current$Description)]]))==0
tmp.loc = tmp.loc.symp==T&tmp.loc.diagn==T 
Depre_symptoms.DSM.tomerge$KSADS.Dep_symptoms_sum_current.p[tmp.loc]=NA
```

Merge with other phenotypes
```{r}
targetdat.clean=merge(targetdat.clean,Depre_symptoms.DSM.tomerge,by='src_subject_id',all.x=T)
```



##### Depressive symptoms based on [DSM-V criteria](https://github.com/xshen796/ABCD_UoEPsych/blob/main/DataManagement/Phenotype/MDD_and_Depressive_symptoms/TableS2_DepreSymptom_Derivation.xlsx?raw=true)

Select data fields for  core symptoms, secondary symptoms and suicidality items
```{r}
Acol.item = 'Depressed.Mood|Anhedonia|Fatigue'
Suicidality.item = 'Suicid|Attempt'
ls.cols.p.core = dic.current.p$Field.name[grep(Acol.item,dic.current.p$Description)]
ls.cols.y.core = dic.current.y$Field.name[grep(Acol.item,dic.current.y$Description)]
ls.cols.p.suicidality = dic.current.p[grep(Suicidality.item,dic.current.p$Description),] %>%
                         .[!grepl('Suicidalideation',.$Description),] %>%
                         .[,'Field.name']
ls.cols.y.suicidality = dic.current.y[grep(Suicidality.item,dic.current.y$Description),] %>%
                         .[!grepl('Suicidalideation',.$Description),] %>%
                         .[,'Field.name']
```

Calculate total scores for core, secondary and suicidality items
```{r}
targetdat.clean$DS.core.total.p=rowSums(tmp.dat.ever.p[,ls.cols.p.core],na.rm=T)
targetdat.clean$DS.secondary.total.p=targetdat.clean$KSADS.Dep_symptoms_sum_ever.p-targetdat.clean$DS.core.total.p
targetdat.clean$DS.suicidality.total.p=rowSums(tmp.dat.ever.p[,ls.cols.p.suicidality],na.rm=T)
targetdat.clean$DS.core.total.y=rowSums(tmp.dat.ever.y[,ls.cols.y.core],na.rm=T)
targetdat.clean$DS.secondary.total.y=targetdat.clean$KSADS.Dep_symptoms_sum_ever.y-targetdat.clean$DS.core.total.y
targetdat.clean$DS.suicidality.total.y=rowSums(tmp.dat.ever.y[,ls.cols.y.suicidality],na.rm=T)
```


Create measures for depressive symptoms
=> Caregiver report
```{r}
targetdat.clean = mutate(targetdat.clean,KSADS.Depressive_symptoms_ever.p=ifelse(((DS.core.total.p==3 & DS.secondary.total.p>3)|(DS.suicidality.total.p>=1))&KSADS.MDD.p==1, 3, 
                                           ifelse((DS.core.total.p>1 & DS.secondary.total.p>2)|(KSADS.Dep_symptoms_sum_ever.p>=7 & KSADS.Dep_symptoms_sum_ever.p<=8),2,
                                           ifelse((DS.core.total.p>1 & DS.secondary.total.p>=1)|(KSADS.Dep_symptoms_sum_ever.p>=5 & KSADS.Dep_symptoms_sum_ever.p<=6),1,0))))
```

=> Child report
```{r}
targetdat.clean = mutate(targetdat.clean,KSADS.Depressive_symptoms_ever.y=ifelse(((DS.core.total.y==3 & DS.secondary.total.y>3)|(DS.suicidality.total.y>=1))&KSADS.MDD.y==1, 3, 
                                           ifelse((DS.core.total.y>1 & DS.secondary.total.y>2)|(KSADS.Dep_symptoms_sum_ever.y>=7 & KSADS.Dep_symptoms_sum_ever.y<=8),2,
                                           ifelse((DS.core.total.y>1 & DS.secondary.total.y>=1)|((KSADS.Dep_symptoms_sum_ever.y>=5 & KSADS.Dep_symptoms_sum_ever.y<=6)),1,0))))
targetdat.clean$KSADS.Depressive_symptoms_ever.p[is.na(targetdat.clean$KSADS.Dep_symptoms_sum_ever.p)]=NA
targetdat.clean$KSADS.Depressive_symptoms_ever.y[is.na(targetdat.clean$KSADS.Dep_symptoms_sum_ever.y)]=NA
```

Merge with other phenotypes and clean up intermediate data columns
```{r}
targetdat.clean$KSADS.Depressive_symptoms_ever.p[is.na(targetdat.clean$KSADS.Dep_symptoms_sum_ever.p)]=NA
targetdat.clean$KSADS.Depressive_symptoms_ever.y[is.na(targetdat.clean$KSADS.Dep_symptoms_sum_ever.y)]=NA
targetdat.clean = targetdat.clean[,!grepl('^DS.core|^DS.secondary|^DS.suici',colnames(targetdat.clean))]
```


* * *
##### Outputs:
In targetdat.clean (all columns below are life-time measures):

 - KSADS.MDD.y: MDD diagnosis based on child report
 - KSADS.MDD.p: MDD diagnosis based on caregiver report
 - KSADS.Dep_symptoms_sum_ever.y: no. of depressive symptoms reported based on child report
 - KSADS.Dep_symptoms_sum_ever.p: no. of depressive symptoms reported based on caregiver report
 - KSADS.Depressive_symptoms_ever.y: DSM-V depressive symptoms reported by children
 - KSADS.Depressive_symptoms_ever.p: DSM-V depressive symptoms reported by caregivers
 
 -----------------END OF SHEN'S SCRIPT---------------------
 
 Inspect new column outputs. Check to see levels within DS severity for youth and parent.
 They are coded as follows: 0=none, 1=mild, 2=moderate 3=severe. 
 
 Export clean variables for use in models, as outlined in Outputs.  
```{r}

sum_dep_measures <- targetdat.clean %>% 
  select (src_subject_id, KSADS.MDD.y, KSADS.MDD.p, KSADS.Dep_symptoms_sum_ever.y, KSADS.Dep_symptoms_sum_ever.p, KSADS.Depressive_symptoms_ever.y, KSADS.Depressive_symptoms_ever.p)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data")

saveRDS(sum_dep_measures,"dep_measures_cleaned_R3.0.rds")
```
 
Get breakdown of MDD and DS for caregiver and child report. 
Note missing data has changed to 0% for release 3.0, which is different to release 2.0 (NAs = DS (~30%) and MDD (~14%))

Need to double check that this is right!
```{r}

targetdat.clean %>% 
  dplyr::count(KSADS.MDD.y)

targetdat.clean %>% 
  dplyr::count(KSADS.MDD.p)

targetdat.clean %>% 
  dplyr::count(KSADS.Depressive_symptoms_ever.y)

targetdat.clean %>% 
  dplyr::count(KSADS.Depressive_symptoms_ever.p)



```
 




 