---
title: "PREP_covs_R3.0"
author: "Niamh MacSweeney"
date: "5/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The purpose of this script is to tidy and quality control the covariates for our analysis.
Covariates: Age, BMI, race/ethnicity, family income, site (for not imaging models), scanner_id (for imaging models) hemisphere (for bilateral brain measures)

This data frame was downloaded via the DEAP portal which works off the most recent ABCD data release (3.0). 

## Setup

Load libraries, set working directory.

```{r, setup}

library(psych)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggstatsplot)
library(readr)
library(gridExtra)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")
getwd()

```

##LOAD DATA

Note: Renaming variables to make the names more intuitive. 

Need to insert table here of all the covariates used.... 

```{r, covariate table}

vars <- factor(c("src_subject_id", "eventname", "sex", "age_years", "bmi", "race.6level", "demo_comb_income_p",  "abcd_site", "high.educ", "fsqc_qu_motion", "asr_scr_depress_r", "rel_family_id"))

var_lables <- factor(c("Participant ID", "Data collection wave", "Sex at birth", "Age in years", "Body Mass Index", "Race (6-level categorisation", "Household income", "Study site", "Highest education of parents", "FreeSurfer QC motion", "Parent (self-report) depressive problems", "Family ID"))

```

Note: While we are primarily using Release 4.0, we will use some covariates (race.6.level, BMI) from DEAP (still using Release 3.0) as these variables have been derived by DEAP and are not available on the NDA site. There have been no updates/no changes on these variables between release 3.0 and 4.0. Additionally, for our covariates, we are using baseline and year 1 data (all available in 3.0), so are confident that this won't impact our analyses. 


```{r, load data}
#import site info from longitudinal tracking file
site <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/Culture_and_Environment/abcd_lt01.rds")

#import family ID to track family
family <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/Culture_and_Environment/acspsw03.rds")

#note: DEAP is still waiting to be updated to release4.0.
covs <-rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/DEAP_data_downloads/covs_puberty_RR.rds") 

#import dti for additional DTI motion covariates 
dti <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/MRI_T_roi/abcd_dti_p101.rds")

#import MRI info for scanner manufacturer ID
mri_info <- rio::import("/Volumes/GenScotDepression/data/abcd/release4.0/iii.data/MRI_T_roi/abcd_mri01.rds")
```

See the following data dictionary pages for further info on variables below:

Longitudinal tracking info (site): https://nda.nih.gov/data_structure.html?short_name=abcd_lt01

Family info: https://nda.nih.gov/data_structure.html?short_name=acspsw03

MRI info: https://nda.nih.gov/data_structure.html?short_name=abcd_mri01

```{r, reduce to variables of interest}
#rename variables so consistent across scripts and more intuitive
covs <- covs %>% 
  rename(eventname = event_name, sex = sex_at_birth, bmi = anthro_bmi_calc)
colnames(covs)

#extract bmi and race, parent depression rating from covs df
covs <- covs %>% 
  select(src_subject_id, eventname, sex, bmi, race.6level, asr_scr_depress_r)

#extract site id (this give site for each event name) and subject id 
site <- site %>% 
  select(src_subject_id, eventname, sex, interview_age, site_id_l)

#extract family id, group id, acs stratification weight 
family <- family %>% 
  select(src_subject_id, eventname, sex, rel_family_id, rel_group_id, acs_raked_propensity_score)

#extract dti motion covariates 
dti <- dti %>% 
  select(src_subject_id, eventname, sex, dmri_dti_meanmotion, dmri_dti_meanrot, dmri_dti_meantrans)

#extract MRI info
mri_info <- mri_info %>% 
  select(src_subject_id, eventname, sex, mri_info_visitid, mri_info_manufacturersmn, mri_info_deviceserialnumber, mri_info_magneticfieldstrength, mri_info_studydate)
```

##MERGE DATAFRAMES

```{r, merge dataframes}

#merge dfs by id, eventname and sex
covs_all <- covs %>% 
  left_join(y=site, by= c("src_subject_id","eventname","sex")) #merge covs and site

covs_all <- covs_all %>% 
  left_join(y=dti, by= c("src_subject_id","eventname","sex")) #merge with dti

covs_all <- covs_all %>% 
  left_join(y=family, by= c("src_subject_id","eventname","sex")) #merge with family
                     
covs_all <- covs_all %>% 
  left_join(y=mri_info, by= c("src_subject_id","eventname","sex")) #merge with mri info

```
##INSPECT DATA

```{r, tidy up variables}

#change age to years instead of months 
temp <- covs_all %>%
  mutate(
    age_years = interview_age/12 
  )

#check number of scanner ids: There are 30 different scanners used across 21 sites 
scanner_id_tbl <- covs_all %>%
  group_by(mri_info_deviceserialnumber) %>% 
  summarise(no_rows = length(mri_info_deviceserialnumber))
print(scanner_id_tbl)

```

###MAKE SEPARATE DATAFRAMES PER TIMEPOINT:

1. Baseline (Pilot analysis)

2. Year 1 (Main analysis - non imaging covs)

3. Year 2 (Main analysis - imaging covs)

```{r, make separate DFs per timepoint}
#### Baseline #####
covs_y0 <- covs_all %>% 
  filter(eventname=="baseline_year_1_arm_1") #N=11,878
  
#### Year 1 ######
covs_y1 <- covs_all %>% 
  filter(eventname=="1_year_follow_up_y_arm_1") #N=11,235


colnames(covs_y0) #check new variable exists

#Let's reduce variables to those we need 
#Note: removing age in months, extra race, education (high.educ = high.educ.bl) and income vars.

covs <- covs %>% 
  select(-age, -female, -race.4level,  -high.educ.bl, - household.income)

#Let's reorder variables so that order is consistent with table above

# Age, BMI, race/ethnicity, family income, site, motion,hemisphere (for bilateral brain measures),
#parent depressive symptoms (fieldname: asr_scr_depress_r )

covs <- covs %>% 
  relocate("src_subject_id", "eventname", "sex", "age_years", "bmi", "race.6level", "demo_comb_income_p",  "abcd_site", "high.educ", "fsqc_qu_motion", "asr_scr_depress_r", "rel_family_id")

```
##Quality control covariates

We will get a summary of each covariate and check for number of NAs. These will be removed when running the linear models later.  

We will use the "rel_family_id" later to reduce sample to unrelated participants only. 

##Core Covariates

```{r, QC covs}
### Core covariates for basic model ###

#Note: Additional imaging covariates motion, hemisphere and ICV will be added later from PREP.imaging file when building models. 

#Age --- looks good, no NAs
summary(covs$age_years)

#BMI --- NAs = 94
summary(covs$bmi)

#Race/ethnicity --- NAs = 171

#Note: We are using a 6-level race measure instead of race_etnicity variable in an effort to make the measure more representative. "race.6level" includes White, Black, Asian, AIAN/NHPI (Asian Indian/Alaskan American/Native Hawaiian/Pacific Islander), Other, Mixed. 

summary(covs$race.6level)

#Study site --- why is there a site 22? 
summary(covs$abcd_site)

```

##Additional Covariates

```{r, QC covs}
### Additional Covariates ###

##Socio-economic variables

#Family income --- Refuse to answer=512, Don't know = 504, NAs= 2
summary(covs$demo_comb_income_p)

#Parent highest education --- NAs = 14.
summary(covs$high.educ)

#Parent depression --- NAs = 5. 
summary(covs$asr_scr_depress_r)

```

All covariates look within a normal range and can be exported. 

##Export cleaned data 

```{r, export data}
setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data")
saveRDS(covs,"covariates_cleaned_R3.0.rds")
```
####------------ END OF MAIN SCRIPT -----------------####



