---
title: "PREP_mplus_data"
author: "Niamh MacSweeney"
date: '2022-09-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction:

The purpose of this script is to prepare a dataframe for our mediation analysis in MPlus to answer Hypotheses 2 and 3 of MacSweeney et al.'s Registered Report. 

We will first load in our predictor (puberty: year 1) outcome (depression: year 3) and covariate data, alongside the mediator (imaging data: year 2).

We will create a confirmatory and exploratory dataframe for males and females separately because we are running independent models. 

Confirmatory df: Contains only the imaging variables identified in our pilot analyses. 


Exploratory df: Contains all imaging variables to test for additional mediating variables above and beyond what was identified in pilot analysis. 

#Set up
```{r, load libraries and set wd}

library(tidyverse)
library(ggplot2)
library(MplusAutomation)
library(tidySEM)
library(lavaan)

library(here)
library(gt)
library(gtsummary)
library(mediation)
library(carData)
library(plotly)
library(viridis)
library(iNZightTools) #to create combined categorical vars

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")

```

#1: LOAD DATA + TIDY

```{r, load data}

#puberty
pds_y1 <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/pds_timing_y1_R4.0.rds")
  
#depression
dep_vars <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/CBCL_dep_all_yrs_R4.0.rds")

#covariates
covs <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/covs_main_R4.0.rds")

#imaging

cortical <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/cortical_year2_R4.0.rds")
  
subcort <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/subcort_year2_R4.0.rds")
  
  
dti <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/dti_year2_R4.0.rds")
  

```

##Extract variables of interest before merge

The depression, puberty and covariate variables are already in reduced format so we just need to extract the imaging variables that we will test in our mediation analyses. 

Proposed mediators:

**Hypothesis 2: Global Measures**
**Global measures**
H2a: Reduced global cortical volume, surface area, thickness and sulcal depth.

H2b: Reduced global FA

**Hypothesis 3: Regional Measures** 
**Regional measures**
H3a: Reduced cortical thickness in temporal regions, namely, the middle temporal gyrus and insula.

H3b: Reduced cortical thickness in frontal regions namely, the lateral orbito-frontal cortex and middle frontal gyri.

H3c: Reduced cortical volume in temporal regions, namely, middle temporal gyrus and bank of the superior temporal sulcus.

H3d: Reduced cortical volume in fronto-parietal regions, namely, the middle frontal and postcentral gyri. 

H3e Reduced FA in the cortico-striatal tract and corpus collosum

H3f: Increased sulcal depth in the pars orbitalis.

H3g: Increased volume in the ventral diencephalon. 


```{r, extract imaging variables}

#open R data files from pilot analyses to get ROIs
#We have four sets of results containing global and regional ROIs

ptImgF <- load("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/revised_pilot/ROIs_pt_brain_females.Rdata")
  
ptImgM <- load("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/revised_pilot/ROIs_pt_brain_males.Rdata")
  
depImgF <- load("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/revised_pilot/ROIs_dep_brain_females.Rdata") 

depImgM <- load("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/revised_pilot/ROIs_dep_brain_males.Rdata")


```

###Get ROI names 

Note (20/09/22): It seems that postcentral surface area was also a ROI in the pilot analysis (ROI_dep_brain_males_regional_results.Rdata analysis) but only postcentral volume was included as an ROI to test at stage 2. I'm not sure how this happened? Given that this ROI does not withstand correction for multiple comparisons, I will not amend stage 1 to include this ROI. If it emerges as a significant mediator in our exploratory analyses, it will be reported as exploratory analyses. 






```{r, ROI names}

#global measures 
globalMeds <- c("bl.vol.APARC.total", "bl.sa.APARC.total", #H2a
                "bl.sulc.APARC.mean", "bl.thk.APARC.mean",
                "bl.dtifa_allfibers") #H2b

globalMeds

#regional measures
regionalMeds <- c(	
                  "avg.thk.APARC.mdtm", #temporal thickness ROIs (H3a)
                  "avg.thk.APARC.insula",
                  "avg.thk.APARC.rrmdfr", #frontal thickness  ROIs (H3b)
                  "avg.thk.APARC.mobfr",
                  "avg.thk.APARC.lobfr",
                  
                  "avg.vol.APARC.bankssts", #temporal volume ROIs (H3c)
                  "avg.vol.APARC.mdtm",
                  "avg.vol.APARC.cdmdfr", #frontal volume ROIs (H3d)
                  "avg.vol.APARC.postcn",
                  
                  
                  "avg.dtifa_pscs",# FA ROIs H3e
                  "dtifa_cc",
                  
                  "avg.sulc.APARC.parsobis", #H3f
                  "avg.vol.ASEG.vedc" #H3g
                  )
regionalMeds

```

##Exploratory mediators

See scripts ANALY_explr... for exploratory data analysis that was undertaken to identify additional ROIs. 

We undertook exploratory analyses to identify any additional ROIs that were not detected in the pilot analyses. The pilot analyses used baseline imaging data from the full sample while the confirmatory analyses used imaging data from year 2 in ~50% of the sample due to the partial year 3 data release at time of analysis. 

We decided to take a more specific approach to identify potential mediators in our exploratory analyses, such that we only included an ROI as a mediator if it was associated with both depression and puberty at year 3 and year 1 respectively. 

The volume of the accumbens area was the only region that met these criteria in females only. No brain structural measures met these criteria in males. 



```{r, exploratory mediator}

explrMeds <- c("avg.vol.ASEG.aa")

```



###Merge imaging variables and extract ROIs

```{r,merge imaging}

#inspect colnames before merge

head(cortical) #no sex column in dataframes so just merge by subject ID
head(subcort)
head(dti)

imgVars <- cortical  %>% 
  left_join(y=subcort, by= "src_subject_id") #merge 1 (N= 7695)

imgVars <- imgVars %>% 
  left_join(y=dti, by= "src_subject_id") 

colnames(imgVars)
  
  
```

#Extract mediators 

Note: Given that there is only one mediator to test from the exploratory analyses, we will add it to the main MPlus data files. However, we will still generate exploratory dataframe with all brain measures for reference (also, the exploratory data files created here in the MPlus naming format were used as input in the ANALY_explr scripts).

```{r, extract mediator vars}

mainImgVars <- imgVars %>% #20 variables
  select(src_subject_id, globalMeds, regionalMeds, explrMeds) 

```

##Generate imaging dataframe for exploratory analyses

This will include the average bilateral measures and the unilateral brain measures.
The current imaging dataframe has all the individual bilateral measures (i.e., left and right), which we aren't using in the current analyses. 

```{r, exploratory imaging vars}

explImgVars <- imgVars %>% 
  .[,grep('^avg|^bl|^dti|^src', colnames(.))]  
  #select cols starting with avg, bl, dti or src_subject 
  #should give us the cols we want

colnames(explImgVars) #col names look correct! 191 variables
```



STEP TWO: MAIN MERGE
## Main analysis dataframe
```{r, main df merge}

df <- covs %>% 
  left_join(y=dep_vars, by= c("src_subject_id", "sex"))
  
#merge 2
df <- df %>% 
  left_join(y=pds_y1, by= c("src_subject_id", "sex"))

#merge 3 (with imaging)

df <- df %>% 
  left_join(y=mainImgVars, by= "src_subject_id")

```

###Reduce dataframe to correct sample

We will only include people with complete exposure (puberty) data and who had available year 3 depression data at release 4.0. 

Total N = 5727

```{r, select correct sample}

df <- df %>% 
  filter(!is.na(pds_tot_all)) %>% #N=10,400
  filter(!is.na(interview_date_y3)) #N = 5727

```

Note: For our first hypothesis test, we are using multiple imputation to deal with missing data. Following our analysis to identify auxiliary variables that predicted out outcome data and/or missingness in our outcome data, we were unable to find suitable auxiliary variables that predicted race and household income. See script (ANALY_H1_females.Rmd) We therefore removed individuals with missing data for these variables from our analysis. 

For our second and third hypotheses, we will be using FIML to handle missing data, which is the default in R. To keep the sample sizes consistent across all hypotheses, we will remove people with missing race and income data also. 

This results in a final N = 5325 (males and females combined)

```{r, further df filtering}

df <- df %>% #5727
  filter(!is.na(household_income_y0)) %>% #5371
  filter(!is.na(race_y0)) #5325

```

##exploratory analysis dataframe

```{r, exploratory df merge}

expl_df <- covs %>% 
  left_join(y=dep_vars, by= c("src_subject_id", "sex"))
  
#merge 2
expl_df <- expl_df %>% 
  left_join(y=pds_y1, by= c("src_subject_id", "sex"))

#merge 3 (with imaging)

expl_df <- expl_df %>% 
  left_join(y=explImgVars, by= "src_subject_id")

```

Reduce to correct sample as per explanation above 

```{r, reduce expl df to correct sample}

expl_df <- expl_df %>% 
  filter(!is.na(pds_tot_all)) %>% #N=10,400
  filter(!is.na(interview_date_y3)) %>%  #N = 5727
    filter(!is.na(household_income_y0)) %>% #5371
  filter(!is.na(race_y0)) #5325 

colnames(expl_df)
head(expl_df) #merge and filter looks okay 

```

Remove extra columns that we don't need: 

- remove extra variables NOT used in imputation
- remove puberty variables.

Note, we will make one large list with all the extra variables and then remove using select function. 

```{r, tidy dataset}

extraVars <- c("site_id_6m", "site_id", "site_id_18m", #no longer need these
               "site_id_y2", "site_id_y3",
               "parent_dep_y3",
               "acs_raked_propensity_score_y1", "bmi_y2",
               "race_6m", "race_y1", "race_18m",
               "race_y2", "dti_mean_motion_y0",
               "interview_date.baseline_year_1_arm_1",
               "parent_high_ed_y0",
               "interview_date_y0", "interview_date_y1", "interview_date_y2",
               "interview_date_y3", 
               "interview_age_y1", "interview_age_y2", "interview_age_y3",
               "pds_p_ss_female_category", 
               "pds_p_ss_male_category",
               "age_years", "age_y1", #extra age variable after merge
               "pds_tot_m", "pds_tot_f" ,"pds_tot_all", 
               "pds_1_p", "pds_2_p",  "pds_3_p", "pds_m5_p",    
               "pds_m4_p", "pds_f4_p","pds_f5b_p",
               "pds_m_pt_valid", "pds_f_pt_valid"
                )

#reduce dataframe by removing extra cols. 
df <- df %>% 
  select(-(all_of(extraVars))) 

colnames(df) #looks good

```

#2: MPLUS DATA prep
Mplus names can't be longer than 8 characters or else they will be truncated 

## Main analysis
###Renaming
```{r, rename for main df for Mplus}

#let's start with imaging
#remove "avg." prefix and "APARC" and "ASEG" joiners

colnames(df)=gsub('avg.','',colnames(df)) #remove avg. prefix
colnames(df)=gsub('APARC.','',colnames(df)) #remove APARC
colnames(df)=gsub('ASEG.','',colnames(df)) # remove ASEG
colnames(df)=gsub('bl.','',colnames(df)) #remove bl
colnames(df)=gsub('total','glb',colnames(df)) #shorten total to global to indicate it's a global measure
colnames(df)=gsub('mean','glb',colnames(df)) #shorten mean to global to indicate global measure. 

#truncate regional metrics: vol = V, thk = T, sa = SA, sulc = SL
colnames(df)=gsub('vol','V',colnames(df))
colnames(df)=gsub('thk','T',colnames(df))
colnames(df)=gsub('sa','SA',colnames(df))
colnames(df)=gsub('sulc','SL',colnames(df))

#some names are still too long e.g., bankssts and parsobis so need to shorten 
colnames(df)=gsub('bankssts','bsts',colnames(df))
colnames(df)=gsub('parsobis','porb',colnames(df))

#finally, shorten the dti measure names 
colnames(df)=gsub('dtifa_cc','FA.cc',colnames(df))
colnames(df)=gsub('dtifa_pscs','FA.pscs',colnames(df))
colnames(df)=gsub('dtifa_allfibers','FA.glb',colnames(df)) #global FA
colnames(df)=gsub('dti_glb_motion_y2','dti_mot',colnames(df)) #dti motion

#check col names
colnames(df)

#### non imaging variables 

df <- df %>% 
  dplyr::rename( id = src_subject_id, 
          p_dep_y2 = parent_dep_y2,
          site_id = site_id_y1,
          fam_id = rel_family_id,
          pws = acs_raked_propensity_score_y0,
          scnr_id = scanner_id_y2,
          income = household_income_y0,
          y_dep_y3 = cbcl_withdep_y3,  #youth depression year 3
          y_dep_y0 = cbcl_withdep_y0, #aux vars needed for FIML
          y_dep_y1 = cbcl_withdep_y1, 
          y_dep_y2 = cbcl_withdep_y2,
          y_anx_y0 = cbcl_anxdep_y0,
          y_anx_y1 = cbcl_anxdep_y1,
          y_anx_y2 =cbcl_anxdep_y2, 
          y_anx_y3 = cbcl_anxdep_y3, 
          p_dep_y0 = parent_dep_y0
          )
        
colnames(df)

```
###Formating

We also need to create a combined cluster variable for scanner_id and family_id due to issues with multiple clustering variables in MPlus. We will use the combineCatVars variable to do this. This needs to be done before variables are converted to numeric due to function requirements. 

All variables need to be numeric and missing values need to be indicated with "-9999". 

Before converting character variables to factors and then numeric, we will get a list of all the factors and values for future reference. 

Note: income is already numeric, described below:

1= Less than $5,000; 2=$5,000 through $11,999; 3=$12,000 through $15,999; 4=$16,000 through $24,999; 5=$25,000 through $34,999; 6=$35,000 through $49,999; 7=$50,000 through $74,999; 8= $75,000 through $99,999; 9=$100,000 through $199,999; 10=$200,000 and greater

N.B: Sex variable labelling: 1= F, 2= M

```{r, Mplus formatting}

# #Create cluster variable
# 
df$fam_id <- as.character((df$fam_id)) #make fam id character var
# 
# df <- combineCatVars(df, 
#                      vars = c("scnr_id", "fam_id"),
#                      sep = "_",
#                      keep_empty = TRUE)
# head(df) #check new var looks okay
# str(df)
# 
# #rename cluster variable so it short enough for MPlus! 
# #N.B. c_vars = scnr_id_fam_id
# 
# df <- df %>% 
#   rename(c_vars = scnr_id_fam_id)

#char vars
charVars <- c("id", "sex", "site_id",
              "scnr_id", "fam_id", "race_y0")

#convert character variables to factor
df <- df %>%
  mutate_at(charVars, as.factor)

#get levels for sex and race for future reference
# the other variables aren't likely needed for further interpretation

df %>% 
  select(sex, race_y0) %>% 
     sapply(levels)
#check order of labeling with 
str(df)

#convert factor variables to numeric 
df <- df %>% 
  mutate_at(charVars, as.numeric)

#check 
str(df) # all variables are either numeric or integer 

summary(df)


### change NA values to 

df[is.na(df)] <- -9999 

head(df) #looks good!


```
##Export data files in R format

```{r, }
#females
F_df <- df %>% 
  filter(sex == 1) %>%  #females only  N= 2533
  select(- c(pt_m))

#males
M_df <- df %>% 
  filter(sex == 2) %>%  #males only M = 2792
  select(- c(pt_f))

#whole sample
```


```{r, save RDS files }
#Save RDS files 

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data")

saveRDS(F_df, "female_df_2022_10_12.R4.0.rds")

saveRDS(M_df, "male_df_2022_10_12.R4.0.rds")

saveRDS(df, "full_sample_df_2022_10_12.R4.0.rds")


```







##Exploratory analysis 

###Renaming
Renaming set up should be the same but for some of the imaging vars, the name will be longer than 8 but MPlus should truncate this. If is causes an issue, we can revisit later. 

```{r, rename exploratory df for Mplus}

#first remove extraVars (should have this object in environment already)

expl_df <- expl_df %>% 
  select(-(all_of(extraVars))) #should have 211 cols

colnames(expl_df) #looks good

#rename imaging vars 
colnames(expl_df)=gsub('avg.','',colnames(expl_df)) #remove avg. prefix
colnames(expl_df)=gsub('APARC.','',colnames(expl_df)) #remove APARC
colnames(expl_df)=gsub('ASEG.','',colnames(expl_df)) # remove ASEG
colnames(expl_df)=gsub('bl.','',colnames(expl_df)) #remove bl
colnames(expl_df)=gsub('total','glb',colnames(expl_df)) #shorten total to global to indicate it's a global measure
colnames(expl_df)=gsub('mean','glb',colnames(expl_df)) #shorten mean to global to indicate global measure. 

#truncate regional metrics: vol = V, thk = T, sa = SA, sulc = SL
colnames(expl_df)=gsub('vol','V',colnames(expl_df))
colnames(expl_df)=gsub('thk','T',colnames(expl_df))
colnames(expl_df)=gsub('sa','SA',colnames(expl_df))
colnames(expl_df)=gsub('sulc','SL',colnames(expl_df))

#some names are still too long e.g., bankssts and parsobis so need to shorten 
colnames(expl_df)=gsub('bankssts','bsts',colnames(expl_df))
colnames(expl_df)=gsub('parsobis','porb',colnames(expl_df))

#finally, shorten the dti measure names 
colnames(expl_df)=gsub('dtifa_','FA.',colnames(expl_df))
colnames(expl_df)=gsub('dtimd_','MD.',colnames(expl_df))
colnames(expl_df)=gsub('FA.allfibers','FA.glb',colnames(expl_df)) #global FA
colnames(expl_df)=gsub('MD.allfibers','MD.glb',colnames(expl_df))
colnames(expl_df)=gsub('dti_glb_motion_y2','dti_mot',colnames(expl_df))

#check 
colnames(expl_df) #looks good!

### non imaging variables 

expl_df <- expl_df %>% 
  rename( id = src_subject_id, 
          p_dep_y2 = parent_dep_y2,
          fam_id = rel_family_id,
          pws = acs_raked_propensity_score_y0,
          scnr_id = scanner_id_y2,
          income = household_income_y0,
          y_dep_y3 = cbcl_withdep_y3,  #youth depression year 3
          y_dep_y0 = cbcl_withdep_y0, #aux vars needed for FIML
          y_dep_y1 = cbcl_withdep_y1, 
          y_dep_y2 = cbcl_withdep_y2,
          y_anx_y0 = cbcl_anxdep_y0,
          y_anx_y1 = cbcl_anxdep_y1,
          y_anx_y2 =cbcl_anxdep_y2, 
          y_anx_y3 = cbcl_anxdep_y3, 
          p_dep_y0 = parent_dep_y0)
        
colnames(expl_df)


```
###R Export for RR exploratory analyses
```{r, for use in R}

F_expl_df <- expl_df %>% 
  filter(sex == "F") %>%  #females only  N= 2533
  select(- c(pt_m))

M_expl_df <- expl_df %>% 
  filter(sex == "M") %>%  #males only M = 2792
  select(- c(pt_f))

#Save RDS files 

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data")

saveRDS(F_expl_df, "explr_analy_F_data.R4.0.rds")

saveRDS(M_expl_df, "explr_analy_M_data.R4.0.rds")

```


###Formatting for Mplus
```{r, formatting for MPlus}

#char vars
charVars <- c("id", "sex", "site_id_y1",
              "scnr_id", "race_y0")

#convert character variables to factor
expl_df <- expl_df %>%
  mutate_at(charVars, as.factor) %>% 
   mutate_at(charVars, as.numeric)

#check 
str(expl_df)
colnames(expl_df)

summary(expl_df)

### change NA values to -9999

expl_df[is.na(expl_df)] <- -9999

head(expl_df) #looks good!

F_expl_df <- expl_df %>% 
  filter(sex == 1) %>%  #females only  N= 2533 N.B - sex = 1 for females
  select(- c(pt_m))

M_expl_df <- expl_df %>% 
  filter(sex == 2) %>%  #males only M = 2792 N.B. sex = 2 for males 
  select(- c(pt_f))


#Save RDS files with Mplus formatting (ID = numeric value)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data")

saveRDS(F_expl_df, "explr_analy_F_Mplus_data.R4.0.rds")

saveRDS(M_expl_df, "explr_analy_M_Mplusdata.R4.0.rds")

```

The data frames are not in the correct format for MPlus

#3: Split by sex 


##Main analysis
```{r, split by sex}

F_df <- df %>% 
  filter(sex == 1) %>%  #females only  N= 2533
  select(- c(pt_m))

M_df <- df %>% 
  filter(sex == 2) %>%  #males only M = 2792
  select(- c(pt_f))


```

##Exploratory analysis
```{r, split by sex}

F_expl_df <- expl_df %>% 
  filter(sex == 1) %>%  #females only  N= 2533
  select(- c(pt_m))

M_expl_df <- expl_df %>% 
  filter(sex == 2) %>%  #males only M = 2792
  select(- c(pt_f))

```

#Explore data

## Mplus
```{r, make data files for Mplus}

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/Mplus_inputs")

female_df <- prepareMplusData(F_df, "female_df.dat")

male_df <- prepareMplusData(M_df, "male_df.dat")

female_expl_df <- prepareMplusData(F_expl_df, "female_explr_df.dat")

male_expl_df <- prepareMplusData(M_expl_df, "male_explr_df.dat")

```

##Exploratory Mediation dataframe 

Note (17/10/22): For ease of data cleaning, we will include the exploratory mediators here and then create a separate dataframe that includes both the confirmatory and exploratory mediators. 

The exploratory mediators were determined by running the PT~ brain and Dep~ brain associations using the follow up data waves PT Year 1, brain Year 2, and dep year 3. See "ANALY_explr_..." scripts for associated code. 

- In this dataframe, we need all the variables included in the confirmatory analysis dataframe, plus the extra exploratory ROIs, and the random factor clustering variable. 

Let's do this separately for females and males


##Females
```{r, females exploratory mediation}

#Load a random factor clustering variable that we will use in MPlus. This was created in Stata by J.A. 
clusterVar <- load("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/MplusAuto/judith_files/sex_1.Rdata")

#keep id and cluster variable only 
clusterVar <- sex_1 %>% 
  select(id, c)

#note, exploratory variables
explorMeds <- c("T.paracn", "T.rrmdfr", "T.insula", "V.sufr", "T.cdmdfr", "V.paracn",
          "V.rrmdfr", "T.sufr", "V.insula" , "V.cdmdfr", "T.lingual", "V.bsts", 
          "V.pc", "V.supl", "V.aa", "T.supl", "T.ifpl", "V.mdtm", "V.fusiform",# pt regional ROIs
          "T.glb", "V.glb", "V.vedc", #only male pt ROI = vedc
          "SL.bsts", "SL.precn", "V.sm", "MD.ifo", #F dep ROIs
          "SA.mobfr", "V.aa", "V.pallidum", "V.tp" #M dep ROIs
          )

confirmMeds <- c("V.glb", "SA.glb", "SL.glb", "T.glb", "FA.glb", "T.mdtm", "T.insula", "T.rrmdfr",
                 "T.mobfr", "T.lobfr", "V.bsts", "V.mdtm", "V.cdmdfr", "V.postcn","FA.pscs", "FA.cc", 
                 "SL.porb", "V.vedc")



```

Merge with exploratory data frame 

N.B- This dataframe is in Mplus format so missing values at -9999 and all variables are numeric in type. 

```{r, merge females}

F_expl_df <- F_expl_df %>% 
  left_join(y=clusterVar, by= "id") #looks good. Should have 227 vars

#main non imaging vars - create list to make variable extraction easier below

mainVars <- colnames(F_expl_df[1:36]) # extract non imaging vars
mainVars #looks good 

cluster <- "c"

#make main list with all vars needed 
allVars <- c(mainVars, confirmMeds, explorMeds, cluster) #85 vars

allVars <- unique(allVars) #remove 8 duplicate ROIs between confirmatory and exploratory. Should have 76
allVars

F_expl_df <- F_expl_df[, allVars] #subset to get vars needed

```

##Males

N.B. Sex = 2 and missing data are = -9999 so we very careful to change missing values back to NA if using this dataframe outside of MPlus. 

```{r, generate exploratory mediation df}

#Load a random factor clustering variable that we will use in MPlus. This was created in Stata by J.A. 
clusterVar <- load("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/MplusAuto/judith_files/sex_2.Rdata")

#keep id and cluster variable only 
clusterVar <- sex_2 %>% 
  select(id, c)

#note, exploratory variables
explorMeds <- c("T.paracn", "T.rrmdfr", "T.insula", "V.sufr", "T.cdmdfr", "V.paracn",
          "V.rrmdfr", "T.sufr", "V.insula" , "V.cdmdfr", "T.lingual", "V.bsts", 
          "V.pc", "V.supl", "V.aa", "T.supl", "T.ifpl", "V.mdtm", "V.fusiform",# pt regional ROIs
          "T.glb", "V.glb", "V.vedc", #only male pt ROI = vedc
          "SL.bsts", "SL.precn", "V.sm", "MD.ifo", #F dep ROIs
          "SA.mobfr", "V.aa", "V.pallidum", "V.tp" #M dep ROIs
          )

confirmMeds <- c("V.glb", "SA.glb", "SL.glb", "T.glb", "FA.glb", "T.mdtm", "T.insula", "T.rrmdfr",
                 "T.mobfr", "T.lobfr", "V.bsts", "V.mdtm", "V.cdmdfr", "V.postcn","FA.pscs", "FA.cc", 
                 "SL.porb", "V.vedc")


M_expl_df <- M_expl_df %>% 
  left_join(y=clusterVar, by= "id") #looks good. Should have 227 vars

#main non imaging vars - create list to make variable extraction easier below

mainVars <- colnames(M_expl_df[1:36]) # extract non imaging vars
mainVars #looks good 

cluster <- "c"

#make main list with all vars needed 
allVars <- c(mainVars, confirmMeds, explorMeds, cluster) #85 vars

allVars <- unique(allVars) #remove 8 duplicate ROIs between confirmatory and exploratory. Should have 76
allVars

M_expl_df <- M_expl_df[, allVars] #subset to get vars needed
head(M_expl_df) #looks good to go! 



```
##Export .dat files 

```{r, export .dat files}

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/Mplus_inputs")

femalesExplMed <- prepareMplusData(F_expl_df, "fExplMed.dat")
malesExplMed <- prepareMplusData(M_expl_df, "mExplMed.dat")


```



############# END OF MAIN SCRIPT ###############


#Column ordering for Mplus 

```{r}

custom_glimpse <- function(x) {
  data.frame(
    col_name = colnames(x),
    col_index = 1:ncol(x),
    col_class = sapply(x, class),
    row.names = NULL
  )
}

x = F_df

female_codebook <- custom_glimpse(x)
female_codebook

write.csv(female_codebook,"/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/FemaleMplusCodebook.csv", row.names = TRUE)

#males

custom_glimpse <- function(y) {
  data.frame(
    col_name = colnames(y),
    col_index = 1:ncol(y),
    col_class = sapply(y, class),
    row.names = NULL
  )
}

y = M_df

male_codebook <- custom_glimpse(y)
male_codebook

write.csv(male_codebook,"/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/MaleMplusCodebook.csv", row.names = TRUE)

```





#4: MplusAutomation 

See this paper: https://www.tandfonline.com/doi/full/10.1080/10705511.2017.1402334

Remember to cite this package in paper!

```{r, get list of mediators}
#need to make new list after earlier renaming
globalMeds <- c("V.glb", "SA.glb", "SL.glb", "T.glb" ,"FA.glb")

regionalMeds <- c("T.mdtm", "T.insula", "T.rrmdfr" , "T.mobfr" , 
                  "T.lobfr" , "V.bsts",   "V.mdtm" , "V.cdmdfr", 
                  "V.postcn", "FA.pscs", "FA.cc", "SL.porb", 
                  "V.vedc")

```


```{r, get variables table with label for Mplus}

#Example mediation model
test_table <- tribble(
   ~"Name",      ~"Labels",                                     
 #--------------|--------------------------------|,
  "pt_f (X)" , "Pubertal timing (females)"         ,
  "V.glb (M)" , "Global brain volume"          ,
  "y_dep_y3 (Y)" , "Youth depresison at year 3"   )

test_table %>% gt()

```
Females 
```{r, run Mplus from R}

#we define our model here
?mplusObject #gives instructions for Mplus syntax
m1 <- mplusObject(
  TITLE = 
    "Females: Global Brain Measures SEM;",

  VARIABLE = 
  
    "usevar = 
      pt_f  V.glb  y_dep_y3;",
  
  ANALYSIS =  
    "bootstrap = 500;  ! set number of bootstrap   draws (small # for demonstration purposes)" , 
  
    # y_dep_y3        !!! outcome (Y)
    # V.glb           !!! mediator (M)
    # pt_f            !!! predictor (X)
  MODEL = 
 
  "y_dep_y3 on V.glb pt_f;   !!! Y on M X
  V.glb on pt_f;             !!! M on X Y
    
    
    Model indirect:
    y_dep_y3 ind V.glb pt_f;",
     
  OUTPUT = 
    "standardized;",
  
  PLOT = "type=plot2;",
  rdata = F_df)     

?mplusModeler
setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/Mplus")

m1_fit <- mplusModeler(m1,
                modelout = 
                  "/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/results/Mplus/m1.inp",
                run = TRUE)

SummaryTable(m1_fit)

```

```{r}


test <- lm(y_dep_y3 ~ FA.glb, data = F_df, na.action = na.exclude)
```


```{r, check dat files}

datFile <- readLines("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/med_preprocess/sex_1.dat", 
           n=10)

```




