---
title: "PREP_hormones.Rmd"
author: "Niamh MacSweeney"
date: "7/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ABCD Hormone Cleaning Script 

## Introduction

The purpose of this script is to tidy and quality control the pubertal hormone data from ABCD release3.0. 

#SETUP

```{r load libraries and set wd }
library(tidyr)
library(dplyr)
library(tidyverse)
library(psych)
library(ggplot2)
library(ggstatsplot)
library(GGally)
library(outliers)
library(jtools) #great package for summarising regression results
library(readr)
library(gridExtra)
library(sandwich) #for linear models

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")
getwd()
```

# LOAD DATA, REDUCE VARIABLES, CREATE PUBERTY DATAFRAME, TIDY VARIABLES
 
```{r load data, tidy variables}

hormones <-rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/abcd_hsss01.rds") #salimetric hormones
hormone_notes <-rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/sph01.rds") ##notes on hormone data

#Reduce to baseline data only. N= 11875
hormones <- subset(hormones, eventname=="baseline_year_1_arm_1")
hormone_notes <- subset(hormone_notes, visit=="baseline_year_1_arm_1")

#Merge and create pubertal hormone master dataframe. 
#drop repeated columns from hormone_notes before merge
hormone_notes <- hormone_notes %>% select(-interview_date, -interview_age, -sex, -visit)
#merge hormones and hormone_notes to form df
ph_df <- merge(hormones, hormone_notes, by="src_subject_id") 


#Convert age in months to years, save new variable
ph_df <- ph_df %>%
  mutate(
    age_years = interview_age/12 
  )

colnames(ph_df) #check new columns
```

#INSPECT AND FILTER DATA

#STEP1: Check for and remove missing values for sex variables. Remove NAs: N=11878 --> 11875)
```{r check for and remove missing values}

ph_df %>% 
  dplyr::count(sex) # no blanks detected. 

ph_df %>% 
  dplyr::count(hormone_sal_sex) # N=3 NAs 
ph_df <-  ph_df[!(is.na(ph_df$hormone_sal_sex) | ph_df$hormone_sal_sex==""), ] 

table(ph_df$sex, ph_df$hormone_sal_sex) #Check missing values have been removed and now check for mismatches. 

```
#STEP 2: Check that sex matches hormone sex given. Remove data that is unusable: N= 11875 --> 11766

18 male tubes miscaslsified as female, 6 female tubes misclassified as male ---> REMOVE (N=33)
59 unable to complete sample, 19 refused to give sample, 7 sample not collected/other ---> REMOVE (N=81)
```{r sex and hormone sex matching}

ph_df %>% 
  dplyr::count(hormone_sal_sex)

#Remove participants if following conditions are met:
ph_df  <- ph_df[-c(which(ph_df$sex == "M" & ph_df$hormone_sal_sex == 1), #male sex, female labelled tube
             which(ph_df$sex =="F" & ph_df$hormone_sal_sex == 2), #female sex, male labelled tube
             which(ph_df$hormone_sal_sex == 3), #participant unable to complete sample 
             which(ph_df$hormone_sal_sex == 4), #participant/parent refused sample:
             which(ph_df$hormone_sal_sex == 5)),] # sample not collected 

```

#STEP 3: Check for and remove hormone missing values.

First check number of NAs to get estimate of complete cases
Count number of participant with NAs for three hormone mean scores
```{r determine hormone complete cases}
sum(is.na(ph_df$hormone_scr_dhea_mean)) #NAs= 866; complete cases= 10900
sum(is.na(ph_df$hormone_scr_ert_mean)) #NAs= 767; complete cases= 10999
sum(is.na(ph_df$hormone_scr_hse_mean)) #NAs= 6672; complete cases = 5094 (females only)

```

Then use the filter scheme that used by Herting & Uban et al., (2021) to check records for any RA saliva collection notes.
If true, then flag the record. 
Then check flagged records and see if the Salimetrics value is out of range per hormone.
If yes, then change value to NA, else keep the existing values for each replicate.
Finally, average the two replicates into a new field.
Note: variable hormon_sal_notes_y___1 is not included here because it is a summary measure, we are interested in the items independently.

#FILTER DATA

#STEP 1: Count number of participants with flagged records.
Method 1: based on yes/no concerns about sample. 
No concerns(1): N= 10,025, Concerns(0): N= 1740
Could remove all sample with any flagged records - this would be the most conservative approach. 

Method 2: Instead, we are going to remove values that fall outside the range for that specific hormone, method outline below. 
```{r, filtering hormone data}
#count concerns based on yes/no criteria
ph_df %>% 
  dplyr::count(hormon_sal_notes_y___1)

#look at number flagged cases and see whether they are in range for each hormone — we will use this approach
ph_df$hormone_notes_sum <- as.numeric(ph_df$hormon_sal_notes_y___2) + 
  as.numeric(ph_df$hormon_sal_notes_y___3) +
  as.numeric(ph_df$hormon_sal_notes_y___4) +
  as.numeric(ph_df$hormon_sal_notes_y___5) +
  as.numeric(ph_df$hormon_sal_notes_y___6)

summary(ph_df$hormone_notes_sum) #check min and max.Values should be within 0-5 range.

rownums <- which(ph_df$hormone_notes_sum >= 1) #get the row numbers that have at least 1 flagged record. N=1948

```

#STEP 2: Filter data to remove participants with unusable hormone levels (i.e., values outside range) for each hormone separately.

Note: DHEA, Testosterone, and Estradiol Ranges, respectively, as per Herting and Uban et al. (2021):
Calibrator ranges (10.2–1,000 pg/ml; 6.1–600 pg/ml; 1–32 pg/ml); 
Lower limits of sensitivity (5 pg/ml; 1 pg/ml; 0.1 pg/ml).

#DHEA

Complete cases for filtered DHEA = 10,931 11011
```{r, filter DHEA}

#Generate new columns for filtered data
ph_df$filtered_dhea <- as.numeric(NA)
ph_df$filtered_dhea_rep1 <- as.numeric(ph_df$hormone_scr_dhea_rep1)
ph_df$filtered_dhea_rep2 <- as.numeric(ph_df$hormone_scr_dhea_rep2)

#Is the replicate too low for detection (e.g., nds)? 
#If Yes (val=1), then change to 0 so that it will be removed a later "value range" QC step.
ph_df %>% 
  dplyr::count(hormone_scr_dhea_rep1_nd) #R1: N=85 (NA = 685)

ph_df %>% 
  dplyr::count(hormone_scr_dhea_rep2_nd) #R2: N=111 (NA = 685)

ph_df$filtered_dhea[which(ph_df$hormone_scr_dhea_rep1_nd == 1)] <- 0 #change any nds=1 values to 0 for R1
ph_df$filtered_dhea[which(ph_df$hormone_scr_dhea_rep2_nd == 1)] <- 0 #change any nds=1 values to 0 for R2
rownums_rep1 <- which(ph_df$hormone_scr_dhea_rep1 < 5 | ph_df$hormone_scr_dhea_rep1 > 1000) #isolate R1 rows that are outside range.
rownums_rep2 <- which(ph_df$hormone_scr_dhea_rep2 < 5 | ph_df$hormone_scr_dhea_rep2 > 1000) #repeat for R2
ph_df$filtered_dhea_rep1[rownums[which(rownums %in% rownums_rep1)]] <- NA #change out of range R1 values to NA
ph_df$filtered_dhea_rep2[rownums[which(rownums %in% rownums_rep2)]] <- NA #repeat for R2

#average two replicates to new variable
ph_df$filtered_dhea <- apply(ph_df[, c("filtered_dhea_rep1", "filtered_dhea_rep2")], 1, function(x) mean(x, na.rm=T)) 

sum(is.na(ph_df$filtered_dhea)) #NAs= 835. Complete cases = 10931  (Remember to use total N=11766 to calc complete cases) 
#Same as Herting & Uban QC when done with release 2.0 data. 

```

#Testosterone

Complete cases for filtered testosterone N = 11011
```{r, filter Testosterone}

#Generate new columns for filtered data
ph_df$filtered_testosterone <- as.numeric(NA)
ph_df$filtered_testosterone_rep1 <- as.numeric(ph_df$hormone_scr_ert_rep1)
ph_df$filtered_testosterone_rep2 <- as.numeric(ph_df$hormone_scr_ert_rep2)

#Is the replicate too low for detection (e.g., nds)? 
#If Yes (val=1), then change to 0 so that it will be removed a later "value range" QC step.
ph_df %>% 
  dplyr::count(hormone_scr_ert_rep1_nd) #R1: N=1 (NA=743)

ph_df %>% 
  dplyr::count(hormone_scr_ert_rep2_nd) #R2: N=7 (NA=743)

ph_df$filtered_testosterone[which(ph_df$hormone_scr_ert_rep1_nd == 1)] <- 0 #change any nds=1 values to 0 for R1
ph_df$filtered_testosterone[which(ph_df$hormone_scr_ert_rep2_nd == 1)] <- 0 #change any nds=1 values to 0 for R2
rownums_rep1 <- which(ph_df$hormone_scr_ert_rep1 < 1 | ph_df$hormone_scr_ert_rep1 > 600) #isolate R1 rows that are outside range.
rownums_rep2 <- which(ph_df$hormone_scr_ert_rep2 < 1 | ph_df$hormone_scr_ert_rep2 > 600) #repeat for R2
ph_df$filtered_testosterone_rep1[rownums[which(rownums %in% rownums_rep1)]] <- NA #change out of range R1 values to NA
ph_df$filtered_testosterone_rep2[rownums[which(rownums %in% rownums_rep2)]] <- NA #repeat for R2

#average two replicates to new variable
ph_df$filtered_testosterone <- apply(ph_df[, c("filtered_testosterone_rep1", "filtered_testosterone_rep2")], 1, function(x) mean(x, na.rm=T)) 

sum(is.na(ph_df$filtered_testosterone)) #NAs = 755 Complete cases= 11011
```

#Estradiol

This is for females only. Complete cases for filtered Estradiol N = 5163
```{r}
#Generate new columns for filtered data
ph_df$filtered_estradiol <- as.numeric(NA)
ph_df$filtered_estradiol_rep1 <- as.numeric(ph_df$hormone_scr_hse_rep1)
ph_df$filtered_estradiol_rep2 <- as.numeric(ph_df$hormone_scr_hse_rep2)

#Is the replicate too low for detection (e.g., nds)? 
#If Yes (val=1), then change to 0 so that it will be removed a later "value range" QC step.
ph_df %>% 
  dplyr::count(hormone_scr_hse_rep1_nd) #R1: N=120 (NA = 6419, higher because females only)

ph_df %>% 
  dplyr::count(hormone_scr_hse_rep2_nd) #R2: N=127 (NA as above)

ph_df$filtered_estradiol[which(ph_df$hormone_scr_hse_rep1_nd == 1)] <- 0 #change any nds=1 values to 0 for R1
ph_df$filtered_estradiol[which(ph_df$hormone_scr_hse_rep2_nd == 1)] <- 0 #change any nds=1 values to 0 for R2
rownums_rep1 <- which(ph_df$hormone_scr_hse_rep1 < 0.1 | ph_df$hormone_scr_hse_rep1 > 32) #isolate R1 rows that are outside range.
rownums_rep2 <- which(ph_df$hormone_scr_hse_rep2 < 0.1 | ph_df$hormone_scr_hse_rep2 > 32) #repeat for R2
ph_df$filtered_estradiol_rep1[rownums[which(rownums %in% rownums_rep1)]] <- NA #change out of range R1 values to NA
ph_df$filtered_estradiol_rep2[rownums[which(rownums %in% rownums_rep2)]] <- NA #repeat for R2

#average two replicates to new variable
ph_df$filtered_estradiol <- apply(ph_df[, c("filtered_estradiol_rep1", "filtered_estradiol_rep2")], 1, function(x) mean(x, na.rm=T)) 

sum(is.na(ph_df$filtered_estradiol)) #NAs = 6603 Complete cases= 5163 
```

#Distribution Checks
Once each hormone has been filtered, we will check the distribution and log-transform if needed

```{r, distribution and log transformation}
#check distribution - data is skewed, as expected, so we will log10 transform raw hormone levels before generating pt measure. 

#histograms
dhea_hist <- ggplot(data=ph_df, aes(x= filtered_dhea)) +
  geom_histogram(color="black", fill = "deepskyblue4",
                 bins =30) +
  ggtitle("Filtered DHEA")
print(dhea_hist)

testosterone_hist <- ggplot(data=ph_df, aes(x= filtered_testosterone)) +
  geom_histogram(color="black", fill = "deepskyblue4",
                 bins =30) +
  ggtitle("Filtered Testosterone")
print(testosterone_hist)

estradiol_hist <- ph_df %>%
  ggplot(aes(x= filtered_estradiol)) +
  geom_histogram(color="black", fill = "deepskyblue4",
                 bins =30) +
  ggtitle("Filtered Estradiol")
print(estradiol_hist)


#Knit histograms together
hormones_hist <- grid.arrange(dhea_hist, testosterone_hist, estradiol_hist, ncol = 3)
print(hormones_hist)

#check skewness 
skewness <- ph_df %>% 
  select(filtered_dhea, filtered_testosterone, filtered_estradiol) %>% 
skew()
skewness

#check kurtosis
kurtosis <- ph_df %>% 
  select(filtered_dhea, filtered_testosterone, filtered_estradiol) %>% 
kurtosi()
kurtosis

#Data is skewed so we will log10 transform the data. 



#Log10 transformed data
dhea_log_hist <- ggplot(ph_df, aes(x = filtered_dhea)) + 
              geom_histogram(fill = "deepskyblue4",
                              bins =30) + 
              scale_x_log10()

#plot pre-post log10 transformation. 
dhea_log_compar_hist <- grid.arrange(dhea_hist, dhea_log_hist, ncol = 2,
             top = "DHEA values before (left) and after log10 transformation (right)")

#change wd and save plot.
setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs")
ggsave("dhea_log_compar_hist.pdf")

#Let's look at summary of data before and after log10 transformation

summary(ph_df$filtered_dhea)

summary(log10(ph_df$filtered_dhea)) #note: if min value=0, should add +1 to avoid neg infinity. 


#log10-transform and save as new variable. 

ph_df <- ph_df %>%
  mutate(dhea_log = log10(filtered_dhea))
  
colnames(ph_df) #check new col exists. 

#Check for outliers after transformation

outl_dhea <- ph_df %>% 
  filter(mean(dhea_log) + 3*(sd(dhea_log)) & mean(dhea_log) - 3*(sd(dhea_log)))
        
#check outlier removal
outl_dhea_upper <- (mean(ph_df$dhea_log) - 3*(sd(ph_df$dhea_log)))
                             
```

#Log10 transforming data 

```{r, log transform}

ph_df <- ph_df %>%
  mutate(dhea_log = log10(filtered_dhea), 
         test_log = log10(filtered_testosterone),
         estr_log = log10(filtered_estradiol))

#Let's look at summary of data before and after log10 transformation
#Note: if min value=0, should add +1 to avoid neg infinity - not the case here. 

ph_df %>% 
  select(filtered_dhea, dhea_log, filtered_testosterone, test_log, filtered_estradiol, estr_log) %>% 
summary()

#Let's plot the values after log transformation

dhea_log_hist <- ggplot(data=ph_df, aes(x= dhea_log)) +
  geom_histogram(colour = "black", fill = "aquamarine4",
                 bins =30) +
  ggtitle("Log10 DHEA")
print(dhea_log_hist)

test_log_hist <- ggplot(data=ph_df, aes(x= test_log)) +
  geom_histogram(colour = "black", fill = "aquamarine4",
                 bins =30) +
  ggtitle("Log10 Testosterone")
print(test_log_hist)

estr_log_hist <- ggplot(data=ph_df, aes(x= estr_log)) +
  geom_histogram(colour = "black", fill = "aquamarine4",
                 bins =30) +
  ggtitle("Log10 Estradiol")
print(estr_log_hist)

#Let's knit all the plots together, before and after log transformation
pre_post_log_ph_hist <- grid.arrange(dhea_hist, testosterone_hist, estradiol_hist, 
                              dhea_log_hist, test_log_hist, estr_log_hist,
                              ncol = 3, nrow = 3)

#change wd and save plot.
setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs")
ggsave("pre_post_log_ph_hist.pdf")


```

Outlier checks 

We are defining outliers as ±5SD from mean as with a sample this large (N= ~10,000), 
we should expect some responses greater than 3SD (commonly used outlier cutoff) if the data are really normally distributed. 

Remember to run sensitivity analyses to examine the effect of outliers on models. 

```{r, outlier checks}
 
#define upper and lower limits for each hormone
outl_dhea_upper <- (mean(ph_df$dhea_log, na.rm=T) +5*(sd(ph_df$dhea_log, na.rm=T))) 
outl_dhea_lower <- (mean(ph_df$dhea_log, na.rm=T) -5*(sd(ph_df$dhea_log, na.rm=T)))
outl_test_upper <- (mean(ph_df$test_log, na.rm=T) +5*(sd(ph_df$test_log, na.rm=T))) 
outl_test_lower <- (mean(ph_df$test_log, na.rm=T) -5*(sd(ph_df$test_log, na.rm=T)))
outl_estr_upper <- (mean(ph_df$estr_log, na.rm=T) +5*(sd(ph_df$estr_log, na.rm=T))) 
outl_estr_lower <- (mean(ph_df$estr_log, na.rm=T) -5*(sd(ph_df$estr_log, na.rm=T)))

#Plot outliers:
#DHEA - NO OUTLIERS PRESENT
hist(ph_df$dhea_log, main="", xlab="DHEA LOG")
abline(v=outl_dhea_upper,col="red")
abline(v=outl_dhea_lower,col="red")

#TESTOSTERONE - OUTLIERS PRESENT
hist(ph_df$test_log, main="", xlab="TESTOSTERONE LOG")
abline(v=outl_test_upper,col="red")
abline(v=outl_test_lower,col="red")

#ESTRADIOL - OUTLIERS PRESENT 
hist(ph_df$estr_log, main="", xlab="ESTRADIOL LOG")
abline(v=outl_estr_upper,col="red")
abline(v=outl_estr_lower,col="red")

#Identify number of outliers for testostrone and estradiol. 
test_outl <- ph_df %>% 
  filter(test_log > outl_test_upper |
           test_log < outl_test_lower) #11 outliers for testosterone

estr_outl <- ph_df %>% 
  filter(estr_log > outl_estr_upper |  #10 outliers for estradiol 
           estr_log < outl_estr_lower)


#create new variable with outliers removed for later sensitivity analysis
#duplicate test_log and name it test_log_no_outl. Same for estradiol. 
ph_df <- ph_df %>% 
  mutate(test_log_no_outl = test_log) 

ph_df <- ph_df %>% 
  mutate(estr_log_no_outl = estr_log) 


#Change outlier values to NA
#tried to do with dplyr syntax below too - think it works but base R def works so stick to that for now. 
ph_df$test_log_no_outl[ph_df$test_log_no_outl > outl_test_upper |ph_df$test_log_no_outl < outl_test_lower] <- NA
ph_df$estr_log_no_outl[ph_df$estr_log_no_outl > outl_estr_upper |ph_df$estr_log_no_outl < outl_estr_lower] <- NA

#dplyr syntax
 #ph_df %>% 
  #mutate(test_log_no_outl = na_if(test_log_no_outl, test_log_no_outl > outl_test_upper |
           #test_log_no_outl < outl_test_lower)

#check it worked - it did! 
outl_test_check  <- ph_df %>% 
  filter(test_log_no_outl > outl_test_upper |
           test_log_no_outl < outl_test_lower)

outl_estr_check  <- ph_df %>% 
  filter(estr_log_no_outl > outl_estr_upper |
           estr_log_no_outl < outl_estr_lower)


```

#Generate pubertal timing score for each hormone measure

We will use the log transformed hormone levels as per Barendse & Byrne et al. (pre-print)
see script: https://github.com/dsnlab/TAG_scripts/blob/master/hormones/cleaningSaliva_concentrations_W2.Rmd

see https://boostedml.com/2019/06/linear-regression-in-r-interpreting-summarylm.html for refresher on interpretation


DHEA
```{r, DHEA PT}

#model summary
dhea_lm <- lm(dhea_log~age_years, na.action=na.exclude, data = ph_df) # create lm, specify na.exclude so df size stays the same.
summ(dhea_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 

ph_df <- ph_df %>% 
mutate(dhea_timing = resid(dhea_lm))

summary(ph_df$dhea_timing)


#Note - no sensitivity analysis to run later due to no outliers being present after Log10 transformation. 
#--------------- for sensitivity analysis with outliers removed ---------------#

```

TESTOSTERONE
```{r, TESTOSTERONE PT}

#model summary
test_lm <- lm(test_log~age_years, na.action=na.exclude, data = ph_df) # create linear model, and specify na.exclude so that df size stays the same.
summ(test_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 

ph_df <- ph_df %>% 
  mutate(test_timing = resid(test_lm))

summary(ph_df$test_timing)

#--------------- test_timing for sensitivity analysis with outliers removed ---------------#

test_lm_no_outl <- lm(test_log_no_outl~age_years, na.action=na.exclude, data = ph_df) # create linear model, and specify na.exclude so that df size stays the same.
summ(test_lm_no_outl,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 

ph_df <- ph_df %>% 
  mutate(test_timing_no_outl = resid(test_lm_no_outl))

summary(ph_df$test_timing_no_outl)


```
ESTRADIOL 

```{r, ESTRADIOL PT}

#model summary
estr_lm <- lm(estr_log~age_years, na.action=na.exclude, data = ph_df) # create linear model, and specify na.exclude so that df size stays the same.
summ(estr_lm,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 

ph_df <- ph_df %>% 
  mutate(estr_timing = resid(estr_lm))

summary(ph_df$estr_timing)

#--------------- estr_timing for sensitivity analysis with outliers removed ---------------#

estr_lm_no_outl <- lm(estr_log_no_outl~age_years, na.action=na.exclude, data = ph_df) # create linear model, and specify na.exclude so that df size stays the same.
summ(estr_lm_no_outl,
     robust = TRUE, #robust standard error = HC3 as per sandwich package
     scale = TRUE,  #scaled beta used
     confint = TRUE, digits = 3) #get confidence intervals. 95% CI by default 

#obtain standardised residual = measure of pubertal timing. Save as new variable. 

ph_df <- ph_df %>% 
  mutate(estr_timing_no_outl = resid(estr_lm_no_outl))

summary(ph_df$estr_timing_no_outl)

```
#Residual plot analysis 
Examine plots to check to validity of the assumptions (i.e., residuals vs. leverage; residuals vs. fitted; Normal Q-Q; scale-location)
This allows us to assess the quality of the regression
see https://data.library.virginia.edu/diagnostic-plots/ for guidance on plot interpretation. 
```{r, residual plot checks}

#Note sure about whether assumptions have been met. 
#Check dhea plots -  
par(mfrow=c(2,2))
plot(dhea_lm)

#Check testosterone plots - 
par(mfrow=c(2,2))
plot(test_lm)

#Check estradiol plots - Worried that the Q-Q plot diverges slightly. Look into later. 
par(mfrow=c(2,2))
plot(estr_lm)
```
# EXPORT CLEANED DATA
We now have single timing measure for each hormone plus other relevant variables. 
We will use this dataframe in our models. 

#Note: clean files should be saved in data folder in ABCD_puberty_depression project folder.

Remember total Ns: DHEA= 10,931 ; Testosterone= 11011 ; Estradiol= 5163 

```{r, export cleaning data}

abcd3.0_hormones_clean <- ph_df %>% select("src_subject_id", "eventname", "sex",
                                "dhea_timing","test_timing", "estr_timing", 
                                "test_timing_no_outl", "estr_timing_no_outl")

#Change working directory and save clean file there. 

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data")
saveRDS(abcd3.0_hormones_clean,"abcd3.0_hormones_clean.rds")
```

-------------END OF SCRIPT----------------------
