---
title: "PREP_covs_R3.0"
author: "Niamh MacSweeney"
date: "5/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The purpose of this script is to tidy and quality control the covariates for our analysis.
Covariates: Age, BMI, race/ethnicity, family income, motion, site, hemisphere (for bilateral brain measures)

This data frame was downloaded via the DEAP portal which works off the most recent ABCD data release (3.0). 

## Setup

Load libraries, set working directory.

```{r, setup}

library(psych)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggstatsplot)
library(readr)
library(gridExtra)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")
getwd()

```

##Load data and tidy variables 

Note: Renaming variables to make the names more intuitive. 

Need to insert table here of all the covariates used.... 

```{r, covariate table}

```

```{r, load and tidy data}

covs <-rio::import("/Volumes/GenScotDepression/users/niamh/ABCD_release3.0/iii.data/DEAP_data_downloads/covs_puberty_RR.rds") 

#rename variables so consistent across scripts and more intuitive
covs <- covs %>% 
  rename(eventname = event_name, sex = sex_at_birth, bmi = anthro_bmi_calc)
colnames(covs)


#Reduce to baseline data only
covs <- covs %>% 
  filter(eventname=="baseline_year_1_arm_1") #N=11,878

#change age to years instead of months 
covs <- covs %>%
  mutate(
    age_years = age/12 
  )

colnames(covs) #check new variable exists

#Let's reduce variables to those we need 
#Note: removing age in months, extra race, education (high.educ = high.educ.bl) and income vars.

covs <- covs %>% 
  select(-age, -female, -race.4level,  -high.educ.bl, - household.income)

#Let's reorder variables so that order is consistent with table above

# Age, BMI, race/ethnicity, family income, site, motion,hemisphere (for bilateral brain measures),
#parent depressive symptoms (fieldname: asr_scr_depress_r )

temp <- covs %>% 
  relocate("src_subject_id", "eventname", "sex", "age_years", "bmi", "race.6level", "demo_comb_income_p",         "abcd_site", "high.educ", "fsqc_qu_motion", "asr_scr_depress_r", "rel_family_id")

```
##Quality control covariates

We will get a summary of each covariate and check for number of NAs. These will be removed when running the linear models later.  

We will use the "rel_family_id" later to reduce sample to unrelated participants only. 

##Core Covariates

```{r, QC covs}
### Core covariates for basic model ###

#Note: Additional imaging covariates motion, hemisphere and ICV will be added later from PREP.imaging file when building models. 

#Age --- looks good, no NAs
summary(covs$age_years)

#BMI --- NAs = 94
summary(covs$bmi)

#Race/ethnicity --- NAs = 171

#Note: We are using a 6-level race measure instead of race_etnicity variable in an effort to make the measure more representative. "race.6level" includes White, Black, Asian, AIAN/NHPI (Asian Indian/Alaskan American/Native Hawaiian/Pacific Islander), Other, Mixed. 

summary(covs$race.6level)

#Study site --- why is there a site 22? 
summary(covs$abcd_site)

```

##Additional Covariates

```{r, QC covs}
### Additional Covariates ###

##Socio-economic variables

#Family income --- Refuse to answer=512, Don't know = 504, NAs= 2
summary(covs$demo_comb_income_p)

#Parent highest education --- NAs = 14.
summary(covs$high.educ)

#Parent depression --- NAs = 5. 
summary(covs$asr_scr_depress_r)

```

All covariates look within a normal range and can be exported. 






