---
title: "ANALY_demo"
author: "Niamh MacSweeney"
date: "12/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

This is a script to quantify demographic statistics for the RR, such as number of case/controls for MDD as well. 

####STEP 1: SETUP Load libraries and set wd

```{r, load libraries, set wd}

library(tidyr)
library(dplyr)
library(tidyverse)
library(psych)
library(ggplot2)
library(readr)
library(pbapply)
library(nlme)
library(gridExtra)
library(sandwich)
library(stringr)
library(kableExtra)
library(epiDisplay)
library(gmodels)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/analy")
```

#Load data 
```{r, }

#Puberty data 
pds <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/pds_timing_INT_cleaned_R3.0.rds")

hormones <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/hormones_cleaned_R3.0.rds")

#Depression data

dep <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/dep_measures_cleaned_R3.0.rds")

#Covariates
covs <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/covariates_cleaned_R3.0.rds")



```
Merge dataframes to create master dataframe

As we have a four data frames to merge, we are going to do it stepwise using dplyr function.
```{r, merge DFs}

df <- hormones %>% 
  left_join(y=pds, by= "src_subject_id") %>% 
  left_join(y=dep, by= "src_subject_id") %>% 
  left_join(y=covs, by= "src_subject_id")

#Remove duplicate columns after merge (i.e, for sex, eventname and age_years)
#This is not the most efficient way of doing thing but will do for now. 
df <- df %>% 
  select(-age_years.y, -eventname.x, -eventname.y, -sex.x, -sex.y)

#rename year variable
df <- df  %>% 
  rename(age_years = age_years.x)

#check col names
colnames(df)


```

#Get demographic details for depression measures
```{r, depression measures details}
df$sex = as.numeric(df$sex)

#select depression measures
dep_demo <- df %>% 
  select (KSADS.MDD.p, KSADS.Depressive_symptoms_ever.p, src_subject_id, sex, age_years)

#Create table 

dep_demo$sex <- 
  factor(dep_demo$sex, levels=c(1,2),
         labels=c("Female",
                  "Male"))

dep_demo$KSADS.MDD.p <-
  factor(dep_demo$KSADS.MDD.p, levels=c(0,1),
         labels=c("Control",
                  "Case"))
dep_demo$KSADS.Depressive_symptoms_ever.p <-
  factor(dep_demo$KSADS.Depressive_symptoms_ever.p, levels=c(3,2,1,0),
         labels=c("Severe",
                  "Moderate",
                  "Mild",
                  "None of the above"))

label(dep_demo$sex) <- "Sex"
label(dep_demo$age_years) <- "Age"
label(dep_demo$KSADS.Depressive_symptoms_ever.p) <- "Depression Symptom Severity"
label(dep_demo$KSADS.MDD.p) <- "MDD Status"

table1(~ sex + age_years + KSADS.Depressive_symptoms_ever.p + KSADS.MDD.p |  )
```


```{r, depression measures details}
#MDD Case Control
tab1(dep_demo$KSADS.MDD.p, cum.percent = TRUE)

#Depressive symptom severity categories
tab1(dep_demo$KSADS.Depressive_symptoms_ever.p, cum.percent = TRUE)






CrossTable(dep_demo$KSADS.MDD.p, dep_demo$KSADS.Depressive_symptoms_ever.p, 
           prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)

CrossTable(dep_demo$KSADS.MDD.p, dep_demo$sex,
           prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)

```













####Misc details for RR

1. Check N for depression measures for year two follow up. 

```{r, follow-up N for dep measures}

dep_year2 <- df %>% 

```

