---
title: "ANALY_pilot"
author: "Niamh MacSweeney"  Credit to X. Shen for function templates. 
date: "30/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#ABCD Pilot Analysis Script

#Introduction

The purpose of this script is to run the pilot analysis for the Puberty RR. For this analysis, we will be using a random sub-sample (N=1,000) to generate ROIs for the pubertal timing ~ brain structure association. 

We will be using an un-related sample of participants for our analysis. The pilot analysis will use a basic model, outlined below. 

Note that we are using inverse normal transformed data for the puberty variables (hormones and PDS).

###Required Inputs

Cleaned data Rds files are located in /Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data

####Independent Variable
Pubertal timing: pds_timing_INT_cleaned_R3.0.rds + hormones_cleaned_R3.0.rds

####Dependent Variable
Brain structure: cortical_cleaned_R3.0.rds + subcort_cleaned_R3.O.rds + dti_cleaned_R3.0.rds

####Covariates
Covariates: covariates_cleaned_R3.0.rds

____________________________________________

####Workflow 
1. Setup: Load data, merge to create master dataframe, keep unrelated participants only
2. Pilot prep: Get subsample of 10% of unrelated sample (N= 9,775). 
3. Set up lme function using Shen's template script


####STEP 1: SETUP
Load libraries and set wd
```{r, load libraries, set wd}

library(tidyr)
library(dplyr)
library(tidyverse)
library(psych)
library(ggplot2)
library(readr)
library(pbapply)
library(nlme)
library(gridExtra)
library(sandwich)
library(stringr)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/analy")
```

Load data
```{r,load data}

#Puberty data 
pds <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/pds_timing_INT_cleaned_R3.0.rds")

hormones <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/hormones_cleaned_R3.0.rds")

#Imaging data
cortical <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/cortical_cleaned_R3.0.rds")

subcort <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/subcort_cleaned_R3.0.rds")

dti <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/dti_cleaned_R3.0.rds")

#Covariates
covs <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/covariates_cleaned_R3.0.rds")

```

Merge dataframes to create master dataframe

As we have a six data frames to merge, we are going to do it stepwise using dplyr function. 
```{r, merge DFs}

df <- hormones %>% 
  left_join(y=pds, by= "src_subject_id") %>% 
  left_join(y=cortical, by= "src_subject_id") %>% 
  left_join(y=subcort, by= "src_subject_id") %>% 
  left_join(y=dti, by= "src_subject_id") %>% 
  left_join(y=covs, by="src_subject_id")

#Remove duplicate columns after merge (i.e, for sex, eventname and age_years)
#This is not the most efficient way of doing thing but will do for now. 
df <- df %>% 
  select(-age_years.y, -eventname.x, -eventname.y, -sex.x, -sex.y)

#rename year variable
df <- df  %>% 
  rename(age_years = age_years.x)

#check col names
colnames(df)

```
Check variable classifcation
```{r, variable classification}

covs_str <- df %>% 
  select("sex", "bmi", "race.6level", "age_years", "abcd_site", "fsqc_qu_motion") %>% 
  str()
```
Reduce to unrelated sample only 
Before N=11766, After N=9,775
```{r, get unrelated sample}

unrel_df =df[!duplicated(df$rel_family_id),]
```

####STEP 2: Generate subsample (10% of unrelated sample) for pilot analysis. N=978

```{r, generate subsample}

targetdata <- unrel_df %>% 
  sample_frac(size = 0.1, replace = FALSE)

#check males and females breakdown F= 482, M=496
targetdata %>% 
  dplyr::count(sex)

```

####STEP 3: Model set up 
We are using X.Shen's phewasStyle.R function to run the linear models.

We need to change "src_subject_id" to "f.eid" so that function will run. 
```{r, colnames}
colnames(targetdata)[1]='f.eid'
colnames(targetdata) #check it worked

```

Set categorical variables 
```{r, set categorical variables}

targetdata$race.6level = as.factor(targetdata$race.6level)
targetdata$sex = as.factor(targetdata$sex)
targetdata$fsqc_qu_motion = as.factor(targetdata$fsqc_qu_motion)
targetdata$abcd_site = as.factor(targetdata$abcd_site)

```

Generate long-format data.

Unsure here about what is happening when I convert to long format, it looks like there are two entries for every participant - one for imaging vars and one for non-imaging vars?
```{r, gen long-format data}
# colnames of imaging data
dat_colnames = colnames(targetdata)[grep('lh\\.|rh\\.',colnames(targetdata))] 

# colnames of non-imaging data/unilateral img data
cols_nonimg = colnames(targetdata)[!grepl('lh\\.|rh\\.',colnames(targetdata))]  
cols_nonimg = cols_nonimg[2:length(cols_nonimg)]

source('../functions/long_format_new.R')
targetdata_longformat <- long_format(targetdata,cols_nonimg,cols_img)

#Exploratory work to see what above script is doing. 
#We have two separate entries for each ID - the values are different so not sure what is happening??? Will revisit. 

test <- targetdata_longformat %>% 
  filter(f.eid ==	"NDAR_INVGY65ZZBV")

```

Define functions
```{r, Phewas function}

source('../functions/reg_phewasStyle.R')

```

Define global variables
```{r, global vars}
targetdata=targetdata
dat_long=targetdata_longformat
```

Set up dependent variables (brain measures)
```{r, set up DVs}
ls.dep.fs.long=colnames(dat_long)[grep('sa\\.|thk\\.|vol\\.|sulc\\.|dtiFA\\.|dtiMD\\.|\\.ASEG\\.',colnames(dat_long))]
ls.dep.fs.short=colnames(targetdata)[grep('^bl\\.',colnames(targetdata))]
ls.dep.all=c(ls.dep.fs.long,ls.dep.fs.short)

ls.dep.all=ls.dep.all[c(grep('sa\\.|thk\\.|vol\\.|sulc\\.',ls.dep.all),
                        grep('\\.ASEG\\.',ls.dep.all),
                        grep('dtiFA\\.',ls.dep.all),
                        grep('dtiMD\\.',ls.dep.all))]
ls.dep.all=ls.dep.all[!duplicated(ls.dep.all)]

```
Set up Independent Variables (or factors) (puberty measures) 
```{r, set up IVs}

ls.factor=colnames(targetdata)[grep('dhea_timing|test_timing|estr_timing|pt_all_INT',colnames(targetdata))]

```
Combine DVs and IVs so that we get a list of the DV and IV
```{r, combine DVs and IVs}

ls.dep.factor.combo=expand.grid(ls.dep.all,ls.factor,stringsAsFactors = F)

```
Set up Covariates 
Note: We are adding hemisphere as a covariate here... 
```{r, set up covs}

ls.models=data.frame(dependent=ls.dep.factor.combo$Var1,
                     factor=ls.dep.factor.combo$Var2,
                     covs='',stringsAsFactors = F)

ls.models$covs=paste0(c('age_years','bmi','abcd_site','race.6level','fsqc_qu_motion'),collapse='+')

ls.models$covs[!grepl('bl\\.',ls.models$dependent)]=paste0(ls.models$covs[!grepl('bl\\.',ls.models$dependent)],'+hemi')

```

#### STEP 4: Specify models

```{r, specify models}

ls.models$model.est=''
ls.models$model.est[grep('hemi',ls.models$covs)]='lme'
ls.models$model.est[ls.models$model.est=='']='glm'

```

Divide models into bulk measures and individual regions

Bulk (global) measures
```{r, bulk measures}
ls.model.bulk=ls.models[grep('mean|total|wholeb',ls.models$dependent),]
ls.model.bulk$p_batch=1
```

Individual regions
```{r, individual regions}

ls.model.region=ls.models[!grepl('mean|total|wholeb',ls.models$dependent),]

ls.model.region$p_batch=99999
target.model=ls.model.region
ls.dep.cate=c('thk\\.','sa\\.','vol\\.APARC','sulc\\.','vol\\.ASEG','dtiFA','dtiMD')
ls.factor.cate=unique(target.model$factor)
cate.no = 1
for (fac in ls.factor.cate){
      for (dep in ls.dep.cate){
            loc = grepl(dep,target.model$dependent)&grepl(fac,target.model$factor)
            target.model$p_batch[loc]=cate.no
            cate.no=cate.no+1
      }
}
ls.model.region=target.model

# individual region models add intracranial volume as a covariate
ls.model.region.covWholeB=ls.model.region
ls.model.region.covWholeB$covs[grep('thk\\.|sa\\.|vol\\.APARC|sulc\\.|vol\\.ASEG',ls.model.region.covWholeB$dependent)]=
      paste0(ls.model.region.covWholeB$covs[grep('thk\\.|sa\\.|vol\\.APARC|sulc\\.|vol\\.ASEG',ls.model.region.covWholeB$dependent)],'+ICV_ASEG')

```

#### STEP 5: Run Models 

```{r, run models}

result.PT_brain.bulk=reg_phewasStyle(ls.model.bulk,dat_short=targetdata,dat_long=dat_long,correctByFactor = T)

result.PT_brain.region.covWholeB=reg_phewasStyle(ls.model.region.covWholeB,dat_short=targetdata,dat_long=dat_long,correctByFactor = T)

save(result.PT_brain.bulk,result.PT_brain.region.covWholeB,file='../results/PT_brain_pilot_analysis.RData')

```




