---
title: "Heatmap"
author: "Nikolaj Høier" and "Niamh MacSweeney"
date: "24/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

This script is created to help produce heatmaps for ABCD studies at Div. Psych. UoE. Imaging Group led by HW. 
We will be utilizing ggseg for starts. ggseg is an amazing tool and a guide can be found  here: https://drmowinckels.io/blog/2018-08-24-introducing-the-ggseg-r-package-for-brain-segmentations/

We will use this script to visualise the ROIs for our Puberty-Depression RR. These ROIs were generated from our pilot analyses. 

#Set Up

Load in packages
```{r, ggseg}

library(tidyverse)
library(ggseg)
library(gridExtra)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs")


```
#General introduction to GGSEG - what can it help us do?

## hemisphere stacking
we can stack hemisphere if we want

```{r, stacked}
ggseg(position="stacked")
```
Now we want to colour in our segmentated area

```{r, fill}
#here we see the novelty of the underlying code, basically works like a GGplot syntax :) 
ggseg(mapping=aes(fill=region), position="stacked")


```

```{r, fill}
#we can also colour and fill but avoid showing the legend from prev. figure and add in a title :) 
ggseg(mapping=aes(fill=region), position="stacked", show.legend = F) +
  ggtitle("first atlas")
  
  # +theme_darkbrain() #can add this on if you want a dark background
```
So this atlas quite nicely helps us view cortical areas. But what about subcortical? Well there is the ASEG
subcortical atlas

```{r, ASEG/SUBCORT}
ggseg(atlas=aseg,mapping=aes(fill=region)) +
  ggtitle("Automatic subcortical segmentation")
```

#Visualise ROIs 

We are going to look at the ROIs from the RR pilot analyses.


#STEP 1: List Hypohtheses

For reference, here are our regional hypotheses:

The association between earlier pubertal timing and increased depressive symptoms (H1) will be mediated by:

•	H3a: Reduced cortical thickness in temporal regions, namely, the middle temporal gyrus and insula.

•	H3b: Reduced cortical thickness in frontal regions namely, the lateral orbito-frontal cortex and middle frontal gyrus.

•	H3c: Reduced cortical volume in temporal regions, namely, middle temporal gyrus and bank of the superior temporal sulcus.

•	H3d: Reduced cortical volume in fronto-parietal regions, namely, the middle frontal and postcentral gyri. 

•	H3e Reduced FA in the cortico-striatal tract and corpus collosum.

•	H3f: Increased sulcal depth in the pars orbitalis

•	H3g: Increased cortical volume in the ventral diencephalon. 

For completeness, H2 refers to global hypotheses:

•	H2a: Reduced global cortical volume, surface area, thickness and sulcal depth.

•	H2b: Reduced global FA

#Step 2: Load in ROI data

```{r, load data}



```







Group by cortical measurement type:

1. Cortical thickness
2. Cortical volume
3. Sulcal Depth 
4. Subcortical regions - ventral diencephalon. 




#Cortical thickness; region, beta value
PDS-Brain
Females
Insula (thickness)	-0.0663
Lateral orbitofrontal cortex (thickness)	-0.0635
Middle temporal gyrus (thickness)	-0.0708
Medial orbitofrontal cortex (thickness)	-0.0640
Rostral middle frontal gyrus (thickness)	-0.0696

Volume
Bank of the superior temporal sulcus (volume)	-0.0545
Caudal middle frontal gyrus (volume)	-0.0449
middle temporal gyrus (volume)	-0.0425

surface area
post central -0.0256

sulcal depth 
pars orbitalis  0.0393

subcort
Ventral diencephalon (volume)	0.0931


Males
Lateral orbitofrontal cortex (thickness)	-0.0525


```{r, test}


someData <- tibble(
  region = rep(c("transverse temporal", "insula",
           "precentral","superior parietal"), 2), 
  p = sample(seq(0,.5,.001), 8),
  groups = c(rep("g1", 4), rep("g2", 4))
)

someData %>%
  group_by(groups) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = p)) +
  facet_wrap(~groups)
#> merging atlas and data by 'region'

```



 We can add in our own heating to these maps:
```{R, add results}
#make dataframe with regions, Beta values and group type
thk.results = tibble(
              Region=c("insula","lateral orbitofrontal", "middle temporal", "medial orbitofrontal", "rostral middle frontal"),
              Beta=c(-0.0663, -0.0635, -0.0708, -0.0640, -0.0696),
              groups = c("thickness", "thickness", "thickness","thickness","thickness"), 
    stringsAsFactors = FALSE)


vol.results = tibble(
              Region=c("bankssts","caudal middle frontal", "middle temporal", "postcentral"),
              Beta=c(-0.0545, -0.0449, -0.0425, -0.0285),
              groups = c("volume", "volume", "volume", "volume"),
              stringsAsFactors = FALSE)

sulc.results = tibble(
                Region = "pars orbitalis",
                Beta = 0.0393,
                groups = c(rep("sulcal depth", 1)),
                stringsAsFactors = FALSE)


sa.results = tibble(
            Region = "postcentral",
            Beta = -0.0256,
            groups = c(rep("surface area", 1)),
            stringsAsFactors = FALSE)


#merge dataframes to display on one figure
df <- rbind(thk.results, vol.results, sulc.results, sa.results) 

df <- as_tibble(df) %>% 
  group_by(groups) #need to group by so that facet_wrap works
  
df %>% 
  ggseg(mapping=aes(fill= Region),
        colour="black", size=.8) +
  ggtitle("Regions of interest") + theme_brain2() +
  facet_wrap(~groups, ncol=2) +
  


df %>% 
  ggplot() +
  geom_brain(atlas=dk, 
              mapping = aes(fill = Beta),
              position = position_brain(hemi ~ side)) +
  facet_wrap(~groups, ncol=2) +
  theme(legend.position = "bottom") +
  scale_fill_gradientn(colours = c("royalblue","firebrick","goldenrod"), na.value="grey")



ggplot(someData) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = beta)) +
  scale_fill_viridis_c(option = "cividis", direction = -1) +
  theme_void() +
  labs(title = "Suicidal ideation and cortical thickness", 
       subtitle = "Associations between changes to cortical thickness and SI",
       caption = "Adjusted for sex, age, BMI, family, ID, race, anxiety and depression")




df %>% 
  ggplot() +
  geom_brain(
    atlas = dk, 
    show.legend = TRUE,
    position = position_brain(side  + hemi ~ .)
  )

  
thk.results %>% 
  ggseg(mapping=aes(fill=as.numeric(Beta)),
        colour="black", size=.8) +
  ggtitle("Cortical thickness ROIs") + theme_brain2() 
  
  
vol_plot <- vol.results %>% 
  ggseg(mapping=aes(fill=Beta),
        colour="black", size=.8, show.legend = F) +
  ggtitle("Coritcal Volume ROIs") + theme_brain2()


thk_plot_2 <- thk.results %>% 
  ggseg(mapping=aes(fill=Beta),
        colour="black", size=.8) +
  ggtitle("Cortical thickness ROIs") + theme_brain2() 
  

vol_plot_2 <- vol.results %>% 
  ggseg(mapping=aes(fill=Beta),
        colour="black", size=.8, show.legend = F) +
  ggtitle("Coritcal Volume ROIs") + theme_brain2()

plot <- grid.arrange(thk_plot,thk_plot_2, vol_plot, vol_plot_2, ncol=2, nrow=2),
                     heights=unit(c(1,10), c("in", "mm")) %>% 
  ggsave("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/ROI_heatmaps.pdf")

vol_plots <- grid.arrange(vol_plot, vol_plot_2, ncol=1, nrow=2)


position="stacked", show.legend = F




```

```{r}

# If you group_by the columns you will facet by, this will work well.
someData = someData %>% 
  group_by(AgeG)

# We can now supply the newAtlas as an atlas to ggseg
ggplot(someData) +
  geom_brain(atlas=dk, 
             colour="white",
             mapping=aes(fill=p)) +
  facet_wrap(~AgeG, ncol=1) +
  theme(legend.position = "bottom") +
  scale_fill_gradientn(colours = c("royalblue","firebrick","goldenrod"),na.value="grey")

```


