---
title: "Heatmap"
author: "Nikolaj Høier" and "Niamh MacSweeney"
date: "24/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

This script is created to help produce heatmaps for ABCD studies at Div. Psych. UoE. Imaging Group led by HW. 
We will be utilizing ggseg for starts. ggseg is an amazing tool and a guide can be found  here: https://drmowinckels.io/blog/2018-08-24-introducing-the-ggseg-r-package-for-brain-segmentations/

We will use this script to visualise the ROIs for our Puberty-Depression RR. These ROIs were generated from our pilot analyses. 

#Set Up

Load in packages
```{r, ggseg}

library(tidyverse)
library(ggseg)
library(gridExtra)
library(hrbrthemes) #nice ggplot themes
library(gcookbook)


setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs")


```
#General introduction to GGSEG - what can it help us do?

## hemisphere stacking
we can stack hemisphere if we want

```{r, stacked}
ggseg(position="stacked")
```
Now we want to colour in our segmentated area

```{r, fill}
#here we see the novelty of the underlying code, basically works like a GGplot syntax :) 
ggseg(mapping=aes(fill=region), position="stacked")


```

```{r, fill}
#we can also colour and fill but avoid showing the legend from prev. figure and add in a title :) 
ggseg(mapping=aes(fill=region), position="stacked", show.legend = F) +
  ggtitle("first atlas")
  
  # +theme_darkbrain() #can add this on if you want a dark background
```
So this atlas quite nicely helps us view cortical areas. But what about subcortical? Well there is the ASEG
subcortical atlas

```{r, ASEG/SUBCORT}
ggseg(atlas=aseg,mapping=aes(fill=region)) +
  ggtitle("Automatic subcortical segmentation")
```

#Visualise ROIs 

We are going to look at the ROIs from the RR pilot analyses.


#STEP 1: List Hypohtheses

For reference, here are our regional hypotheses:

The association between earlier pubertal timing and increased depressive symptoms (H1) will be mediated by:

•	H3a: Reduced cortical thickness in temporal regions, namely, the middle temporal gyrus and insula.

•	H3b: Reduced cortical thickness in frontal regions namely, the lateral orbito-frontal cortex and middle frontal gyrus.

•	H3c: Reduced cortical volume in temporal regions, namely, middle temporal gyrus and bank of the superior temporal sulcus.

•	H3d: Reduced cortical volume in fronto-parietal regions, namely, the middle frontal and postcentral gyri. 

•	H3e Reduced FA in the cortico-striatal tract and corpus collosum.

•	H3f: Increased sulcal depth in the pars orbitalis

•	H3g: Increased cortical volume in the ventral diencephalon. 

For completeness, H2 refers to global hypotheses:

•	H2a: Reduced global cortical volume, surface area, thickness and sulcal depth.

•	H2b: Reduced global FA

#Step 2: Load in ROI data (for both puberty and depression associations)


Group by cortical measurement type:

1. Cortical thickness
2. Cortical volume
3. Sulcal Depth 
4. Subcortical regions - ventral diencephalon. 


#Cortical thickness; region, beta value
PDS-Brain
Females
Insula (thickness)	-0.0663
Lateral orbitofrontal cortex (thickness)	-0.0635
Middle temporal gyrus (thickness)	-0.0708
Medial orbitofrontal cortex (thickness)	-0.0640
Rostral middle frontal gyrus (thickness)	-0.0696

Volume
Bank of the superior temporal sulcus (volume)	-0.0545
Caudal middle frontal gyrus (volume)	-0.0449
middle temporal gyrus (volume)	-0.0425

surface area
post central -0.0256

sulcal depth 
pars orbitalis  0.0393

subcort
Ventral diencephalon (volume)	0.0931

Males
Lateral orbitofrontal cortex (thickness)	-0.0525

White matter - ggseg doesn't have atlas to match the DTI parcellation in ABCD. 

Lower FA in cortico-striate tract (beta = -0.0327) and corpus callosum (beta = -0.0280)

 We can add in our own heating to these maps:
 
```{R, add results}
#make dataframe with regions, Beta values and group type
thk.results = tibble(
              region=c("insula","lateral orbitofrontal", "middle temporal", "medial orbitofrontal", "rostral middle frontal"),
              beta=c(-0.0663, -0.0635, -0.0708, -0.0640, -0.0696),
              group = c("thickness", "thickness", "thickness","thickness","thickness"))
    #stringsAsFactors = FALSE


vol.results = tibble(
              region=c("bankssts","caudal middle frontal", "middle temporal", "postcentral"),
              beta=c(-0.0545, -0.0449, -0.0425, -0.0285),
              group = c("volume", "volume", "volume", "volume"))
              #stringsAsFactors = FALSE)

sulc.results = tibble(
                region = "pars orbitalis",
                beta = 0.0393,
                group = "sulcal depth")
                #stringsAsFactors = FALSE)


sa.results = tibble(
            region = "postcentral",
            beta = -0.0256,
            group = "surface area")
            #stringsAsFactors = FALSE)


#merge dataframes to display on one figure
df <- rbind(thk.results, vol.results, sulc.results, sa.results)
```



```{r, plot cortical}

###Cortical ROIs #####
#via ggseg
df %>% 
  group_by(group) %>% 
  ggseg(atlas = dk, 
      colour = "black",
      size = .2, 
      hemi = "right",
      position = "stacked",
      mapping = aes(fill = beta)) +
  scale_fill_gradient(low="deepskyblue",high="midnightblue") +
  labs(title = "Pilot Analysis — Cortical Regions of Interest") +
  theme_custombrain(plot.background = "white",text.colour = "black", text.size = 12, text.family = "mono") +
  facet_wrap(~group)
ggsave("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/cortical_ROI_heatmap.png")



## via geom_brain
df %>%
  group_by(group) %>%
  ggplot() +
  geom_brain(atlas = dk,colour="white",
             position = position_brain(hemi ~ side),
             aes(fill = beta)) +
  theme_brain() +
  scale_fill_gradient(low="deepskyblue",high="midnightblue") +
  labs(title = "Pilot Analysis — Regions of Interest") +
  facet_wrap(~group)

```


```{r, subcortical}

#make df for subcortical ROI

subcort_df  = tibble(
                region = "ventral DC",
                beta = 	0.0931)
subcort_df %>% 
  ggseg(atlas = aseg, 
        side = "sagittal",
      colour = "black",
      size = .1, 
      mapping = aes(fill = beta)) +
  labs(title = "Pilot Analysis — Subcortical Regions of Interest") +
  theme_custombrain(plot.background = "white",text.colour = "black", text.size = 12, text.family = "mono")
ggsave("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/subcortical_ROI_heatmap.png")

  
```


``` {r, individual plots}

  
vol_plot <- vol.results %>% 
  ggseg(mapping=aes(fill=Beta),
        colour="black", size=.8, show.legend = F) +
  ggtitle("Coritcal Volume ROIs") + theme_brain2()


thk_plot_2 <- thk.results %>% 
  ggseg(mapping=aes(fill=Beta),
        colour="black", size=.8) +
  ggtitle("Cortical thickness ROIs") + theme_brain2() 
  

vol_plot_2 <- vol.results %>% 
  ggseg(mapping=aes(fill=Beta),
        colour="black", size=.8, show.legend = F) +
  ggtitle("Coritcal Volume ROIs") + theme_brain2()

plot <- grid.arrange(thk_plot,thk_plot_2, vol_plot, vol_plot_2, ncol=2, nrow=2),
                     heights=unit(c(1,10), c("in", "mm")) %>% 
  ggsave("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/ROI_heatmaps.pdf")

vol_plots <- grid.arrange(vol_plot, vol_plot_2, ncol=1, nrow=2)


```

#####Puberty Brain Associations ROIs

Global cortical thickness	-0.0659
Global sulcal depth	-0.0540

Insula (thickness)	-0.0663
Lateral orbitofrontal cortex (thickness)	-0.0635
Middle temporal gyrus (thickness)	-0.0708
Medial orbitofrontal cortex (thickness)	-0.0640
Rostral middle frontal gyrus (thickness)	-0.0696
Bank of the superior temporal sulcus (volume)	-0.0545
Caudal middle frontal gyrus (volume)	-0.0449
middle temporal gyrus (volume)	-0.0425
Lateral orbitofrontal cortex (thickness)	-0.0525

Ventral diencephalon (volume)	0.0931
	

```{R, add results}
#make dataframe with regions, Beta values and group type
pub_thk.results = tibble(
              region=c("insula","lateral orbitofrontal", "middle temporal", "medial orbitofrontal", "rostral middle frontal"),
              beta=c(-0.0663, -0.0635, -0.0708, -0.0640, -0.0696),
              group = c("thickness", "thickness", "thickness","thickness","thickness"))
    #stringsAsFactors = FALSE


pub_vol.results = tibble(
              region=c("bankssts","caudal middle frontal", "middle temporal"),
              beta=c(-0.0545, -0.0449, -0.0425),
              group = c("volume", "volume", "volume"))
              #stringsAsFactors = FALSE)


#merge dataframes to display on one figure
pub_df <- rbind(pub_thk.results, pub_vol.results)
```

Plot
```{r, plot puberty ROIs}

###Cortical ROIs #####
#via ggseg
pub_df %>% 
  group_by(group) %>% 
  ggseg(atlas = dk, 
      colour = "black",
      size = .2, 
      hemi = "right",
      position = "stacked",
      mapping = aes(fill = beta)) +
  scale_fill_gradient(low="deepskyblue",high="midnightblue") +
  labs(title = "Pubertal Timing — Cortical Regions of Interest") +
  theme_custombrain(plot.background = "white",text.colour = "black", text.size = 10, text.family = "mono") +
  theme(plot.margin=unit(c(2,2,2,3),'mm'),
        plot.title = element_text(lineheight=1, face="bold", vjust=1, hjust=0.5,size=12)) +
  facet_wrap(~group)
ggsave("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/PT_cortical_ROI_heatmap.png")

```

Only one subcortical ROI for puberty-brain associations. 
```{r, subcortical}

#make df for subcortical ROI

subcort_df  = tibble(
                region = "ventral DC",
                beta = 	0.0931)
subcort_df %>% 
  ggseg(atlas = aseg, 
        side = "sagittal",
      colour = "black",
      size = .1, 
      mapping = aes(fill = beta)) +
  labs(title = "Pubertal Timing — Subcortical Regions of Interest") +
  theme_custombrain(plot.background = "white",text.colour = "black", text.size = 10, text.family = "mono") +
  theme(plot.margin=unit(c(2,2,2,3),'mm'),
        plot.title = element_text(lineheight=1, face="bold", vjust=1, hjust=0.5,size=12)) 
ggsave("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/PT_subcortical_ROI_heatmap.png")


```

####Depression - Brain Measures (Cortical/subcortical)

Postcentral gyrus (surface area)	-0.0256
Middle temporal gyrus (volume)	-0.0283
Postcentral gyrus (volume)	-0.0285
Pars orbitalis
(Sulcal depth)	0.0393


```{R, add results}
#make dataframe with regions, Beta values and group type


dep_vol.results = tibble(
              region=c("postcentral", "middle temporal"),
              beta=c(-0.0285, -0.0425),
              group = c("volume", "volume"))
            


dep_sulc.results = tibble(
                region = "pars orbitalis",
                beta = 0.0393,
                group = "sulcal depth")
                #stringsAsFactors = FALSE)


dep_sa.results = tibble(
            region = "postcentral",
            beta = -0.0256,
            group = "surface area")
            #stringsAsFactors = FALSE)


#merge dataframes to display on one figure
dep_df <- rbind(dep_vol.results, dep_sulc.results, dep_sa.results)

```

Plot Cortical Measures

```{r, plot depression ROIs}

###Cortical ROIs #####
#via ggseg
dep_df %>% 
  group_by(group) %>% 
  ggseg(atlas = dk, 
      colour = "black",
      size = .2, 
      hemi = "right",
      position = "stacked",
      mapping = aes(fill = beta)) +
  scale_fill_gradient(low="deepskyblue",high="midnightblue") +
  labs(title = "Depressive Symptoms — Cortical Regions of Interest") +
  theme_custombrain(plot.background = "white",text.colour = "black", text.size = 10, text.family = "mono") +
  theme(plot.margin=unit(c(2,2,2,3),'mm'),
        plot.title = element_text(lineheight=1, face="bold", vjust=1, hjust=0.5,size=12)) +
  facet_wrap(~group, ncol = 3)
ggsave("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/dep_cortical_ROI_heatmap.png")

```

Plot Global Measures

Puberty
Females
Global cortical thickness	-0.0659
Global sulcal depth	-0.0540


Depression
Females
Global surface area	-0.0396
Global cortical volume	-0.0393
Global FA	-0.0290

Males
Global cortical surface area	-0.0607
Global cortical volume	-0.0636

For plotting purposes, get average for values that are shared between males and females
Surface area: -.0501
Cortical volume: -0.0513


```{r, global data}

global_results <- data.frame(
                measure = c("volume", "surface area", "thickness", "sulcal depth","FA"),
                beta = 	c(-.0513, -.0501, -.0959, -.0540 , -0.029),
                group = c("Depressive Symptoms", "Depressive Symptoms", "Pubertal Timing", "Pubertal Timing", "Depressive Symptoms"))

```



Make bar graph 
```{r}

global_fig=
  ggplot(global_results, aes(x=measure, y=beta)) + 
  geom_bar(fill='deepskyblue4', position=position_dodge(), stat="identity", width=0.5) +
  ggtitle("Global Results")+
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(hjust=0.5),
    axis.line.x = element_line(size=0.3),
    axis.text=element_text(size=10, hjust=-0.01), axis.title=element_text(size=10),
    plot.title = element_text(lineheight=1, face="bold", vjust=1, hjust=0.5,size=14),
    strip.text = element_text(size=10),
    panel.spacing = unit(2, "lines"),
    plot.margin=unit(c(2,2,2,3),'mm')) +
  # geom_text(aes(label=sig), colour="black", hjust=12, size=3)+
  # ylab("Beta") + xlab("\n\n") +
  # ylim(c(-0.07,0.06))+
  #scale_y_reverse()+
  scale_x_discrete(position='top')+
  geom_hline(yintercept=0,size=0.5,color='white')+
  geom_vline(xintercept=2.5,size=0.5,color='white', linetype = "dashed")+
  coord_flip() +
      facet_wrap(~group) 
print(global_fig)
ggsave("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/global_results_barplot.png")


```




