---
title: "ANALY_pilot_dep_brain_males"
author: "Niamh MacSweeney"  Credit to X. Shen for function templates. 
date: "20/04/2022"
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
#ABCD Pilot Analysis Script - Pubertal timing ~ Brain Structure -> Males

####Introduction

The purpose of this script is to run the pilot analysis (PT~BS) for the Puberty RR. For this analysis, we will be using a random sub-sample of 10% (N= ~1,000) to generate ROIs for the pubertal timing ~ brain structure association.

This sub-sample will be split into males and females and the script will be run independently for each.

Note that this sample includes unrelated individuals. 

Please note that in this script, brain structure is set as the dependent variable due to the treatment of hemisphere mixed effects. However, in the RR, our hypotheses state that depression is the dependent or outcome variable. As this analysis is cross-sectional, the Y and X terms are statistically interchangeable so Y ~ X = pubertal timing ~ brain = brain ~ pubertal timing. For our later mediation models, the order of the terms is important so depression will be the dependent or outcome variable. See RR manuscript for further details on mediation model set up.

Required Inputs

Cleaned data Rds files are located in /Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data

Independent Variable — Pubertal timing: pds_timing_baseline_R4.0.rds

Dependent Variable — Brain structure: cortical_baseline_R4.0.rds + subcort_baseline_R4.O.rds + dti_baseline_R4.0.rds

Covariates — Covariates: covs_pilot_R4.0.rds

------------------------------------------------------------------------

####Workflow

1.  Setup: Load data, merge to create master dataframe, keep unrelated participants only
2.  Pilot prep: Get subsample of 10% of sample (N= ~6,000). Split into males and females.
3.  Set up glm/lme function

#####STEP 1: SETUP Load libraries and set wd

```{r, load libraries, set wd}


library(tidyverse)
library(lme4)
library(lmerTest)
library(nlme)
library(psych)
library(ggplot2)
library(readr)
library(pbapply)
library(gridExtra)
library(sandwich)
library(stringr)
library(kableExtra)
#library(epiDisplay)
library(gmodels)
library(caret)


setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/analy")
```

Load data:

```{r,load data}

#Pubertal timing data 
pds <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/pds_timing_baseline_R4.0.rds")

#Imaging data
cortical <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/cortical_baseline_R4.0.rds")

subcort <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/subcort_baseline_R4.0.rds")

dti <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/dti_baseline_R4.0.rds")

#Covariates
covs <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/covs_pilot_R4.0.rds")

```

Merge dataframes to create master dataframe

As we have a five data frames to merge, we are going to do it stepwise using dplyr function.

```{r, merge DFs}

#before merge, remove extra variables in pds

pds <- pds %>%
  select(-c("eventname", "sex", "age_years"))

df <- pds %>% 
  left_join(y=cortical, by= "src_subject_id") %>% 
  left_join(y=subcort, by= "src_subject_id") %>% 
  left_join(y=dti, by= "src_subject_id") %>% 
  left_join(y=covs, by="src_subject_id")

#check col names
colnames(df)

```

Check variable classifcation

```{r, variable classification}

df %>% 
 select("rel_family_id", "race.6level", "age_years", "scanner_id", "dmri_dti_meanmotion") %>% 
  str()
```


Get sample demographics: Age = 9.91 years; SD = 0.62

```{r, sample demographics}

df %>% 
  dplyr::select(age_years, sex) %>% 
  summary()

#get sd
sd(df$age_years)
#get mean
mean(df$age_years)

#make df with males only for demographic table purposes
males <- df %>% #N= 6193
    filter(sex == "M")
  
```

IGNORE: Inspect Ns per nested condition
```{r, inspect Ns per condition}
# 
# test <- females %>% 
#   group_by(rel_family_id, scanner_id) %>% 
#   count()

```


#####STEP 2: Generate subsample (10%/20% of unrelated sample) for pilot analysis. N=1188 Get males only

Note: name of object needs to be targetdata for function

1. Get unrelated sample N= 9850
```{r, get unrelated sample using randomisation approach}

#Assign random number to each family ID. Set random 

df$random <- NA #generate new column of NAs
unrel_df <- data.frame() #make new df

#iterate through family IDs using for loop (this will take a few minutes)
# set.seed(2507) #to make sure random sample is the same random sample each time
unique_fam_id <- unique(df$rel_family_id)
for (i in unique_fam_id[1:length(unique_fam_id)]){
  
  set.seed(2507)
  
  #subset each family and assign random number to each individual within each family
  df$random[df$rel_family_id == i] <- sample(nrow(df[df$rel_family_id == i,]),nrow(df[df$rel_family_id == i,]), replace= FALSE)
  
  #select individual with highest number for unrelated dataframe
  unrel_individual <- df[df$rel_family_id == i,][which.max(df$random[df$rel_family_id == i]),]
    
  #append unrelated individual to unrelated dataframe
  unrel_df <- rbind(unrel_df, unrel_individual)
  
}

# i <- 8781 #for trouble shooting. 
nrow(unrel_df) #check that N= 9146
length(unique(unrel_df$rel_family_id)) #check that no. of unique IDs is correct. 

#Check that set.seed worked so that we get the same IDs each timescript is run

test_set_seed <- unrel_df$src_subject_id #make new vector of un_rel individual ids

#the rerun df$random lines + for loop, without clearing global environment
#check if new participant id list matches test_set_seed

all(unrel_df$src_subject_id == test_set_seed) #it worked! 

```

```{r, generate subsample, get males only}

set.seed(2507) #set seed for reproducibility
targetdata <- df %>% 
  sample_frac(size = 0.1, replace = FALSE)

#check males and females breakdown F= 564, M=623
targetdata %>% 
  dplyr::count(sex)

#get mean age and sd for pilot sample. Mean age = 9.91, SD = 0.62
mean(targetdata$age_years)
sd(targetdata$age_years)

#Get Female sample only 
targetdata <- targetdata %>% 
  filter(sex == "M")

```

Make demographics table for CBCL breakdown

```{r, CBCL demographics table}

#DS category breakdown
CrossTable(targetdata$cbcl_scr_syn_withdep_r, 
           prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)

```

Remove columns not needed for pilot analyis
```{r, remove cols not needed for pilot analysis}
# targetdata <- targetdata %>% 
#   select(-c("bmi", "household_income", "asr_scr_depress_r", "interview_age", "site_id_l", "dmri_dti_meanrot", "dmri_dti_meantrans", "acs_raked_propensity_score" ,"mri_info_visitid", "mri_info_studydate", "rel_group_id", "random"))
```

Check for multicolinearity

```{r, multicolinearity}

#only include numeric variables
mydata <- targetdata %>% 
  dplyr::select(where(is.numeric))

#check colinearity with dti variables only 

dti_corr <- targetdata %>% 
  dplyr::select(matches('age_years|race.6level|scanner_id|dti'))

dti_corr <- dti_corr %>% 
  dplyr::select(-(contains("allfibers")))

#numeric only
dti_corr <- dti_corr %>% 
  dplyr::select(where(is.numeric))
#make correlation matrix
dti_corr_mat = cor(dti_corr, method = "pearson", use = "pairwise.complete.obs")

#get column names that have a correlation >0.95
tmp <- dti_corr_mat > 0.7
diag(tmp) <- FALSE
names(Filter(function(x) x > 0, rowSums(tmp) > 0 | colSums(tmp) > 0))


#it seems that the superior longitudinal fasiculus and the superior cortical striatal tracts are collinear. 
```


\#\#\#\#STEP 3: Model set up We are using X.Shen's reg_phewasStyle_withCI.R function to run the glm and lme for unilateral and bilateral brain regions respectively.
We need to change "src_subject_id" to "f.eid" so that function will run.

```{r, colnames}
colnames(targetdata)[1]='f.eid'
colnames(targetdata) #check it worked

#check for duplicate columns
unique_names <- unique(colnames(targetdata)) #no duplicates 

```

Define Basic Model for pilot analysis

IV: Pubertal timing

DV: Brain Measures

Covariates: Fixed = age, ethnicity; scanner ID, dti mean motion for (DTI models), hemisphere for bilateral (LME) models

Random: ID for LME models. 


Set categorical variables

```{r, set categorical variables}

targetdata$race.6level = as.factor(targetdata$race.6level)
targetdata$scanner_id = as.factor(targetdata$scanner_id)

```
Standardise variables before running models to avoid scaling issues. Note, we will only scale the numeric variables

```{r, scale variables}

targetdata <- targetdata %>% 
mutate_if(is.numeric, scale)

```

Generate long-format data.

Note: for the lme models (bilateral brain regions only), the data needs to be in long-format.  Hemisphere is controlled for as a fixed, within-participant effect.

```{r, gen long-format data}
# colnames of imaging data
dat_colnames = colnames(targetdata)[grep('lh\\.|rh\\.',colnames(targetdata))] 

# colnames of non-imaging data/unilateral img data
cols_nonimg = colnames(targetdata)[!grepl('lh\\.|rh\\.',colnames(targetdata))]  
cols_nonimg = cols_nonimg[2:length(cols_nonimg)]

source('../functions/long_format_new.R')
targetdata_longformat <- long_format(targetdata,cols_nonimg,cols_img)

```

Define functions

Note: Make sure that the "family" argument in function is correct (i.e., we want it to be "poisson" not "binomial")


```{r, Phewas function}

source('../functions/reg_phewasStyle_withCI.R')

```

Define global variables

```{r, global vars}
targetdata=targetdata
dat_long=targetdata_longformat

```

Set up dependent variables (brain measures: cortical(APARC files); subcortical(ASEG files); whitematter(dti)

```{r, set up DVs}

ls.dep.fs.long=colnames(dat_long)[grep('sa\\.|thk\\.|vol\\.|sulc\\.|\\.ASEG\\.|dtifa\\_|dtimd\\_',colnames(dat_long))] #select vars of interest
ls.dep.fs.long=colnames(dat_long)[-grep('avg\\.',colnames(dat_long))] #remove average DTI measures 

ls.dep.fs.short=colnames(targetdata)[grep('^bl\\.|\\_cc|\\_fmaj|\\_fmin',colnames(targetdata))]
ls.dep.all=c(ls.dep.fs.long,ls.dep.fs.short)

ls.dep.all=ls.dep.all[c(grep('sa\\.|thk\\.|vol\\.|sulc\\.',ls.dep.all),
                        grep('\\.ASEG\\.',ls.dep.all),
                        grep('dtifa\\_',ls.dep.all),
                        grep('dtimd\\_',ls.dep.all))]
ls.dep.all=ls.dep.all[!duplicated(ls.dep.all)]

```

Set up Independent Variable (or factor) (Pubertal timing for males)


```{r, set up IVs}

ls.factor=colnames(targetdata)[grep('^pt_m',
                                    colnames(targetdata))]

```

Combine DVs and IVs so that we get a list of the DV and IV

```{r, combine DVs and IVs}

ls.dep.factor.combo=expand.grid(ls.dep.all,ls.factor,stringsAsFactors = F)

```

Set up Covariates Note: We are adding hemisphere as a covariate here...

```{r, set up covs}

ls.models=data.frame(dependent=ls.dep.factor.combo$Var1,
                     factor=ls.dep.factor.combo$Var2,
                     covs='',stringsAsFactors = F)

ls.models$covs=paste0(c('age_years','race.6level', 'scanner_id'),collapse='+')

#For all models that are not bilateral, add hemisphere as a covariate. 
#Note that these DTI measures: CC, fmaj and fmin are unilateral, so these need to be listed here too. 
ls.models$covs[!grepl('bl\\.|\\_cc|\\_fmaj|\\_fmin',ls.models$dependent)]=paste0(ls.models$covs[!grepl('bl\\.|\\_cc|\\_fmaj|\\_fmin',ls.models$dependent)],'+hemi')

#add dti mean motion (fd) as cov for dti models
ls.models$covs[grepl('dtifa\\_|dtimd\\_',ls.models$dependent)]=paste0(ls.models$covs[grepl('dtifa\\_|dtimd\\_',ls.models$dependent)],'+dmri_dti_meanmotion')

```

#### STEP 4: Specify models

```{r, specify models}

ls.models$model.est=''
ls.models$model.est[grep('hemi',ls.models$covs)]='lme'
ls.models$model.est[ls.models$model.est=='']='glm'


```

Divide models into bulk measures and individual regions

Bulk (global) measures

```{r, bulk measures}
ls.model.bulk=ls.models[grep('mean|total|allfibers',ls.models$dependent),]
ls.model.bulk$p_batch=1

```

Individual regions

```{r, individual regions}

ls.model.region=ls.models[!grepl('mean|total|allfibers',ls.models$dependent),] 

ls.model.region$p_batch=99999
target.model=ls.model.region
ls.dep.cate=c('thk\\.','sa\\.','vol\\.APARC','sulc\\.','vol\\.ASEG','dtifa\\_','dtimd\\_')
ls.factor.cate=unique(target.model$factor)
cate.no = 1
for (fac in ls.factor.cate){
      for (dep in ls.dep.cate){
            loc = grepl(dep,target.model$dependent)&grepl(fac,target.model$factor)
            target.model$p_batch[loc]=cate.no
            cate.no=cate.no+1
      }
}
ls.model.region=target.model

# individual region models add whole brain volume as a covariate
ls.model.region.covWholeB=ls.model.region
ls.model.region.covWholeB$covs[grep('thk\\.|sa\\.|vol\\.APARC|sulc\\.|vol\\.ASEG',ls.model.region.covWholeB$dependent)]=
      paste0(ls.model.region.covWholeB$covs[grep('thk\\.|sa\\.|vol\\.APARC|sulc\\.|vol\\.ASEG',ls.model.region.covWholeB$dependent)],'+WBV')


```

#### STEP 5: Run Models

```{r, run models}

#Global (bulk) models
result.males.pt_brain.bulk=reg_phewasStyle(ls.model.bulk,dat_short=targetdata,dat_long=dat_long,correctByFactor = T)

#Regional (Individual) models
result.males.pt_brain.region.covWholeB=reg_phewasStyle(ls.model.region.covWholeB,dat_short=targetdata,dat_long=dat_long,correctByFactor = T)

#Save as R.Data file
save(result.males.pt_brain.bulk,result.males.pt_brain.region.covWholeB,file='../results/revised_pilot/males_pt_brain_pilot.RData')

#add TRUE column if nominal p value < 0.005 for pilot ROIs

#global results
result.males.pt_brain.bulk$Sig <- as.numeric((as.character(result.males.pt_brain.bulk[,"p.value"]))) < 0.01

#regional results
result.males.pt_brain.region.covWholeB$Sig <- as.numeric((as.character(result.males.pt_brain.region.covWholeB[,"p.value"]))) < 0.01

#Save results as .csv files

#global results
write.csv(result.males.pt_brain.bulk, file = '../results/revised_pilot/tables/pilot_males.pt_brain.bulk.csv')

#regional results 
write.csv(result.males.pt_brain.region.covWholeB, file = '../results/revised_pilot/tables/pilot_males_pt_brain.regional.csv')

```

STEP 6: Identify ROIs for confirmatory analysis. Threshold: Nominal p-value \< 0.01

```{r, make tables}

ROI_pt_brain_males_regional_results <- result.males.pt_brain.region.covWholeB %>%
  filter(p.value <0.01)

#save as R.data
save(ROI_pt_brain_males_regional_results,file='../results/revised_pilot/ROIs_pt_brain_males.RData')

#save as .csv
write.csv(ROI_pt_brain_males_regional_results,file= "../results/revised_pilot/tables/ROIs_pt_brain_males.csv")

#save as Kable Table

# ROI_table <- ROI_dep_brain_females_regional_results %>%
#        kable(digits=4, format = "html",
#       caption = "Females: Depressive symptoms-brain structure models and associated statistics with significant ROI associations") %>%
#                kable_styling(font_size = 15,
#                              position = "left",
#                              full_width = F) #%>%
#       #save_kable(file = "../figs/female_ROIs_PT_brain_pilot_analysis.png")
# 
# str(ROI_dep_brain_females_regional_results) #check variable chars


```

# \#\#\#\#\#\#\#\#\#\#\#\#\#\# End of Script \#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#

Make demographic table for paper - where we have a breakdown of the MDD and DS categories for each sex

```{r, demographic table}


```
