---
title: "ANALY_pilot_dep_brain_females"
author: "Niamh MacSweeney"  Credit to X. Shen for function templates. 
date: "08/9/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\#ABCD Pilot Analysis Script - Depression ~ Brain Structure -> Females

\#Introduction

The purpose of this script is to run the pilot analysis (DR~BS) for the Puberty RR. For this analysis, we will be using a random sub-sample of 10% (N= ~1,000) to generate ROIs for the depression ~ brain structure association.

This sub-sample will be split into males and females and the script will be run independently for each.

We will be using an un-related sample of participants for our analysis. The pilot analysis will use a basic model, outlined below.

\#\#\#Required Inputs

Cleaned data Rds files are located in /Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data

\#\#\#\#Independent Variable Depression Rating (Parent report): dep_measures_cleaned_R3.0.rds

\#\#\#\#Dependent Variable Brain structure: cortical_cleaned_R3.0.rds + subcort_cleaned_R3.O.rds + dti_cleaned_R3.0.rds

\#\#\#\#Covariates Covariates: covariates_cleaned_R3.0.rds

------------------------------------------------------------------------

####Workflow 

1. Setup: Load data, merge to create master dataframe, keep unrelated participants only 
2. Pilot prep: Get subsample of 10% of unrelated sample (N= 9,775). Split into males and females. 
3. Set up glm/lme function using Shen's template script

####STEP 1: SETUP Load libraries and set wd

```{r, load libraries, set wd}

library(tidyr)
library(dplyr)
library(tidyverse)
library(psych)
library(ggplot2)
library(readr)
library(pbapply)
library(nlme)
library(gridExtra)
library(sandwich)
library(stringr)
library(kableExtra)
library(epiDisplay)
library(gmodels)
library(plyr)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/analy")
```

Load data:

Note on depression variables:

As noted in Mental Health release notes for 4.0, the ABCD Team spotted inaccuracies in the algorithm used to derive the MDD diagnosis. For this reason, we have decided to not use the MDD diagnosis as a depression rating measure in the pilot analysis. Instead, we will use the ordinal measure of DS only (KSADS.Depressive_symptoms_ever).

The following measures are all available in the dep_measures_cleaned.rds file but we will only focus on the KSADS.Depressive_symptoms_ever.p for the pilot analysis. 

- KSADS.MDD.y: MDD diagnosis based on child report
- KSADS.MDD.p: MDD diagnosis based on caregiver report
- KSADS.Dep_symptoms_sum_ever.y: no. of depressive symptoms reported based on child report
- KSADS.Dep_symptoms_sum_ever.p: no. of depressive symptoms reported based on caregiver report
- KSADS.Depressive_symptoms_ever.y: DSM-V depressive symptoms reported by children
- KSADS.Depressive_symptoms_ever.p: DSM-V depressive symptoms reported by caregivers

```{r,load data}

#Depression data 
dep <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/dep_measures_cleaned_R3.0.rds")

#Imaging data
cortical <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/cortical_cleaned_R3.0.rds")

subcort <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/subcort_cleaned_R3.0.rds")

dti <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/dti_cleaned_R3.0.rds")

#Covariates
covs <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/covariates_cleaned_R3.0.rds")

```

Merge dataframes to create master dataframe

As we have a five data frames to merge, we are going to do it stepwise using dplyr function.

```{r, merge DFs}

df <- dep %>% 
  left_join(y=cortical, by= "src_subject_id") %>% 
  left_join(y=subcort, by= "src_subject_id") %>% 
  left_join(y=dti, by= "src_subject_id") %>% 
  left_join(y=covs, by="src_subject_id")

#check col names
colnames(df)

```

Check variable classifcation

```{r, variable classification}

#covs_str <- df %>% 
#  select("sex", "bmi", "race.6level", "age_years", "abcd_site", "fsqc_qu_motion") %>% 
#  str()
```

Reduce to unrelated sample only Before N=11878, After N=9,856

We are going to randomise which family member is included in the study 

```{r, get unrelated sample using randomisation approach}

#Assign random number to each family ID. Set random 

df$random <- NA #generate new column of NAs
unrel_df <- data.frame() #make new df

#iterate through family IDs using for loop (this will take a few minutes)
set.seed(2507) #to make sure random sample is the same random sample each time
for (i in unique(df$rel_family_id)){
  
  #subset each family and assign random number to each individual within each family
  df$random[df$rel_family_id == i] <- sample(nrow(df[df$rel_family_id == i,]),nrow(df[df$rel_family_id == i,]), replace= FALSE)
  
  #select individual with highest number for unrelated dataframe
  unrel_individual <- df[which.max(df$random[df$rel_family_id == i]),] 
  #append unrelated individual to unrelated dataframe
  unrel_df <- rbind(unrel_df, unrel_individual)
  
}

nrow(unrel_df) #check that N= 9856

```

Get sample demographics 
```{r, sample demographics}

unrel_df %>% 
  dplyr::select(age_years, sex) %>% 
  summary()
#get sd
sd(unrel_df$age_years)
#get mean
mean(unrel_df$age_years)

#make df with unrelated females only for demographic table purposes
unrel_females <- unrel_df %>% #N= 4673
    filter(sex == "F")
  

```

####STEP 2: Generate subsample (10% of unrelated sample) for pilot analysis. N=986 Get females only

```{r, generate subsample, get females only}

set.seed(2507) #set seed for reproducibility
targetdata <- unrel_df %>% 
  sample_frac(size = 0.1, replace = FALSE)

#check males and females breakdown F= 475, M=511
targetdata %>% 
  dplyr::count(sex)

#get mean age and sd for pilot sample
mean(targetdata$age_years)
sd(targetdata$age_years)

#Get Female sample only 
targetdata <- targetdata %>% 
  filter(sex == "F")

```
Make demographics table for MDD/DS breakdown 
```{r, MDD/DS demographics table}

####TOTAL FEMALE UNRELATED SAMPLE

#no longer using MDD binary measure in this project due to error in MDD algorithm spotted in 4.0 release notes.

# #MDD case control
# CrossTable(unrel_females$KSADS.MDD.p, 
#            prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)
# 
# #MDD and age breakdown
# ddply(unrel_females,~KSADS.MDD.p,summarise,mean=mean(age_years),sd=sd(age_years))

#DS category breakdown
CrossTable(unrel_females$KSADS.Depressive_symptoms_ever.p, 
           prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)

#DS and age breakdown
ddply(unrel_females,~KSADS.Depressive_symptoms_ever.p,summarise,mean=mean(age_years),sd=sd(age_years))

#### PILOT ANALYSIS SAMPLE
# #MDD case control
# CrossTable(targetdata$KSADS.MDD.p, 
#            prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)

#MDD and age breakdown
# ddply(targetdata,~KSADS.MDD.p,summarise,mean=mean(age_years),sd=sd(age_years))

#DS category breakdown
CrossTable(targetdata$KSADS.Depressive_symptoms_ever.p, 
           prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)

#DS and age breakdown
ddply(targetdata,~KSADS.Depressive_symptoms_ever.p,summarise,mean=mean(age_years),sd=sd(age_years))


```

####STEP 3: Model set up We are using X.Shen's phewasStyle_withCI.R function to run the linear models.

We need to change "src_subject_id" to "f.eid" so that function will run.

```{r, colnames}
colnames(targetdata)[1]='f.eid'
colnames(targetdata) #check it worked

```

Set categorical variables

```{r, set categorical variables}

targetdata$race.6level = as.factor(targetdata$race.6level)
targetdata$sex = as.factor(targetdata$sex)
targetdata$fsqc_qu_motion = as.factor(targetdata$fsqc_qu_motion)
targetdata$abcd_site = as.factor(targetdata$abcd_site)

```

Generate long-format data.

Note: for the lme models (bilaterlal brain regions only), the data needs to be in long-format (i.e, there is a row per hemisphere (lh and rh) for each participant. This is why the number of obs/rows doubles. In other words we are using a repeated measures design that looks at lh and rh independently and then combines them to generate a single association metric. [This is a very rough explanation to help facilitate understanding]

```{r, gen long-format data}
# colnames of imaging data
dat_colnames = colnames(targetdata)[grep('lh\\.|rh\\.',colnames(targetdata))] 

# colnames of non-imaging data/unilateral img data
cols_nonimg = colnames(targetdata)[!grepl('lh\\.|rh\\.',colnames(targetdata))]  
cols_nonimg = cols_nonimg[2:length(cols_nonimg)]

source('../functions/long_format_new.R')
targetdata_longformat <- long_format(targetdata,cols_nonimg,cols_img)

```

Define functions

**Note: 30/11/21 - had to update the "select" function in reg_phewasStyle_withCI.R function to specify that it was dplyr::select as it wasn't working propertly. 

```{r, Phewas function}

source('../functions/reg_phewasStyle_withCI.R')

```

Define global variables

```{r, global vars}
targetdata=targetdata
dat_long=targetdata_longformat
```

Set up dependent variables (brain measures: cortical(APARC files); subcortical(ASEG files); whitematter(dti)

```{r, set up DVs}
ls.dep.fs.long=colnames(dat_long)[grep('sa\\.|thk\\.|vol\\.|sulc\\.|dtifa\\_|dtimd\\_|\\.ASEG\\.',colnames(dat_long))]
ls.dep.fs.short=colnames(targetdata)[grep('^bl\\.',colnames(targetdata))]
ls.dep.all=c(ls.dep.fs.long,ls.dep.fs.short)

ls.dep.all=ls.dep.all[c(grep('sa\\.|thk\\.|vol\\.|sulc\\.',ls.dep.all),
                        grep('\\.ASEG\\.',ls.dep.all),
                        grep('dtifa\\_',ls.dep.all),
                        grep('dtimd\\_',ls.dep.all))]
ls.dep.all=ls.dep.all[!duplicated(ls.dep.all)]

```

Set up Independent Variables (or factors) (depression measures)

  Depression symptom severity = (ordinal measure:none, mild, moderate, severe (rated 0-3, respectively)) 
                              = KSADS.Depressive_symptoms_ever
  
**Note: We are no longer using MDD diagnosis = KSADS.MDD (1=yes, 0= no, binary) due to possible errors in the KSADS MDD diagnosis algorithm flagged by the ABCD team in 4.0 release notes (November 2021)

```{r, set up IVs}

ls.factor=colnames(targetdata)[grep('^KSADS\\.Depressive_symptoms_ever',
                                    colnames(targetdata))]
ls.factor=c(ls.factor[grep('.p$',ls.factor)]) #ls.factor[grep('.y$',ls.factor)]) #we are not including youth report in current analysis. 

```

Combine DVs and IVs so that we get a list of the DV and IV

```{r, combine DVs and IVs}

ls.dep.factor.combo=expand.grid(ls.dep.all,ls.factor,stringsAsFactors = F)

```

Set up Covariates Note: We are adding hemisphere as a covariate here...

```{r, set up covs}

ls.models=data.frame(dependent=ls.dep.factor.combo$Var1,
                     factor=ls.dep.factor.combo$Var2,
                     covs='',stringsAsFactors = F)

ls.models$covs=paste0(c('age_years','bmi','abcd_site','race.6level','fsqc_qu_motion'),collapse='+')

ls.models$covs[!grepl('bl\\.',ls.models$dependent)]=paste0(ls.models$covs[!grepl('bl\\.',ls.models$dependent)],'+hemi')

```

#### STEP 4: Specify models

```{r, specify models}

ls.models$model.est=''
ls.models$model.est[grep('hemi',ls.models$covs)]='lme'
ls.models$model.est[ls.models$model.est=='']='glm'

```

Divide models into bulk measures and individual regions

Bulk (global) measures

```{r, bulk measures}
ls.model.bulk=ls.models[grep('mean|total|allfibers',ls.models$dependent),]
ls.model.bulk$p_batch=1
```

Individual regions

```{r, individual regions}

ls.model.region=ls.models[!grepl('mean|total|allfibers',ls.models$dependent),]

ls.model.region$p_batch=99999
target.model=ls.model.region
ls.dep.cate=c('thk\\.','sa\\.','vol\\.APARC','sulc\\.','vol\\.ASEG','dtifa\\_','dtimd\\_')
ls.factor.cate=unique(target.model$factor)
cate.no = 1
for (fac in ls.factor.cate){
      for (dep in ls.dep.cate){
            loc = grepl(dep,target.model$dependent)&grepl(fac,target.model$factor)
            target.model$p_batch[loc]=cate.no
            cate.no=cate.no+1
      }
}
ls.model.region=target.model

# individual region models add intracranial volume as a covariate
ls.model.region.covWholeB=ls.model.region
ls.model.region.covWholeB$covs[grep('thk\\.|sa\\.|vol\\.APARC|sulc\\.|vol\\.ASEG',ls.model.region.covWholeB$dependent)]=
      paste0(ls.model.region.covWholeB$covs[grep('thk\\.|sa\\.|vol\\.APARC|sulc\\.|vol\\.ASEG',ls.model.region.covWholeB$dependent)],'+ICV_ASEG')

```

#### STEP 5: Run Models

```{r, run models}

#Global (bulk) models
result.females.dep_brain.bulk=reg_phewasStyle(ls.model.bulk,dat_short=targetdata,dat_long=dat_long,correctByFactor = T)

#Regional (Individual) models
result.females.dep_brain.region.covWholeB=reg_phewasStyle(ls.model.region.covWholeB,dat_short=targetdata,dat_long=dat_long,correctByFactor = T)

#Save as R.Data file
save(result.females.dep_brain.bulk,result.females.dep_brain.region.covWholeB,file='../results/females_dep_brain_pilot_analysis_no_MDD.RData')

#add TRUE column if nominal p value < 0.005 for pilot ROIs

#global results
result.females.dep_brain.bulk$Sig <- as.numeric((as.character(result.females.dep_brain.bulk[,"p.value"]))) < 0.05

#regional results
result.females.dep_brain.region.covWholeB$Sig <- as.numeric((as.character(result.females.dep_brain.region.covWholeB[,"p.value"]))) < 0.05

#Save results as .csv files

#global results
write.csv(result.females.dep_brain.bulk, file = "../results/tables/pilot_analy_females.dep_brain.bulk_no_MDD.csv")

#regional results 
write.csv(result.females.dep_brain.region.covWholeB, file = "../results/tables/pilot_analy_females.dep_brain.regional_no_MDD.csv")

```

STEP 6: Identify ROIs for confirmatory analysis. Threshold: Nominal p-value < 0.005

```{r, make tables}

ROI_dep_brain_females_regional_results <- result.females.dep_brain.region.covWholeB %>% 
  filter(p.value <0.05)


#save as R.data
save(ROI_dep_brain_females_regional_results,file='../results/ROIs_dep_brain_females_pilot_analysis_no_MDD.RData')

#save as .csv
write.csv(ROI_dep_brain_females_regional_results, file = "../results/tables/ROIs_dep_brain_females_no_MDD.csv")

str(ROI_dep_brain_females_regional_results) #check variable chars

# ROI_males_regional %>% 
#        kable(digits=4, format = "html",
#        caption = "Female ROIs (PT~Regional Brain Measures)") %>% 
#                kable_styling(font_size = 15,
#                              position = "left",
#                              full_width = F) #%>% 
#       #save_kable(file = "../figs/female_ROIs_PT_brain_pilot_analysis.png")

```

############## End of Script ##################

