---
title: "ANALY_pilot_dep_brain_males"
author: "Niamh MacSweeney"  Credit to X. Shen for function templates. 
date: "20/04/2022"
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
#ABCD Pilot Analysis Script - Pubertal timing ~ Brain Structure -> Males

####Introduction

The purpose of this script is to run the pilot analysis (PT~BS) for the Puberty RR. For this analysis, we will be using a random sub-sample of 20% (N= ~1000 for males and females each, Total N=~1,700) to generate ROIs for the pubertal timing ~ brain structure association.

This sub-sample will be split into males and females and the script will be run independently for each.

Note that this sample includes unrelated individuals. 

Please note that in this script, brain structure is set as the dependent variable due to the treatment of hemisphere mixed effects. However, in the RR, our hypotheses state that depression is the dependent or outcome variable. As this analysis is cross-sectional, the Y and X terms are statistically interchangeable so Y ~ X = pubertal timing ~ brain = brain ~ pubertal timing. For our later mediation models, the order of the terms is important so depression will be the dependent or outcome variable. See RR manuscript for further details on mediation model set up.

Required Inputs

Cleaned data Rds files are located in /Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data

Independent Variable — Pubertal timing: pds_timing_baseline_R4.0.rds

Dependent Variable — Brain structure: cortical_baseline_R4.0.rds + subcort_baseline_R4.O.rds + dti_baseline_R4.0.rds

Covariates — Covariates: covs_pilot_R4.0.rds

------------------------------------------------------------------------

####Workflow

1.  Setup: Load data, merge to create master dataframe, keep unrelated participants only
2.  Pilot prep: Get subsample of 20% of sample (N= ~1,700). Split into males and females.
3.  Set up  function
4.  Run models

####STEP 1: SETUP Load libraries and set wd

```{r, load libraries, set wd}


library(tidyverse)
library(lme4)
library(lmerTest)
library(nlme)
library(psych)
library(ggplot2)
library(readr)
library(pbapply)
library(gridExtra)
library(sandwich)
library(stringr)
library(kableExtra)
#library(epiDisplay)
library(gmodels)
library(caret)


setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/analy")
```

Load data:

```{r,load data}

#Pubertal timing data 
pds <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/pds_timing_baseline_R4.0.rds")

#Imaging data
cortical <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/cortical_baseline_R4.0.rds")

subcort <-rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/subcort_baseline_R4.0.rds")

dti <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/dti_baseline_R4.0.rds")

#Covariates
covs <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/covs_pilot_R4.0.rds")

```

Merge dataframes to create master dataframe

As we have a five data frames to merge, we are going to do it stepwise using dplyr function.

```{r, merge DFs}

#before merge, remove extra variables in pds

pds <- pds %>%
  select(-c("eventname", "sex", "age_years"))

df <- pds %>% 
  left_join(y=cortical, by= "src_subject_id") %>% 
  left_join(y=subcort, by= "src_subject_id") %>% 
  left_join(y=dti, by= "src_subject_id") %>% 
  left_join(y=covs, by="src_subject_id")

#remove duplicate cols after merge (e.g., ICV and WBV)

df <- df %>% 
  select(-c(ICV.y, WBV.y))
#rename remaining cols
df <- df %>% 
  rename(ICV = ICV.x)

df <- df %>% 
  rename(WBV = WBV.x)

#check col names
colnames(df)

```

Check variable classifcation

```{r, variable classification}

df %>% 
 select("rel_family_id", "race.6level", "age_years", "scanner_id", "dmri_dti_meanmotion") %>% 
  str()
```


Get sample demographics: Age = 9.91 years; SD = 0.62

```{r, sample demographics}

df %>% 
  dplyr::select(age_years, sex) %>% 
  summary()

#get sd
sd(df$age_years)
#get mean
mean(df$age_years)

#make df with males only for demographic table purposes
males <- df %>% #N= 6193
    filter(sex == "M")
  
```

IGNORE for pilot analysis: Inspect Ns per nested condition
```{r, inspect Ns per condition}

#let's look at how the data is nested for females
nest_str_all <- df %>%
  group_by(rel_family_id, src_subject_id, scanner_id) %>%
  count()

#check if individuals from each family went to the same scanner
#This is because we plan to nest family id within scanner id 
#group by family id and then count number of unique scanner_ids per family group
#N=47 individuals within each family that went to different scanners
scanner_test <- df %>% 
  group_by(rel_family_id) %>%  
  na.omit() %>%  #some missing values for scanner_id 
  distinct(scanner_id) %>% 
  summarise(scanner_match = n()) %>% 
  count()

```


#####STEP 2: Generate subsample (20% of unrelated sample) for pilot analysis. N=1188 Get males only

Note: name of object needs to be targetdata for function

1. Get unrelated sample N= 9850
```{r, get unrelated sample using randomisation approach}

#Assign random number to each family ID. Set random 

# df$random <- NA #generate new column of NAs
# unrel_df <- data.frame() #make new df
# 
# #iterate through family IDs using for loop (this will take a few minutes)
# # set.seed(2507) #to make sure random sample is the same random sample each time
# unique_fam_id <- unique(df$rel_family_id)
# for (i in unique_fam_id[1:length(unique_fam_id)]){
#   
#   set.seed(2507)
#   
#   #subset each family and assign random number to each individual within each family
#   df$random[df$rel_family_id == i] <- sample(nrow(df[df$rel_family_id == i,]),nrow(df[df$rel_family_id == i,]), replace= FALSE)
#   
#   #select individual with highest number for unrelated dataframe
#   unrel_individual <- df[df$rel_family_id == i,][which.max(df$random[df$rel_family_id == i]),]
#     
#   #append unrelated individual to unrelated dataframe
#   unrel_df <- rbind(unrel_df, unrel_individual)
#   
# }
# 
# # i <- 8781 #for trouble shooting. 
# nrow(unrel_df) #check that N= 9146
# length(unique(unrel_df$rel_family_id)) #check that no. of unique IDs is correct. 
# 
# #Check that set.seed worked so that we get the same IDs each timescript is run
# 
# test_set_seed <- unrel_df$src_subject_id #make new vector of un_rel individual ids
# 
# #the rerun df$random lines + for loop, without clearing global environment
# #check if new participant id list matches test_set_seed
# 
# all(unrel_df$src_subject_id == test_set_seed) #it worked! 

```

```{r, generate subsample, get males only}

set.seed(2507) #set seed for reproducibility
targetdata <- df 
# %>% 
#   sample_frac(size = 0.2, replace = FALSE)

#check males and females breakdown F= 564, M=623
targetdata %>% 
  dplyr::count(sex)

#get mean age and sd for pilot sample. Mean age = 9.91, SD = 0.62
mean(targetdata$age_years)
sd(targetdata$age_years)

#Get Female sample only 
targetdata <- targetdata %>% 
  filter(sex == "M")

```

Remove columns not needed for pilot analyis
```{r, remove cols not needed for pilot analysis}
# targetdata <- targetdata %>% 
#   select(-c("bmi", "household_income", "asr_scr_depress_r", "interview_age", "site_id_l", "dmri_dti_meanrot", "dmri_dti_meantrans", "acs_raked_propensity_score" ,"mri_info_visitid", "mri_info_studydate", "rel_group_id", "random"))
```

TROUBLESHOOTING: Check for multicolinearity

```{r, multicolinearity}
# 
# #only include numeric variables
# mydata <- targetdata %>% 
#   dplyr::select(where(is.numeric))
# 
# #check colinearity with dti variables only 
# 
# dti_corr <- targetdata %>% 
#   dplyr::select(matches('age_years|race.6level|scanner_id|dti'))
# 
# dti_corr <- dti_corr %>% 
#   dplyr::select(-(contains("allfibers")))
# 
# #numeric only
# dti_corr <- dti_corr %>% 
#   dplyr::select(where(is.numeric))
# #make correlation matrix
# dti_corr_mat = cor(dti_corr, method = "pearson", use = "pairwise.complete.obs")
# 
# #get column names that have a correlation >0.95
# tmp <- dti_corr_mat > 0.7
# diag(tmp) <- FALSE
# names(Filter(function(x) x > 0, rowSums(tmp) > 0 | colSums(tmp) > 0))

```

####STEP 3: Model set up We are using X.Shen's reg_phewasStyle_Random.R function to run the mixed effects models for both unilateral and bilateral brain regions.
We need to change "src_subject_id" to "f.eid" so that function will run.

```{r, colnames}
colnames(targetdata)[1]='f.eid'
colnames(targetdata) #check it worked

#check for duplicate columns
unique_names <- unique(colnames(targetdata)) #no duplicates 

```

Define Basic Model for pilot analysis

IV: Pubertal timing (pt_f)

DV: Brain Measures

Covariates: Fixed = age, ethnicity; dti mean motion for (DTI models), hemisphere for bilateral (LME) models

Random: participant id (f.eid), scanner ID.

For main analysis, family id will be nested in scanner_id. 


Set categorical variables

```{r, set categorical variables}

targetdata$race.6level = as.factor(targetdata$race.6level)
targetdata$rel_family_id = as.factor(targetdata$rel_family_id)
targetdata$scanner_id = as.factor(targetdata$scanner_id)

```
Standardise variables before running models to avoid scaling issues. Note, we will only scale the numeric variables

```{r, scale variables}

targetdata <- targetdata %>% 
mutate_if(is.numeric, scale)

```

Generate long-format data.

Note: for the lme models (bilateral brain regions only), the data needs to be in long-format.  Hemisphere is controlled for as a fixed, within-participant effect.

```{r, gen long-format data}
# # colnames of imaging data
# dat_colnames = colnames(targetdata)[grep('lh\\.|rh\\.',colnames(targetdata))] 
# 
# # colnames of non-imaging data/unilateral img data
# cols_nonimg = colnames(targetdata)[!grepl('lh\\.|rh\\.',colnames(targetdata))]  
# cols_nonimg = cols_nonimg[2:length(cols_nonimg)]
# 
# source('../functions/long_format_new.R')
# targetdata_longformat <- long_format(targetdata,cols_nonimg,cols_img)

```

Define functions

Note: Make sure that the "family" argument in function is correct (i.e., we want it to be "poisson" not "binomial")


```{r, Phewas function}

source('../functions/reg_phewasStyle_Random.R')

```

Define global variables

```{r, global vars}
targetdata=targetdata
# dat_long=targetdata_longformat

```

Set up dependent variables (brain measures: cortical(APARC files); subcortical(ASEG files); whitematter(dti)

```{r, set up DVs}

#select vars of interest
#ls.dep.fs.long=colnames(dat_long)[grep('sa\\.|thk\\.|vol\\.|sulc\\.|\\.ASEG\\.|dtifa\\_|dtimd\\_',colnames(dat_long))] 
#For all models that are not bilateral, we will keep data in short format
#Note that these DTI measures: CC, fmaj and fmin are unilateral, so these need to be listed here too.
ls.dep.fs.short=colnames(targetdata)[grep('^bl\\.|^avg\\.|\\_cc|\\_fmaj|\\_fmin',colnames(targetdata))]
ls.dep.all=ls.dep.fs.short

# ls.dep.all=ls.dep.all[c(grep('sa\\.|thk\\.|vol\\.|sulc\\.',ls.dep.all),
#                         grep('\\.ASEG\\.',ls.dep.all),
#                         grep('dtifa\\_',ls.dep.all),
#                         grep('dtimd\\_',ls.dep.all))]
ls.dep.all=ls.dep.all[!duplicated(ls.dep.all)]

```

Set up Independent Variable (or factor) (Pubertal timing for males)


```{r, set up IVs}

ls.factor=colnames(targetdata)[grep('^pt_m',
                                    colnames(targetdata))]

```

Combine DVs and IVs so that we get a list of the DV and IV

```{r, combine DVs and IVs}

ls.dep.factor.combo=expand.grid(ls.dep.all,ls.factor,stringsAsFactors = F)

```

Set up Covariates Note: We are adding hemisphere as a covariate here...

```{r, set up covs}

ls.models=data.frame(dependent=ls.dep.factor.combo$Var1,
                     factor=ls.dep.factor.combo$Var2,
                     covs='',stringsAsFactors = F)

ls.models$covs=paste0(c('age_years','race.6level'),collapse='+')

#For all models that are not bilateral, add hemisphere as a covariate. 
#Note that these DTI measures: CC, fmaj and fmin are unilateral, so these need to be listed here too. 
# ls.models$covs[!grepl('bl\\.|\\_cc|\\_fmaj|\\_fmin',ls.models$dependent)]=paste0(ls.models$covs[!grepl('bl\\.|\\_cc|\\_fmaj|\\_fmin',ls.models$dependent)],'+hemi')

#add dti mean motion (fd) as cov for dti models
ls.models$covs[grepl('dtifa\\_|dtimd\\_',ls.models$dependent)]=paste0(ls.models$covs[grepl('dtifa\\_|dtimd\\_',ls.models$dependent)],'+dmri_dti_meanmotion')

```

#### STEP 4: Specify models

```{r, specify models}

ls.models$model.est=''
ls.models$model.est[grep('hemi',ls.models$covs)]='bilateral'
ls.models$model.est[ls.models$model.est=='']='unilateral'


```

Divide models into bulk measures and individual regions

Bulk (global) measures

```{r, bulk measures}
ls.model.bulk=ls.models[grep('mean|total|allfibers',ls.models$dependent),]
ls.model.bulk$p_batch=1

```

Individual regions

```{r, individual regions}

ls.model.region=ls.models[!grepl('mean|total|allfibers',ls.models$dependent),] 

ls.model.region$p_batch=99999
target.model=ls.model.region
ls.dep.cate=c('thk\\.','sa\\.','vol\\.APARC','sulc\\.','vol\\.ASEG','dtifa\\_','dtimd\\_')
ls.factor.cate=unique(target.model$factor)
cate.no = 1
for (fac in ls.factor.cate){
      for (dep in ls.dep.cate){
            loc = grepl(dep,target.model$dependent)&grepl(fac,target.model$factor)
            target.model$p_batch[loc]=cate.no
            cate.no=cate.no+1
      }
}
ls.model.region=target.model

# individual region models add whole brain volume as a covariate
ls.model.region.covWholeB=ls.model.region
ls.model.region.covWholeB$covs[grep('thk\\.|sa\\.|vol\\.APARC|sulc\\.|vol\\.ASEG',ls.model.region.covWholeB$dependent)]=
      paste0(ls.model.region.covWholeB$covs[grep('thk\\.|sa\\.|vol\\.APARC|sulc\\.|vol\\.ASEG',ls.model.region.covWholeB$dependent)],'+WBV')


```

#### STEP 5: Run Models

```{r, run models}

#Global (bulk) models
result.males.pt_brain.global=reg_phewasStyle(ls.model.bulk,dat_short=targetdata,dat_long=dat_long,correctByFactor = T)

#Regional (Individual) models
result.males.pt_brain.region.covWholeB=reg_phewasStyle(ls.model.region.covWholeB,dat_short=targetdata,dat_long=dat_long,correctByFactor = T)

#Save as R.Data file
save(result.males.pt_brain.global,result.males.pt_brain.region.covWholeB,file='../results/revised_pilot/males_pt_brain_pilot.RData')

#add TRUE column if nominal p value < 0.02 for pilot ROIs, which means that all values of ≤0.01... will be included due to the large number of decimal places. 

#global results
result.males.pt_brain.global$Sig <- as.numeric((as.character(result.males.pt_brain.global[,"p.value"]))) < 0.02

#regional results
result.males.pt_brain.region.covWholeB$Sig <- as.numeric((as.character(result.males.pt_brain.region.covWholeB[,"p.value"]))) < 0.02

#Save results as .csv files

#global results
write.csv(result.males.pt_brain.global, file = '../results/revised_pilot/tables/pilot_males.pt_brain.bulk.csv')

#regional results 
write.csv(result.males.pt_brain.region.covWholeB, file = '../results/revised_pilot/tables/pilot_males_pt_brain.regional.csv')

```

STEP 6: Identify ROIs for confirmatory analysis. Threshold: Nominal p-value \< 0.01

Note: The threshold is indicated as <0.02 to account for the large number of decimal places given for p-values.
```{r, make tables}

ROI_pt_brain_males_regional_results <- result.males.pt_brain.region.covWholeB %>%
  filter(p.value <0.02)

#save as R.data
save(ROI_pt_brain_males_regional_results,file='../results/revised_pilot/ROIs_pt_brain_males.RData')

#save as .csv
write.csv(ROI_pt_brain_males_regional_results,file= "../results/revised_pilot/tables/ROIs_pt_brain_males.csv")

#save as Kable Table

# ROI_table <- ROI_dep_brain_females_regional_results %>%
#        kable(digits=4, format = "html",
#       caption = "Females: Depressive symptoms-brain structure models and associated statistics with significant ROI associations") %>%
#                kable_styling(font_size = 15,
#                              position = "left",
#                              full_width = F) #%>%
#       #save_kable(file = "../figs/female_ROIs_PT_brain_pilot_analysis.png")
# 
# str(ROI_dep_brain_females_regional_results) #check variable chars


```

# \#\#\#\#\#\#\#\#\#\#\#\#\#\# End of Script \#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#

Make demographic table for paper - where we have a breakdown of the MDD and DS categories for each sex

```{r, demographic table}


```
