---
title: "ANALY_descriptives.Rmd"
author: "Niamh MacSweeney"
date: "2022-08-31"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales) #for pie chart
library(waffle)
library(tidyverse)
library(psych)
library(ggplot2)
library(waffle)
library(ggpubr)
library(pbapply)
library(gridExtra)
library(sandwich)
library(stringr)
library(kableExtra)
library(gtsummary)


setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")
```

##SET UP

#Load data

```{r, load data}

pds_y1 <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/pds_timing_y1_R4.0.rds")
  
dep_vars <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/CBCL_dep_all_yrs_R4.0.rds")
  
covs <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/covs_main_R4.0.rds")

```

#Tidy data
```{r, tidy up variables}

pds_y1 <- pds_y1 %>%  
  select (-c("pds_m_pt_valid", "pds_f_pt_valid", "age_years")) #remove extra age column too. 

```

## Merge
```{r, merge}
#inspect vars before merge
colnames(covs)
colnames(dep_vars)
colnames(pds_y1)

#merge by src_subject_id and sex as they are common to all dataframes 

#merge 1
df <- covs %>% 
  left_join(y=dep_vars, by= c("src_subject_id", "sex"))
  
#merge 2
df <- df %>% 
  left_join(y=pds_y1, by= c("src_subject_id", "sex"))

#check if merge worked okay
colnames(df)

```

#REDUCE TO SAMPLE NEEDED

For this RR, our final sample size requires participants to have complete pubertal data (i.e., complete PDS total score). Participants also need to have attended year 3 follow up. Due to the partial data release of year 3 follow up data in Release 4.0, only ~50% of depression data is available. For participants that attended year 3 follow up but do not have complete depression outcome measures, we will impute the depression scores (see PREP_imputation script for more details)

###Total N = 5727

```{r, filter sample}

#get summary first
df %>% 
  select(src_subject_id, pds_tot_all, interview_date_y3) %>% 
summary() 

#get number of complete cases for pds total and interview date y3
sum(complete.cases(df$pds_tot_all)) #N = 10,400
sum(complete.cases(df$interview_date_y3)) #6251


final_df <- df %>% 
  filter(!is.na(pds_tot_all)) %>% #N=10,400
  filter(!is.na(interview_date_y3)) #N = 5727

```


```{r, get list of vars}


#all vars including aux vars

main_vars <- c("src_subject_id", #character variable
               "age_y1",
               "age_y2",
               "age_y3",
               "sex",    #character variable                    
               "site_id",  
              "parent_dep_y2", #continuous score
               "acs_raked_propensity_score_y0" , 
               "bmi_y1", 
               "household_income_y0", #character variable
               "race_y0", 
              "dti_mean_motion_y2",  #continuous score       
               "scanner_id_y2", #character variable
               "cbcl_withdep_y3", #continuous score
               "pds_tot_all",
              "pt_f",
              "pt_m", #continuous score
              "gonad_avg",
              "adren_avg",
              "pds_p_ss_female_category",
              "pds_p_ss_male_category"
               )

# main_vars <- c("Study ID" = "src_subject_id", #character variable
#                "Sex",    #character variable                    
#                "Site ID" = "site_id",  
#                "Parent depressive symptoms" = "parent_dep_y0", 
#                "Population stratification weight" = "acs_raked_propensity_score_y0",
#                "BMI" = "bmi_y0",
#                "Household income" = "household_income_y0", 
#                "Race/ethnicity" = "race_y0", 
#                "DTI mean FD" = "dti_mean_motion_y2",  #continuous score       
#                "Scanner ID" = "scanner_id_y2",
#                "Age" = "age_years", #continuous score
#               "Youth depressive symptoms" = "cbcl_withdep_y3", #continuous score
#                "PDS total score" = "pds_tot_all",
#               "PT females" = "pt_f",
#               "PT males" = "pt_m", #continuous score
#                )
```


```{r, tidy}

#Extract main variables 

main_df <- final_df[,main_vars]

#reorder columns for table 

main_df <- main_df %>% 
  select("src_subject_id", 
          "sex",
          "age_y1",
          "age_y2",
          "age_y3",
          "pds_tot_all",
          "cbcl_withdep_y3", 
          "race_y0",
          "bmi_y1", 
          "household_income_y0",
          "dti_mean_motion_y2",
          "parent_dep_y2",
          "acs_raked_propensity_score_y0" , 
          "site_id", 
          "scanner_id_y2",
          "pt_f",
          "pt_m",
         "gonad_avg",
          "adren_avg",
          "pds_p_ss_female_category",
          "pds_p_ss_male_category"
               )

#tidy household income levels 
str(main_df$household_income_y0)
summary(main_df$household_income_y0)
#make factor 
main_df$household_income_y0 <- as.factor(main_df$household_income_y0)
main_df$household_income_y0 <- recode_factor(main_df$household_income_y0, 
                                             "1" = "<$5000",
                                             "2" = "$5,000-$11,999",
                                             "3" = "$12,000-$15,999",
                                             "4" = "$16,000-$24,999",
                                             "5" = "$25,000-$34,999",
                                             "6" = "$35,000-$49,999",
                                             "7" = "$50,000-$74,999",
                                             "8" = "$75,000-$99,999",
                                             "9" = "$100,000-$199,999",
                                             "10" = ">$200,000")

```

#MAKE TABLE

Make descriptive table
```{r, descriptives table}

des_table <- main_df %>% 
  select(-c(src_subject_id, site_id, scanner_id_y2, pt_f, pt_m, acs_raked_propensity_score_y0, gonad_avg, adren_avg, 
            pds_p_ss_female_category, pds_p_ss_male_category)) %>% 
  tbl_summary(
    by = sex,
    statistic = list(all_continuous() ~ c("{mean} ({sd})"),
                     all_categorical() ~ c("{n} / {N} ({p}%)")),
    digits = list(all_continuous() ~ c(2,2,0,0),
                  all_categorical() ~ c(0,0,2)),
    label = c(cbcl_withdep_y3 ~ "Youth depressive symptoms", pds_tot_all ~  "PDS total score",
            age_y1 ~ "Age (Y1)",
            age_y2 ~ "Age (Y2)",
            age_y3 ~ "Age (Y3)",
            race_y0 ~ "Race/ethnicity",
            parent_dep_y2 ~ "Parent depressive symptoms",
            bmi_y1 ~ "BMI (Y1)",
            household_income_y0 ~ "Household income",
            dti_mean_motion_y2 ~ "DTI mean FD"),
    missing_text = "Missing",
    type = list (all_continuous() ~ "continuous"),
    missing = "ifany"
    ) %>% 
  bold_labels () %>% 
  italicize_levels() %>% 
  add_p() %>% 
  modify_table_body(
    dplyr::mutate,
    label = ifelse(label == "Missing",
                   "Missing (N)",
                   label)) %>% 
   as_gt() %>%             # convert to gt table
  gt::gtsave(             # save table as image
    filename = "/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/figs/descriptives_tbl.png"
)


```

#MAKE DISTRIBUTION FIGURES 

##Puberty figs

```{r, puberty figures}

#convert sex variable to factor

colnames(main_df)
str(main_df)

main_df$sex <- as.factor(main_df$sex)

#plot pds total
pds_tot <- ggplot(main_df, aes(pds_tot_all))

pds_tot + geom_bar(aes(fill = sex)) +
  ggtitle("PDS total score") +
  labs (x = "Total PDS") +
  theme(plot.title = element_text(face = "bold"))
pds_tot


#make histogram of gonadal and adrenal development within each group 
gonad_hist <- main_df %>%
  ggplot(aes(x=gonad_avg, color=sex, fill=sex)) +
  geom_histogram(bins = 7, alpha = 0.5, position = "identity") +
  ggtitle("Gonadal score") +
   labs (x = "Gonadal PDS") +
  theme(plot.title = element_text(face = "bold"))
gonad_hist

adren_hist <-  main_df %>%
  ggplot(aes(x=adren_avg, color=sex, fill=sex)) +
  geom_histogram(bins = 7, alpha = 0.5, position = "identity") +
  ggtitle("Adrenal score") +
   labs (x = "Adrenal PDS") +
  theme(plot.title = element_text(face = "bold"))
adren_hist

#PDS category data - there seems to be some missing data in these variables provided by ABCD. 
#let's check
summary(main_df$pds_p_ss_female_category)
  
#tidy pds category data
cat_df <- main_df %>% mutate(pds_cat = coalesce(pds_p_ss_female_category, pds_p_ss_male_category)) %>%
         select(src_subject_id, pds_cat, sex)

str(cat_df)
cat_df$pds_cat <- as.factor(cat_df$pds_cat) # N = 60 missing

#get levels in factor
levels <- cat_df %>% 
  group_by(pds_cat) %>%
  summarise(no_rows = length(pds_cat))

#rename factor levels 
cat_df$pds_cat <- recode_factor(cat_df$pds_cat, 
                      `1` = "Pre",
                      `2` = "Early",
                      `3` = "Mid",
                      `4` = "Late",
                      `5` = "Post",
                      `NA` = "NA"
)

cat_hist <- cat_df %>%
  drop_na(pds_cat) %>% 
  ggplot(aes(x = pds_cat, fill = sex)) +
  geom_bar()  +
   ggtitle("PDS Category") +
  labs(x = "") +
    theme(plot.title = element_text(face = "bold"))
cat_hist

#with count scores
cat_hist2 <- cat_df %>%
  drop_na(pds_cat) %>% 
  ggplot(aes(x = pds_cat, fill = sex)) +
  geom_bar(position = "fill") + 
  geom_text(stat='count', aes(label=..count..), position = "fill", vjust=1.5) +
   ggtitle("PDS Category") +
  labs(x = "") +
    theme(plot.title = element_text(face = "bold"))
cat_hist2



  
```

## Depression figs
```{r, depression figs}


dep_hist <- main_df %>%
  ggplot(aes(x=cbcl_withdep_y3, color=sex, fill=sex)) +
  geom_histogram(bins = 7, alpha = 0.3, position = "dodge2") +
  ggtitle("Youth depressive symptoms") +
   labs (x = "CBCL Withdrawn/Depressed (raw score)") +
  theme(plot.title = element_text(face = "bold"))
dep_hist

```





####END OF MAIN SCRIPT #######

Variables:

(split by sex)
- PDS total
-CBCL depression
- BMI

(overall sample)
- Income
- Race

## Density plots

```{r, density plots}
# #create separate df with variables we want to plot
# plot_df <- main_df %>% 
#   select(src_subject_id, sex, pds_tot_all, cbcl_withdep_y3, bmi_y0, household_income_y0, race_y0)
# 
# summary(plot_df)
# 
# #PDS TOTAL 
# 
# #get mean and save as tmp variable mu
# mu <- ddply(plot_df, "sex", summarise, grp.mean=mean(pds_tot_all))
# head(mu)
# 
# (pds_plot <- ggplot(main_df, aes(x=pds_tot_all, fill=sex, group=sex)) +
#   geom_density(adjust=4, alpha=.4) +
#      geom_vline(data=mu, aes(xintercept=grp.mean, color=sex),
#              linetype="dashed") +
#   theme_classic()) +
#   labs(title=" Perceived pubertal development density curve",x="Pubertal Development Scale (total score)", y = "density") +
#   scale_fill_brewer(palette="Dark2") 
# 
# #CBCL depression
# mu <- ddply(plot_df, "sex", summarise, grp.mean=mean(cbcl_withdep_y3))
# head(mu)
# 
# (dep_plot <- ggplot(main_df, aes(x=cbcl_withdep_y3, fill=sex, group=sex)) +
#   geom_density(adjust=4, alpha=.4) +
#      geom_vline(data=mu, aes(xintercept=grp.mean, color=sex),
#              linetype="dashed") + 
#     xlim(0,15) +
#   theme_classic()) +
#    labs(title="Depressive symptoms density curve",x="CBCL Withdrawn Depressed (raw score)", y = "density") +
#     scale_fill_brewer(palette="Dark2")
#  
# #BMI 
# mu <- ddply(plot_df, "sex", summarise, grp.mean=mean(bmi_y0))
# head(mu)
# 
# (bmi_plot <- ggplot(main_df, aes(x=bmi_y0, fill=sex, group=sex)) +
#   geom_density(adjust=4, alpha=.4) +
#      geom_vline(data=mu, aes(xintercept=grp.mean, color=sex),
#              linetype="dashed") +
#   theme_classic()) +
#   xlab("Body Mass Index (BMI)") +
#   ylab("density") +
#    labs(title="BMI density curve",x="Body Mass Index (BMI)", y = "density") +
#     scale_fill_brewer(palette="Dark2")
# 
# 
# des_fig_2 <- 
# #display plots together
# (des_fig <- ggarrange(
#   pds_plot, 
#   ggarrange(dep_plot, bmi_plot, ncol= 2, labels = c("B", "C")),
#                nrow = 2,
#             labels = "A")
#             )
# 
# ggarrange(
#   lp,                # First row with line plot
#   # Second row with box and dot plots
#   ggarrange(bxp, dp, ncol = 2, labels = c("B", "C")), 
#   nrow = 2, 
#   labels = "A"       # Label of the line plot
#   ) 


```

















##Waffle charts for Race and income  - waffle charts too complicated. 

Stick to bar chats 

```{r, bar charts}
# # Data transformation
# plot_df$race_y0 <- as.character(plot_df$race_y0)
# 
# race <- plot_df %>% 
#   select(race_y0) %>% #select race variable only to get correct freq. 
#   dplyr::group_by(race_y0) %>% 
#   count() %>% 
# dplyr::ungroup() %>% 
#   mutate(perc = `freq` / sum(`freq`)) %>%
#   arrange(perc) %>%
#   mutate(labels = scales::percent(perc))
# 
# race_chart <- ggplot(race, aes(x = "", y = perc, fill = (race_y0)) +
#   geom_col() +
#   geom_text(aes(label = ""),
#             position = position_stack(vjust = 0.5)) +
#   coord_polar(theta = "y") +
#   theme_void()
# )
# 
# 
# 
# 
# 
# ggplot(race, aes(x = "", y = perc, fill = (race_y0)) +
#        geom_bar(width = 1, stat = "identity") +
#        coord_polar("y", start = 0) +
#        geom_label_repel(aes(label = perc), size=5, show.legend = F, nudge_x = 1) +
#        guides(fill = guide_legend(title = "Group"))
#               )
#        
# 
# ###code working above this####
# 
# 
# ggplot(race, aes(values = perc, fill = race_y0)) +
#   geom_waffle(n_rows = 8, size = 0.33, colour = "white") +
#   scale_fill_manual(name = NULL,
#                     values = perc,
#                     labels = label) +
#   coord_equal() +
#   theme_void()
# 
#   geom_text(aes(label = labels),
#             position = position_stack(vjust = 0.5)) +
#   coord_polar(theta = "y")
  
  



```


