---
title: "ANALY_descriptives.Rmd"
author: "Niamh MacSweeney"
date: '2022-08-31'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales) #for pie chart
library(plyr) #for distribution plots #N.B- make sure to load before tidyverse and dplyr
library(waffle)
library(tidyverse)
library(psych)
library(ggplot2)
library(waffle)
library(ggpubr)
library(readr)
library(pbapply)
library(gridExtra)
library(sandwich)
library(stringr)
library(kableExtra)
library(gtsummary)
library(ggrepel)
library(forcats)

setwd("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/prep")
```

##SET UP

#Load data

```{r, load data}

pds_y1 <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/pds_timing_y1_R4.0.rds")
  
dep_vars <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/CBCL_dep_all_yrs_R4.0.rds")
  
covs <- rio::import("/Volumes/GenScotDepression/users/niamh/puberty_ABCD/ABCD_puberty_depression/data/covs_main_R4.0.rds")

```

#Tidy data
```{r, tidy up variables}

pds_y1 <- pds_y1 %>% 
  select (-c("pds_m_pt_valid", "pds_f_pt_valid"))

#the dep and covs dataframes are in wide format so before merge, check that all of pds variables are definitely from year 1
#they should be based on earlier QC

pds_y1 %>% 
  select(eventname) %>% 
  as.factor() #only one level when converted to factor --- so looks good! 

#remove eventname column before merge because we know that the pds vars are all from year 1

pds_y1 <- pds_y1 %>% 
  select(- c(eventname))
colnames(pds_y1)
```

## Merge
```{r, merge}
#inspect vars before merge
colnames(covs)
colnames(dep_vars)
colnames(pds_y1)

#merge by src_subject_id and sex as they are common to all dataframes 

#merge 1
df <- covs %>% 
  left_join(y=dep_vars, by= c("src_subject_id", "sex"))
  
#merge 2
df <- df %>% 
  left_join(y=pds_y1, by= c("src_subject_id", "sex"))

#check if merge worked okay
colnames(df)

```

#REDUCE TO SAMPLE NEEDED

For this RR, our final sample size requires participants to have complete pubertal data (i.e., complete PDS total score). Participants also need to have attended year 3 follow up. Due to the partial data release of year 3 follow up data in Release 4.0, only ~50% of depression data is available. For participants that attended year 3 follow up but do not have complete depression outcome measures, we will impute the depression scores (see PREP_imputation script for more details)

###Total N = 5727

```{r, filter sample}

#get summary first
df %>% 
  select(src_subject_id, pds_tot_all, interview_date_y3) %>% 
summary() 

#get number of complete cases for pds total and interview date y3
sum(complete.cases(df$pds_tot_all)) #N = 10,400
sum(complete.cases(df$interview_date_y3)) #6251


final_df <- df %>% 
  filter(!is.na(pds_tot_all)) %>% #N=10,400
  filter(!is.na(interview_date_y3)) #N = 5727

```


```{r, get list of vars}


#all vars including aux vars

main_vars <- c("src_subject_id", #character variable
               "sex",    #character variable                    
               "site_id",  
              "parent_dep_y2", #continuous score
               "acs_raked_propensity_score_y0" , 
               "bmi_y0", 
               "household_income_y0", #character variable
               "race_y0", 
              "dti_mean_motion_y2",  #continuous score       
               "scanner_id_y2", #character variable
               "age_years", #continuous score
               "cbcl_withdep_y3", #continuous score
               "pds_tot_all",
              "pt_f",
              "pt_m" #continuous score
               )

# main_vars <- c("Study ID" = "src_subject_id", #character variable
#                "Sex",    #character variable                    
#                "Site ID" = "site_id",  
#                "Parent depressive symptoms" = "parent_dep_y0", 
#                "Population stratification weight" = "acs_raked_propensity_score_y0",
#                "BMI" = "bmi_y0",
#                "Household income" = "household_income_y0", 
#                "Race/ethnicity" = "race_y0", 
#                "DTI mean FD" = "dti_mean_motion_y2",  #continuous score       
#                "Scanner ID" = "scanner_id_y2",
#                "Age" = "age_years", #continuous score
#               "Youth depressive symptoms" = "cbcl_withdep_y3", #continuous score
#                "PDS total score" = "pds_tot_all",
#               "PT females" = "pt_f",
#               "PT males" = "pt_m", #continuous score
#                )
```


```{r, tidy}

#Extract main variables 

main_df <- final_df[,main_vars]

#reorder columns for table 

main_df <- main_df %>% 
  select("src_subject_id", 
               "sex",
                "age_years",
          "pds_tot_all",
           "cbcl_withdep_y3", 
            "race_y0",
            "bmi_y0", 
            "household_income_y0",
            "dti_mean_motion_y2",
              "parent_dep_y2",
               "acs_raked_propensity_score_y0" , 
            "site_id", 
               "scanner_id_y2",
              "pt_f",
              "pt_m"
               )


```

#MAKE TABLE

Make descriptive table
```{r, descriptives table}

main_df %>% 
  select(-c(src_subject_id, site_id, scanner_id_y2, pt_f, pt_m, acs_raked_propensity_score_y0)) %>% 
  tbl_summary(
    by = sex,
    statistic = list(all_continuous() ~ "{mean} ({sd}) [{p_miss}]",
                     all_categorical() ~ "{n} / {N} ({p}%) [{p_miss}]"),
    digits = all_continuous() ~ 2,
    label = c(cbcl_withdep_y3 ~ "Youth depressive symptoms", pds_tot_all ~  "PDS total score",
            age_years ~ "Age",
            race_y0 ~ "Race/ethnicity",
            parent_dep_y2 ~ "Parent depressive symptoms",
            bmi_y0 ~ "BMI",
            household_income_y0 ~ "Household income",
            dti_mean_motion_y2 ~ "DTI mean FD"),
    missing_text = "Missing",
    
  ) %>% 
  italicize_levels()

```

#MAKE DISTRIBUTION FIGURES 

Variables:

(split by sex)
- PDS total
-CBCL depression
- BMI

(overall sample)
- Income
- Race

## Density plots

```{r, density plots}
#create separate df with variables we want to plot
plot_df <- main_df %>% 
  select(src_subject_id, sex, pds_tot_all, cbcl_withdep_y3, bmi_y0, household_income_y0, race_y0)

summary(plot_df)

#PDS TOTAL 

#get mean and save as tmp variable mu
mu <- ddply(plot_df, "sex", summarise, grp.mean=mean(pds_tot_all))
head(mu)

(pds_plot <- ggplot(main_df, aes(x=pds_tot_all, fill=sex, group=sex)) +
  geom_density(adjust=4, alpha=.4) +
     geom_vline(data=mu, aes(xintercept=grp.mean, color=sex),
             linetype="dashed") +
  theme_classic()) +
  labs(title=" Perceived pubertal development density curve",x="Pubertal Development Scale (total score)", y = "density") +
  scale_fill_brewer(palette="Dark2") 

#CBCL depression
mu <- ddply(plot_df, "sex", summarise, grp.mean=mean(cbcl_withdep_y3))
head(mu)

(dep_plot <- ggplot(main_df, aes(x=cbcl_withdep_y3, fill=sex, group=sex)) +
  geom_density(adjust=4, alpha=.4) +
     geom_vline(data=mu, aes(xintercept=grp.mean, color=sex),
             linetype="dashed") + 
    xlim(0,15) +
  theme_classic()) +
   labs(title="Depressive symptoms density curve",x="CBCL Withdrawn Depressed (raw score)", y = "density") +
    scale_fill_brewer(palette="Dark2")
 
#BMI 
mu <- ddply(plot_df, "sex", summarise, grp.mean=mean(bmi_y0))
head(mu)

(bmi_plot <- ggplot(main_df, aes(x=bmi_y0, fill=sex, group=sex)) +
  geom_density(adjust=4, alpha=.4) +
     geom_vline(data=mu, aes(xintercept=grp.mean, color=sex),
             linetype="dashed") +
  theme_classic()) +
  xlab("Body Mass Index (BMI)") +
  ylab("density") +
   labs(title="BMI density curve",x="Body Mass Index (BMI)", y = "density") +
    scale_fill_brewer(palette="Dark2")


des_fig_2 <- 
#display plots together
(des_fig <- ggarrange(
  pds_plot, 
  ggarrange(dep_plot, bmi_plot, ncol= 2, labels = c("B", "C")),
               nrow = 2,
            labels = "A")
            )

ggarrange(
  lp,                # First row with line plot
  # Second row with box and dot plots
  ggarrange(bxp, dp, ncol = 2, labels = c("B", "C")), 
  nrow = 2, 
  labels = "A"       # Label of the line plot
  ) 


```

##Waffle charts for Race and income  - waffle charts too complicated. 

Stick to bar chats 

```{r, bar charts}
# Data transformation
plot_df$race_y0 <- as.character(plot_df$race_y0)

race <- plot_df %>% 
  select(race_y0) %>% #select race variable only to get correct freq. 
  dplyr::group_by(race_y0) %>% 
  count() %>% 
dplyr::ungroup() %>% 
  mutate(perc = `freq` / sum(`freq`)) %>%
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

race_chart <- ggplot(race, aes(x = "", y = perc, fill = (race_y0)) +
  geom_col() +
  geom_text(aes(label = ""),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  theme_void()
)





ggplot(race, aes(x = "", y = perc, fill = (race_y0)) +
       geom_bar(width = 1, stat = "identity") +
       coord_polar("y", start = 0) +
       geom_label_repel(aes(label = perc), size=5, show.legend = F, nudge_x = 1) +
       guides(fill = guide_legend(title = "Group"))
              )
       

###code working above this####


ggplot(race, aes(values = perc, fill = race_y0)) +
  geom_waffle(n_rows = 8, size = 0.33, colour = "white") +
  scale_fill_manual(name = NULL,
                    values = perc,
                    labels = label) +
  coord_equal() +
  theme_void()

  geom_text(aes(label = labels),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y")
  
  



```


